### 2018-05-22 projects/dullemond/paper_dusttraps/ms.tex
Zeroth version of dust trapping paper, including analytic dust trap model and some thoughts about the streaming instability and whether we can find observational evidence for it.

%\documentclass[referee]{aa}
\documentclass{aa}
%\documentclass{mn2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\def\remark#1{{{\bf remark:} \bf #1}}
\def\action#1{{\bf #1}}
\def\putin#1{{\it #1}}
\def\revised#1{{\bf #1}}
\newcommand{\todo}[1]{\textcolor{red}{\bf #1}}
%
\def\aap{A\&A}
\def\apss{A\&SS}
\def\pasp{Proc.Astr.Soc.Pacific}
\def\pasj{Publ.Astr.Soc.Japan}
\def\araa{AnnRevA\& A}
\def\mnras{MNRAS}
\def\icarus{Icarus}
\def\nat{Nature}
\def\apj{ApJ}
\def\aj{AJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\jqsrt{JQSRT}
\def\max{\mathrm{max}}
\def\erg{\hbox{erg}}
\def\sec{\hbox{s}}
\def\Hz{\hbox{Hz}}
\def\gram{\hbox{g}}
\def\cm{\hbox{cm}}
\def\km{\hbox{km}}
\def\kel{\hbox{K}}
\def\KeV{\hbox{KeV}}
\def\MeV{\hbox{MeV}}
\def\eV{\hbox{eV}}
\def\yr{\hbox{yr}}
\def\AU{\hbox{AU}}
\def\ster{\hbox{ster}}
%\def\in{\mathrm{in}}
\def\out{\mathrm{out}}
\def\eff{\mathrm{eff}}
\def\rad{\mathrm{rad}}
\def\vrt{\mathrm{vert}}
\def\vset{\ensuremath{v_\mathrm{sett}}\xspace}
\def\fact{\ensuremath{\xi}\xspace}
\def\cs{\ensuremath{c_\mathrm{s}}\xspace}
\def\siggas{\ensuremath{\sigma_\mathrm{g}}\xspace}
\def\sigcoag{\ensuremath{\sigma_\mathrm{c}}\xspace}
\def\Tgas{\ensuremath{T_\mathrm{gas}}\xspace}
%\def\rhogas{\ensuremath{\rho_\mathrm{gas}}\xspace}
\def\rhogas{\ensuremath{\rho}\xspace}
\def\rhodust{\ensuremath{\rho_\mathrm{d}}\xspace}
\def\mugas{\ensuremath{\mu_\mathrm{gas}}\xspace}
\def\mp{\ensuremath{m_\mathrm{p}}\xspace}
\def\Hp{\ensuremath{H_\mathrm{p}}\xspace}
\def\Hs{\ensuremath{H_\mathrm{s}}\xspace}
\def\Omegak{\ensuremath{\Omega_\mathrm{K}}\xspace}
\def\Mstar{\ensuremath{M_\star}\xspace}
\def\Sc{\ensuremath{\mathrm{Sc}}\xspace}
\def\St{\ensuremath{\mathrm{St}}\xspace}
\def\Pt{\ensuremath{\mathrm{Pt}}\xspace}
\def\comma{\,,}
\def\fullstop{\,.}
%
\def\submitted{submitted\xspace}
\def\inprep{in prep.\xspace}
%
\begin{document}
%\thesaurus{02.01.2,08.03.4,08.06.2,08.16.5,13.09.6}
\title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
\authorrunning{Dullemond, Birnstiel, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
\institute{
  (1) Zentrum f\"ur Astronomie, Heidelberg University, Albert Ueberle Str.~2, 69120 Heidelberg, Germany\\
  (2) {\bf XXXX}
  (3) {\bf XXXX}
  (4) {\bf XXXX}
  (5) {\bf XXXX}
  (6) {\bf XXXX}
} \date{\today}

\abstract{A large fraction of the protoplanetary disks observed with the ALMA
  Large Programme {\bf XXXX} display multiple well-defined and nearly perfectly
  circular rings in the continuum, in many cases with substantial peak-to-valley
  contrast. Several of these rings are very narrow in radial extent.
  In this paper we analyze these results using the assumption that
  these dust rings are caused by dust trapping in radial pressure bumps. We
  model this process in a 1-D axisymmetric way, initially with a simple analytic
  model, then with a more detailed numerical model. We find that .....
%  including time-dependent viscous disk evolution coupled to time-dependent
%  radial dust drift. We perform model calculations without and with dust growth,
%  the latter with a simplified dust growth recipe. We test several scenarios
%  for the origin of the pressure bumps, including planetary gaps and dead zones.
{\bf CONCLUSIONS}
}

\maketitle

\begin{keywords}
accretion, accretion disks -- circumstellar matter 
-- stars: formation, pre-main-sequence -- infrared: stars 
\end{keywords}

\section{Introduction}
The concept of dust trapping in local pressure maxima has become a central theme
in studies of planet formation and protoplanetary disk evolution, as it may
provide an elegant solution to several problems in these fields of
study. Theories of planet formation are plagued by the ``radial drift barrier'':
the problem that as dust aggregates grow by coagulation, they tend to radially
drift toward the star before they reach planetsimal size
\citep[e.g.][]{2010A&A...513A..79B}. A natural solution to this problem could be
the trapping of dust particles in local pressure maxima
\citep{1972fpp..conf..211W, 2007ApJ...664L..55K, 1995A&A...295L...1B,
  1997Icar..128..213K}. From an observational perspective, the radial drift
problem manifests itself by the presence of large grains in the outer regions of
protoplanetary disks \citep{2003A&A...403..323T, 2009ApJ...700.1502A}, which
appears to be in conflict with theoretical predictions
\citep{2007A&A...469.1169B}. One possible solution to this observational
conundrum could be that the disks are much more massive in the gas than
previously suspected, leading to a higher gas friction for millimeter grains and
thus longer drift times \citep{2017ApJ...840...93P}. But another explanation is
to invoke, again, pressure traps. The most striking observational evidence for
dust trapping seems to come from large transitional disks, which feature giant
dust rings, sometimes lopsided, in which large quantities of dust appears to be
concentrated \citep{2013Natur.493..191C, 2013Sci...340.1199V}. These observations
appear to be well explained by the dust trapping scenario
\citet{2012A&A...545A..81P}. But these transitional disks seem to be rather
violent environments, possibly with strong warps \citep{2015ApJ...798L..44M,
  2017A&A...597A..42B} and companion-induced spirals
\citep{2016ApJ...816L..12D}. For more ``normal'' protoplanetary disk the dust
traps would have to be more subtle. \cite{2012A&A...538A.114P} explored the
possibility that the disk features many axisymmetric local pressure maxima, and
calculated how the dust drift and growth would behave under such conditions. It
was found that, if the pressure bumps are strong enough, the dust trapping can
keep a sufficient fraction of the dust mass at large distances from the star to
explain the observed dust millimeter flux. It would leave, however, a detectable
pattern of rings that should be discernable with ALMA observations.  Since the
multi-ringed disk observation of HL Tau \citep{2015ApJ...808L...3A} a number of
such multi-ringed disks have been detected \citep{2016ApJ...820L..40A,
  2016PhRvL.117y1101I, 2017ApJ...851L..23C, 2017A&A...600A..72F,
  2018A&A...610A..24F}. It is therefore very tempting to see also these
multi-ringed disks as evidence for dust trapping, and as an explanation for
the retention of dust in the outer regions of protoplanetary disks.

The data from the ALMA Large Programme 2016.1.00484.L offers an exciting new
opportunity to put this concept to the test, and to put constraints on the
physics of dust trapping in axisymmetric pressure maxima. This is an opportunity
which we explore in this paper.

If we assume that the rings seen in our data are caused by dust trapping, the
question arises: what constraints do the data impose on the physics of dust
trapping? To explore this we start by studying the rings individually, assuming
that the dust does not escape from the ring. This makes it possible to look for
a steady-state dust trapping model in which the radial drift forces (that push
the dust to the pressure peak) are balanced by turbulent mixing (that tends to
smear out the dust away from the pressure peak). In Section
\ref{sec-steady-state-analytic-trap-model} we will construct a very simplified
analytic dust trapping model, and confront this with the most well-isolated
rings from our sample. {\bf Describe the next sections.} But before we do the
modeling, we review, in Section \ref{sec-data}, the key aspects of the data from
our ALMA Large Programme that we compare our models to.

\section{The high-contrast rings of AS 209, Elias 24, GW Lup, HD 143006 and HD 163296}
\label{sec-data}
%
In this paper we discuss in particular those sources of the ALMA Large Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry about 3-D
lie-of-sight aspects. These are AS 209, Elias 24, GW Lup, HD 143006 and HD
163296. The high-contrast rings of these sources provide ``clean laboratories''
for testing the theory of dust trapping, both in a ring-by-ring manner, as well
as by comparing to global dust drift models. The most striking case is AS 209,
which features two high-contrast radially thin rings which are isolated and
large ({\bf see Guzm\'an et al.~(2018) for the detailed description of this
  dataset}). Fig.\ref{fig-obs-as209} shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the disk around AS 209. The
brightness is expressed as linear brightness temperature
($T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$), where $I_\nu$ is the
intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
frequency. It should be noted that the full (non-linear) brightness temperature
(defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
Planck function) can have very different values, in particular at low
brightness. As a result $T_{\mathrm{bright,full}}(r)$ tends to vastly
underestimate the contrast between the rings and the inter-ring
regions. Therefore, in this paper we plot the brightness using the linear
brightness temperature.

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{as209_gauss_modelfit.pdf}}
\caption{\label{fig-obs-as209}The linear brightness temperature in band 6 of the
  disk around AS 209, as a function of deprojected distance to the star (see
  {\bf Guzm\'an et al.~(2018)} for a detailed description and analysis of these data). The
  vertical axis is logarithmic to better show the contrast. The two most
  prominent rings are fitted by Gaussian rings. Ring 1 is fitted by
  $T_{\mathrm{bright,linear}}(r)=
  3.5\,\exp(-0.5(r-77\,\mathrm{au})^2/(4.0\,\mathrm{au})^2)$, while ring 2 is fitted by
  $T_{\mathrm{bright,linear}}(r)=
  2.8\,\exp(-0.5(r-124.7\,\mathrm{au})^2/(4.4\,\mathrm{au})^2)$.}
\end{figure}

The two most prominent rings (seen as bumps in Fig.\ref{fig-obs-as209}) are
fitted by Gaussian profiles with a standard deviation width of
$4.0\,\mathrm{au}$ and $4.4\,\mathrm{au}$ respectively. This corresponds to a
full-width-at-half-maximum (FWHM) of $9.3\,\mathrm{au}$ and $10.3\,\mathrm{au}$.
The ALMA beam FWHM for these observations is roughly $4.4\,\mathrm{au}$, meaning
that the rings are 2.1 and 2.3 beams wide.

These Gaussian profiles fit the observed data remarkably well, even in the wings
of the profile (except the outer wing of the outer ring). This is somewhat
surprising, given that the rings are resolved. The Gauss profiles therefore do
not represent the beam profile, but tell something about the actual dust
distribution within the ring.

The source Elias 24 also shows a prominent high-contrast isolated ring,
albeit flanked on the outside by a plateau of emission (see {\bf Andrews
  et al.~(2018) for a detailed description of this dataset}).

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{elias24_gauss_modelfit.pdf}}
\caption{\label{fig-obs-elias24}As Fig.~\ref{fig-obs-as209}, but now for the
  source Elias 24 (see {\bf Andrews et al.~(2018)} for a detailed description
  and analysis of these data). The most prominent ring is fitted by a Gaussian
  profile $T_{\mathrm{bright,linear}}(r)=
  5.3\,\exp(-0.5(r-79.9\,\mathrm{au})^2/(5.0\,\mathrm{au})^2)$.}
\end{figure}

The ring can be reasonably well fitted with a Gaussian profile with standard
deviation width of $5.0\,\mathrm{au}$ corresponding to a FWHM of
$11.7\,\mathrm{au}$. Given a beam size in band 6 of about $4.9\,\mathrm{au}$,
the ring of Elias 24 is also resolved by about 2.4 beam widths. The Gaussian fit
is, however, less convincing in the case of Elias 24 than in the case of AS 209,
so that it cannot be clearly argued that the ring has a Gaussian shape in the
case of Elias 24.

{\bf [TODO: GW Lup, HD 143006 and HD 163296]}

\section{Steady-state dust distribution in a ringlike trap}
\label{sec-steady-state-analytic-trap-model}
%
Given the nicely separated narrow rings in the above sources, we can analyze
these rings on a one-by-one basis. This is what we will do in this Section. We
will assume that the dust trapping in these rings is absolute: no dust escapes.
But increased turbulent mixing can spread the dust over a broader ring. Before
we start constructing a model, we will discuss an important feature of the
observed rings: that they appear to be radially resolved, Gaussian in shape, yet
sub-Planckian in brightness. We will present a simple analytic dust trapping
model, and show which radial brightness profiles it predicts, depending on the
dust size distribution, the turbulent strength and the gas pressure bump width.
This, when compared to the data, will allow us to draw some conclusions about
the size distribution {\bf [and perhaps more?]}.

\subsection{Resolved ring width and the brightness temperature}
If our finding is correct that the rings in our sample are all spatially
resolved in radial direction, then this poses an interesting question. In many
cases the brightness of these rings is arguably sub-Planckian. This is, for
instance, the case for the high-contrast rings of AS 209, as well as {\bf
  XXXXXXX}. For the outer two rings of AS 209 (at 77 au and 125 au,
respectively, the non-linear brightness temperatures are $3.5\,\mathrm{K}$ and
$2.8\,\mathrm{K}$, respectively. If these rings would have been {\em
  un}resolved, then this sub-Planckian emission is easily explained by the ring
not filling the beam. But because these rings are radially resolved, this
scenario should be rejected. The remaining possibilities are that the disk is
colder than we think or that the rings are optically thin. In the latter case
the spectral slope $\alpha_{\mathrm{spec}}$ at the location of these rings, once
measured, will determine if the grains that are trapped in this ring are
smaller, similar or larger than about $\lambda/2\pi$, where $\lambda$ is the
wavelength around which the slope is measured. If we use
$\lambda=0.13\,\mathrm{cm}$ and $\lambda=0.31\,\mathrm{cm}$ (representing ALMA
band 6 and 3, respectively) to determine $\alpha_{\mathrm{spec}}$ this would
mean we can distinguish between grains smaller, similar or larger than about 300
$\mu$m. A low $\alpha_{\mathrm{spec}}$ would imply grains that are substantially
larger than 300 $\mu$m, i.e.~of the order of millimeter or larger.

In several recent observations of the spectral index across ringed disks
\citep{2015ApJ...808L...3A,2018ApJ...852..122H} one clearly sees that
$\alpha_{\mathrm{spec}}$ varies across these rings, being closer to $2$ at the
ring center and substantially larger between the rings. This makes sense in
terms of the dust trapping scenario in which we expect larger grains to be
trapped more efficiently (and thus dominate the peak of the ring) than smaller
grains, because the smaller grains will be more subject to turbulent
mixing. This scenario requires a grain size distribution, so that the width of
the dust ring is smaller for the bigger grains and bigger for the smaller ones.

While this scenario is plausible and appealing, one may worry that it requires
some fine-tuning. Why would the rings be dominated by a population of grains
that are larger than 300 $\mu$m, but not too large, for else they would be so
effectively trapped that they would produce an even narrower ring? For one or
two sources that could be coincidence, but for a multitude of sources this is
less plausible. One possible explanation is that millimeter grains are indeed
the largest grains in the outer regions protoplanetary disks, because larger
grains would easily get destroyed and/or simply bounce instead of stick when
they collide, and thus not grow further \citep{2010A&A...513A..56G,
  2010A&A...513A..57Z}. These models also predict larger grains to drift quickly
inward, but that is only valid for a model without dust traps. At any rate:
perhaps millimeter grains are a natural consequence of dust growth. If so, it
then remains to be understood why the combination of turbulent strength and gas
surface density conspire to trap these millimeter grains, but not too much, for
else we might expect at least some rings to be narrower than a pressure scale
height. We are aware that, even with a sample of 20 sources, we are still in the
small-number-statistics regime, so these considerations are limited in their
strength.

In this section we will study this question using a very simple model of a dust
trap: a parameterized gaussian pressure profile. The goal is to find out if the
variations of $\alpha_{\mathrm{spec}}$ across the ring are naturally produced,
while still keeping the dust ring radially resolved, without having to fine-tune
the size distribution of the grains, the turbulence of the gas or other
parameters. We will do this in several steps, starting with analytic solutions
in Subsection \ref{sec-analytic-trap}, and proceeding with DISKLAB models in
Subsection \ref{sec-disklab-trap}.

\subsection{Analytic approximate solution of dust trapping}
\label{sec-analytic-model-of-trapping}
%
Let us consider a narrow gas ring around the star at radius $r_0$ with a
midplane pressure given by
\begin{equation}\label{eq-gaussian-pressure-bump}
p(r) = p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right)
\end{equation}
where $w\ll r_0$ is the parameter setting the width of this gaussian ring.
We assume that the gas is weakly turbulent with turbulent diffusion coefficient
$D$. Dust grains get trapped in this ring, and the dust will acquire a
radial density profile that is in equilibrium between the radial dust drift
pointing toward the peak of the gas pressure and radial turbulent diffusion
pointing away from that position. The radial dust drift velocity is
\citep[see e.g.][]{}:
\begin{equation}\label{eq-v-radial-drift}
  v_{\mathrm{dr}} = \frac{1}{\mathrm{St}+\mathrm{St}^{-1}}
  \left(\frac{d\ln p}{d\ln r}\right)\frac{c_s^2}{\Omega_Kr}
\end{equation}
where $c_s$ is the isothermal sound speed and the Stokes number $\mathrm{St}$ is
defined as
\begin{equation}
\mathrm{St} = \Omega_Kt_{\mathrm{stop}}
\end{equation}
where $t_{\mathrm{stop}}$ is the stopping time of the grains. The diffusion
coefficient for the dust is \citep{2007Icar..192..588Y}:
\begin{equation}
D_{\mathrm{d}} = \frac{D}{1+\mathrm{St}^2}
\end{equation}
We take $D$ to be equal to the turbulent viscosity $\nu$ divided by the
Schmidt number $\mathrm{Sc}$, which we usually set to $\mathrm{Sc}=1$.
We use the usual $\alpha$-prescription for the turbulence:
\begin{equation}
D=\frac{\nu}{\mathrm{Sc}}= \alpha_{\mathrm{turb}}\frac{c_s^2}{\mathrm{Sc}\,\Omega_K}
\end{equation}
If $D$ is sufficiently small, the dust will get concentrated into a ring with
width $w_d$ that is substantially smaller than the width of the gas ring
$w$. Also the dust will all be settled vertically into a layer that has a
vertical thickness less than the gas pressure scale height. In that case the
solution to the drift-mixing equation for the dust becomes simple, because
we can then assume everything to be constant with $r$ (within this limited
domain), except for the gas pressure gradient. We will also ignore any terms
arising from the curvature of the coordinates. The steady-state radial
dift-mixing equation for the dust then becomes, in its approximate form:
\begin{equation}
  \frac{d}{dr}
  \left(\Sigma_{\mathrm{d}}v_{\mathrm{dr}}-D_{\mathrm{d}}\frac{d\Sigma_{\mathrm{d}}}{dr}\right) = 0
\end{equation}
Integrating this equation once, with integration constant zero (which
amounts to a zero net radial flux), yields
\begin{equation}\label{eq-drift-mix-equil-eq}
  \Sigma_{\mathrm{d}}v_{\mathrm{dr}} = D_{\mathrm{d}}\frac{d\Sigma_{\mathrm{d}}}{dr}
\end{equation}
From Eqs.(\ref{eq-v-radial-drift},\ref{eq-gaussian-pressure-bump}) we can express
$v_{\mathrm{dr}}$ as
\begin{equation}
v_{\mathrm{dr}} = -\left(\frac{c_s^2}{w^2\Omega_K(\mathrm{St}+\mathrm{St}^{-1})}\right)(r-r_0)
\end{equation}
With this expression we can solve Eq.~(\ref{eq-drift-mix-equil-eq}) for
$\Sigma_{\mathrm{d}}$:
\begin{equation}\label{eq-analytic-sol-radial-trapping}
\Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}^2}\right)
\end{equation}
with 
\begin{equation}
  w_{\mathrm{d}} = w \sqrt{\frac{\Omega_KD_{\mathrm{d}}(\mathrm{St}+\mathrm{St}^{-1})}{c_s^2}}
  = w\sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
\end{equation}
Note that this is only valid as long as $\alpha_{\mathrm{turb}}\ll \mathrm{Sc}\,\mathrm{St}$.
This solution is, in fact, the radial version of the vertical settling-mixing
equilibrium solutions of \citet{1995Icar..114..237D}.

Now let us introduce a grain size distribution. Since the trapping width
$w_{\mathrm{d}}$ for the different grain sizes is different, we can only
impose the grain size distribution for the total (radially integrated)
dust mass. We define the size distribution according to the following powerlaw
\begin{equation}
m(a)\frac{dN}{d\ln a} = \frac{dM}{d\ln a} \propto a^{q}
\end{equation}
where $a$ is the grain size, $m(a)$ the corresponding grain mass, $N$ the
cumulative particle number and $M$ the cumulative dust mass. The parameter $q$
is the size distribution powerlaw coefficient, and it is $q=1/2$ for the usual
MRN distribution (this corresponds to $dN/da\propto a^{-3.5}$). We also need to
define limits $a_{\mathrm{min}}$ and $a_{\mathrm{max}}$. The size distribution
is then normalized such that its integral over $d\ln(a)$ is the total dust mass
$M$. The radial surface density solution,
Eq.~(\ref{eq-analytic-sol-radial-trapping}), then becomes:
\begin{equation}\label{eq-analytic-sol-radial-trapping-sizedistr}
\frac{d\Sigma_{\mathrm{d}}(r)}{d\ln a} = \frac{1}{(2\pi)^{3/2}r_0
    w_{\mathrm{d}}(a)}\frac{dM}{d\ln a}
\exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}(a)^2}\right)
\end{equation}

\subsection{Resulting profiles for a size distribution}
\label{sec-anmodel-profiles}
%
The analytic model of the dust trapping for a single grain size speaks for
itself. But how does it look for a grain size distribution? In
Fig.~\ref{fig-anmodel-distribution-sigma} we show the radial profiles for an MRN size
distribution between $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and
$a_{\mathrm{min}}=1\,\mathrm{cm}$ and a total dust mass of $M=1.3\times
10^{-5}\,M_\odot$. The dust material density is taken to be
$\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
distribution with 10 grain sizes, evenly distributed in $\ln(a)$. These are the
10 curves in Fig.~\ref{fig-anmodel-distribution-sigma}. These surface densities are
defined as being the local grain size distribution integrated over the width of
the grain size bin:
$\Sigma_{\mathrm{bin}}=\int_{\mathrm{bin}}(d\Sigma_{\mathrm{dust}}/da)da$. The
stellar mass is $M_{*}=0.9\,M_\odot$ and we set $\mathrm{Sc}=1$.

{\bf MUST ALSO SPECIFY GAS DENSITY, TO GET STOKES NUMBERS. MUST SAY
(AND CHECK) THAT EPSTEIN REGIME.}

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_sigma_profiles.pdf}}
\caption{\label{fig-anmodel-distribution-sigma}Results of the analytic dust
  trapping model for an MRN powerlaw grain size distribution between
  $a=0.01\,\mathrm{cm}$ and $a=1\,\mathrm{cm}$. The gas pressure bump has a FWHM
  of 70 au and a peak at $r=77\,\mathrm{au}$. Shown here are the surface density
  profiles for 10 logarithmically binned grain sizes. The curve with the
  narrowest peak corresponds to a grain size of $a=1\,\mathrm{cm}$, while that
  with the widest peak (and lowest maximum value) corresponds to a grain size of
  $a=0.01\,\mathrm{cm}$. The intermediate curves correspond to $a=0.0167$,
  0.0278, 0.0464, 0.0774, 0.1291, 0.2154, 0.3593 and 0.5995 $\mathrm{cm}$
  respectively. The dashed curve shows the total surface density of the dust,
  which is, by definition of the bin-integrated surface density, the sum of all
  10 curves. For a description of the model setup and model parameters, see
  Section \ref{sec-anmodel-profiles}.}
\end{figure}

As expected, the largest grains are concentrated the most and the smallest ones
the least. The largest grains, however, have the largest peak value of their
surface density at the exact location of the gas pressure peak.

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_tau_profiles.pdf}}
\caption{\label{fig-anmodel-distribution-tau}As Fig.~\ref{fig-anmodel-distribution-sigma},
  but now the vertical optical depth at $\lambda=1.3\,\mathrm{mm}$ (ALMA band 6)
  is shown. The opacity model of \citet{1997MNRAS.291..121I} was used to compute these
from the size distribution.} 
\end{figure}

When it comes to the contribution to the optical depth, however, the situation
looks a bit different. Let us assume a very simple dust opacity law, which
roughly mimicks the real behavior of dust opacities: the model of
\citet{1997MNRAS.291..121I}, where we take the scattering opacity zero for
simplicity. In this simple but illustrative opacity model the cross
section of a grain with radius $a$ is the geometric cross section $\pi a^2$
for $2\pi a>\lambda$ (with $\lambda$ the wavelength), while it is reduced by
a factor $(2\pi a/\lambda)$ for $2\pi a<\lambda$. This means that for
$2\pi a>\lambda$ the opacity (that is, the cross section per unit mass)
is $\kappa=3/4\xi_{\mathrm{dust}} a$ where $\xi_{\mathrm{dust}}$ is the material
density, here taken to be $\xi_{\mathrm{dust}}=2\,\mathrm{g}/\cm^3$. For
$2\pi a<\lambda$ this is reduced by a factor $(2\pi a/\lambda)$ and
thus becomes independent of $a$.

It can be seen from Fig.~\ref{fig-anmodel-distribution-tau} that at the peak of
the trap all grain sizes contribute equally to the optical depth. That is
because, with our opacity model, the larger grains (the ones with $2\pi a >
\lambda$) have an opacity equal to their geometric cross section per unit dust
mass, which goes as $1/a$. The smallest grains (the ones with $a=0.01$ and
$a=0.167$) have $2\pi a < \lambda$, which means that their opacity remains
constant with $a$. Hence, for these two curves in
Fig.~\ref{fig-anmodel-distribution-tau} the contribution to the peak optical
depth drops below those of the other grains. Of course, with a more realistic
opacity model this can change somewhat, but the rough principle remains valid.

More important is to notice that the relative contributions of the grain sizes
to the peak optical depth may change if a different grain size distribution is
chosen. Here we chose the MRN size distribution ($q=1/2$), but that was a rather
arbitrary choice. Given the complexity of the dust growth processes in
protoplanetary disks, it is not at all said that the MRN distribution should be
considered a preferred one. The reader may experiment with other values of
$q$ or entirely different grain size distribution shapes altogether.

\subsection{Effect of the degree of trapping on the appearance of the ring}
\label{sec-anmodel-degree-trapping-alpha-series}
%
With this simple analytic dust trapping model we can investigate how the degree
of trapping affects the appearance of the dust ring. For a model with a single
grain size this is easy to predict: if the turbulence is weak enough that the
gaussian profile of $\Sigma_{\mathrm{d}}(r)$ is narrower than the ALMA beam,
then we observe a gaussian profile with the width of the beam. Even if the dust
in this narrow ring would be optically thick, the observed ring would have
sub-Planckian brightness due to beam-dilution. For stronger turbulence,
$\Sigma_{\mathrm{d}}(r)$ is spread out wider, and we observe a gaussian profile
with the width of the dust profile itself. As in the case of the strongly
trapped dust, also the wide rust ring has a Gaussian profile, except that if it
is optically thick, the observed profile is flattened at the top due to
saturation to the Planck function.

The situation gets more complex if we have a grain size distribution. For a
given turbulence strength $\alpha_{\mathrm{turb}}$ some grains may be
concentrated into a ring that is narrower than the ALMA beam, while others may
be more spread out. The strongly concentrated grain population produces a cusp
in the brightness, which will be partly smeared out by the beam convolution,
while the weakly concentrated grain population produces a broad underlying
brightness bump.

To see how this looks in practice we set up a realization of our analytic
model. The peak of the pressure is put at $r_{0}=77\,\mathrm{au}$ (corresponding
to the inner of the two main rings of AS 209), and the width is chosen to be
$w=30\,\mathrm{au}$ (corresponding to a FWHM of 70 au).  Given that the ALMA
beam in band 6 has a FWHM of about 35 milliarcseconds, corresponding to 4.4 au
for the distance to AS 209, the gas bump is much wider than the beam. The two
main dust rings of AS 209 are about 3 beams wide, so that for our wide gas bump
model the dust is strongly concentrated. This requires a low turbulent
$\alpha_{\mathrm{turb}}$ or large grains. As in Section
\ref{sec-anmodel-profiles} we choose a grain size distribution between
$a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and $a_{\mathrm{min}}=1\,\mathrm{cm}$,
following the usual MRN powerlaw of $q=1/2$ and a total dust mass of
$M=1.3\times 10^{-5}\,M_\odot$. The dust material density is taken to be
$\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
distribution with 10 grain sizes, evenly distributed in $\ln(a)$. For the disk
temperature we choose a simple powerlaw $T(r)=200\,(r/\mathrm{au})^{-1/2}$. The
stellar mass is $M_{*}=0.9\,M_\odot$, and the source distance is taken to be
$d=126\,\mathrm{pc}$. All these setting are taken to represent the inner of the
two main high-contrast rings of the source AS 209. We set $\mathrm{Sc}=1$.

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_series_alpha.pdf}}
\caption{\label{fig-anmodel-series-alpha}Results of the analytic dust trapping
  model for various turbulent strengths, for an MRN powerlaw grain size
  distribution between $a=0.01\,\mathrm{cm}$ and $a=1\,\mathrm{cm}$. The gas
  pressure bump has a FWHM of 70 au and a peak at $r=77\,\mathrm{au}$. Shown
  here is the predicted brightness temperature at a wavelength
  $\lambda=1.3\,\mathrm{mm}$ before (dashed) and after (solid) convolution with
  the ALMA beam, assuming a source distance of $d=126\,\mathrm{pc}$. For a
  description of the model setup and model parameters, see Section
  \ref{sec-anmodel-degree-trapping-alpha-series}.}
\end{figure}

The results for several values of $\alpha_{\mathrm{turb}}$ are shown in
Fig.~\ref{fig-anmodel-series-alpha}. In the top panel, representing weak
turbulence, most of the dust is trapped into a ring more narrow than the
beam. A clear cusp can be seen in the un-convolved brightness profile,
which is due to the strongly trapped large grains of the size distribution.
However, the convolution washes out this cusp for the most part. This
situation is equivalent to the simple case of a single dust species being
trapped to a narrow unresolved ring. The only difference is that for a single
dust size the ring is either optically thick or optically thin, while in the
multi-size case this can be a mix. That does not make a difference for the
observed brightness profile, which is a Gaussian either way.

For the case of the strongest turbulence, the bottom panel in
Fig.~\ref{fig-anmodel-series-alpha}, the cusp is still seen in the un-convolved
profile, but now the convolution does not wash it out completely. The resulting
convolved brightness profile is clearly non-Gaussian as a result. It has a
quasi-Gaussian cusp with broad wings. When seen with ALMA such a cuspy profile
(narrow at the peak, broad in the wings) should be recognizable.

These findings show that if the turbulent strength is weak, the predicted dust
ring width equals the ALMA beam width, and the brightness shape should have the
shape of the beam profile (here assumed to be Gaussian). For stronger
turbulence, the dust ring becomes indeed broader than the ALMA beam, but then
the shape of the observed brightness profile may teach us something about the
grain size distribution. Roughly speaking, if the largest grains of the size
distribution still get trapped in a narrow ring, this shows up as a cusp in the
brightness profile. If the smallest grains of the grain size distribution are
too small, the predicted brightness profile would be too broad compared to the
observed ring width. In other words: if the observed brightness profile of a
observed ring in given source is clearly radially resolved, and if it has a
Gaussian shape, then the most likely fit to our analytic model is that of
a narrow size distribution (quasi a single grain size) with the right turbulent
strength to fit the observed ring width. A broader size distribution, with the
appropriately chosen turbulent strength, would yield a cuspy profile, not a
Gaussian one.

The two high-contrast rings of AS 209 both appear to fit a Gaussian profile with
2.5 times the width of the ALMA beam (Fig.~\ref{fig-as209-rings-gauss-fit}). In
both cases it is therefore likely, according to this analysis, that the dust
that is trapped in these rings has a narrow size distribution instead of a broad
MRN powerlaw distribution.

\subsection{Spectral slope variations across the ring}
\label{sec-anmodel-slope-variations}
%

\subsection{Model degeneracies}
\label{sec-anmodel-degeneracies}
%
When trying to infer the effectiveness of dust trapping within the rings with
the analytic model of Section \ref{sec-analytic-model-of-trapping}, one notices
that there are degeneracies in the possible combinations of parameters. Most
obvious is the degeneracy between grain size and turbulence: smaller grains
and stronger turbulence may yield the same dust trapping width. But also the
width of the gas pressure bump can be compensated: a narrower bump but stronger
turbulence (or smaller grains) will yield the same ring width.

{\bf [CONTINUE THIS; SOME DEGENERACIES FOLLOW DIRECTLY FROM THE EQUATIONS.]}

\subsection{Model fits to the ALMA data}
\label{sec-anmodel-fits-to-data}
%
We now apply this model to the isolated rings in our ALMA sample (Section
\ref{sec-data}). Given the degeneracies of the model in many of the parameters,
we perform the fitting procedure in three parameters, while fixing the others
to a particular choice of values.

{\bf [CONTINUE THIS; SHOW THE FIT FIGURES; SHOW THE MCMC CORNER PLOTS]}

\subsubsection{Could the rings be optically thick at the center?}
\label{sec-anmodel-fits-optically-thick}
%
All rings in our sample show sub-Planckian brightness, at least for the disk
temperature assumed in our model. We have calculated the disk midplane temperature
using a simple ``flaring angle recipe''. If the disks are, however, substantially
colder (by up to a factor of 2.5 to 3) then the peak flux of the rings might be
consistent with saturated emission, i.e.\ emission expected from an optically thick
disk. So the question is: how sure are we that the rings are sub-Planckian?

To reduce the midplane temperature by a factor of $2$ the irradiation flux must
be reduced roughly by a factor of $2^4=16$. Within the flaring angle recipe this
would lead to unrealistically low flaring angle. Presently we assumed $\varphi=0.02$,
so that it would have to reduce to $\varphi=0.0013$, which is so low that we can no
longer trust the flaring angle recipe. However, maybe the outer disk lies in the
shadow of the inner disk. To see whether this can reduce the temperature to low
enough values, we perform a proper radiative transfer calculation. The results are
shown in Appendix ({\bf XXXX}). It shows that {\bf XXXXX}.

There is another issue, however. If the emission saturates to the optically thick
brightness, i.e.\ to the Planck function, then the radial profile of the ring should
be flat-topped. This is clearly not what we see in the rings in our sample, with
the possible exception of the ring of GW Lup which shows a very slight hint of that.
In general, however, it seems that the simple picture of collecting so much dust into
the ring that the ring becomes optically thick does not seem to explain the observed
radial brightness profiles. On the other hand, optical thickness can also occur
on smaller scales, as we will discuss in Section \ref{sec-si-clump}.

{\bf [TO BE DONE: So far the CMB or cloud background is not included in the model.
    This has to be improved, because we are working at temperatures close to the
    CMB. It might reduce the sub-Planckianness, I guess.]}

{\bf [Question to the experts: The CMB and/or cloud background is automatically
filtered out of the interferometric observations, I presume?]}

\section{Streaming instability, clumping, and the spectral index}
\label{sec-si-clump}
%
Whenever the dust trapping becomes so effective that the volume density of the
dust at the midplane of the disk exceeds the gas volume density, the streaming
instability may set in \citep{2005ApJ...620..459Y}. This produces turbulence and
leads to clumping \citep{2007ApJ...662..627J, 2008A&A...479..883L,
  2010ApJ...722.1437B}. If conditions are right, some of these clumps may
gravitationally collapse and form planetesimals \citep{2007Natur.448.1022J,
  2017ApJ...847L..12S, 2017A&A...597A..69S}. It is believed that this mechanism
is the trigger that starts the process of the formation of planets
\citep[see][for a review]{2014prpl.conf..547J}. If one could find observational
evidence of this dust clumping occurring in protoplanetary disks, this would
be a major step forward for our understanding of planet formation.

The clumps that are formed in simulations of the streaming instability are,
however, typically substantially smaller than a pressure scale height of the
disk \citep{2007ApJ...662..627J, 2013MNRAS.434.1460K}. This makes it very hard
to spatially resolve these structures, even with ALMA at its highest resolution.

It might, however, be possible to find indications of particle clumping in an
indirect way, if it leads to the formation of optically thick clouds of
particles surrounded by optically thin regions. Even if we cannot spatially
resolve these clouds, the spatially averaged thermal emission from such a clumpy
medium has a different spectral slope than that of a homogeneous distribution of
dust. If all the dust is concentrated in a multitude of small and very optically
thick clumps, then the brightness of this unresolved clumpy medium would be the
Planck function times the covering fraction $C$ of the clumps:
\begin{equation}\label{eq-bright-clumpy}
I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,C
\end{equation}
where the limit $C\rightarrow 0$ means concentrating all the dust in infinitely
dense clumps so that the covering fraction is zero, and $C\rightarrow 1$ means
having so many clumps that they multiply overlap, when seen in projection from
above. In contrast, if all this dust is homogeneously distributed, then the
brightness is
\begin{equation}\label{eq-bright-homogeneous}
I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,(1-e^{-\tau_\nu})
\end{equation}
with $\tau_\nu$ being the vertical optical depth of the dust in the disk.
Reality will likely lie in between these two extreme scenarios, with
optically thin dust interdispersed between optically thick clouds.

There are at least two ways by which one could use Eqs.~(\ref{eq-bright-clumpy},
\ref{eq-bright-homogeneous}) to search for hints of clumpiness. The most direct
way is to make use of the different dependencies of these two equations on
wavelength $\lambda=c/\nu$. In the case of the clumpy medium with perfectly
optically thick clumps (Eq.~\ref{eq-bright-clumpy}) the brightness follows the
Planck curve, whereas for the homogeneous medium
(Eq.~\ref{eq-bright-homogeneous}), if it is optically thin
(i.e.~$1-e^{-\tau_\nu}\simeq \tau_\nu$), it follows the Planck curve times the
opacity law. If we are in the Rayleigh-Jeans domain of the Planck curve, then
these two limits are usually referred to as spectral slope
$\alpha_{\mathrm{spec}}=2$ and $\alpha_{\mathrm{spec}}=2+\beta$, respectively,
where $\beta$ is defined by $\kappa_\nu\propto \nu^\beta$. At ALMA band 6, for
the low temperatures we expect in these disks at large radii, we are no longer
strictly in the Rayleigh-Jeans domain, so the spectral slope analysis becomes
temperature-dependent, which complicates matters a bit, but the principle
remains the same. If the streaming instability is operating in the rings of our
sources, and if the numbers work out (see below), then we expect to see
$\alpha_{\mathrm{spec}}\simeq 2$ within the rings, and
$\alpha_{\mathrm{spec}}>2$ between the rings.

There is an important caveat to this idea: the grain size and the corresponding
opacity slope may mimick the same spectral signatures as the clumping. This is,
in fact, the same degeneracy of the interpretation of the millimeter spectral
slope as usual \citep[e.g.][]{2003A&A...403..323T}, just on a much smaller
spatial scale: a scale that remains unresolved even by ALMA.  Nevertheless it is
worthwhile to investigate if we can find clues that point to clumpiness.

In the scope of this paper we will only do a very simple analysis: we will
verify if, under the most benign conditions, the above mentioned optically
thick/thin clumpy medium is at all possible. This is not granted, because
these rings are several tens of au large, and therefore it requires a lot of
mass to make dust clumps optically thick at 1.3 millimeter wavelength.
Furthermore, these dust clouds cannot have a density in excess of the
Roche density, because then they would immediately gravitationally collapse.
Therefore, we cannot make the clumps arbitrary dense and compact.

So let us construct the most optimistic model of a dust cloud. The highest
possible density of a cloud is the Roche density, defined by
\begin{equation}
\rho_{\mathrm{Roche}}(r) = \frac{9}{4\pi}\frac{M_{*}}{r^3}
\end{equation}
where $M_{*}$ is the stellar mass and $r$ is the distance to the star. We take
$\rho_{\mathrm{cloud}}=\rho_{\mathrm{Roche}}$. The simplest opacity model for a
dust grain is that of \cite{1997MNRAS.291..121I}, where we omit the scattering
part:
\begin{equation}
\kappa(a) = \frac{3}{4}\frac{1}{\xi\,a}\left\{\begin{matrix}
1 & \hbox{for} & \lambda<2\pi a \\
2\pi a/\lambda & \hbox{for} & \lambda>2\pi a \\
\end{matrix}\right.
\end{equation}
where $a$ is the grain radius and $\xi$ is the material density of the dust
grain ($\xi=3.6$ for amorphous olivine, $\xi=2.0$ for silica {\bf CHECK}). For a
given wavelength $\lambda$ (in our case $\lambda=0.13\,\mathrm{cm}$ for ALMA
band 6), the most optimistic opacity is therefore $\kappa=(3\pi/2)/\xi\lambda$,
for all $a\le \lambda/2\pi$. For ALMA band 6 and $\xi=2\,\mathrm{g/cm}^3$ this
is $\kappa=18.1\,\mathrm{cm}^2/\mathrm{g}$. The largest grain size to still
have this opacity is therefore $a=\lambda/2\pi$, which for ALMA band 6 is
$a=0.02\,\mathrm{cm}$.  We make a spherical cloud of this dust with radius
$R$. The optical depth through the cloud center is then
\begin{equation}
\tau = 2R\rho_{\mathrm{cloud}}\kappa
\end{equation}
To obtain the effect of an optically thick/thin clumpy medium, these clouds
have to be sufficiently optically thick, not just marginally. So let us demand
a minimal optical depth of $\tau=4$. This gives a minimal cloud radius of
\begin{equation}
  R\gtrsim \frac{2}{\rho_{\mathrm{Roche}}\kappa} = 0.11\frac{1}{\rho_{\mathrm{Roche}}}
  = 0.15\frac{r^3}{M_{*}}
\end{equation}
in CGS units (for ALMA band 6). Note that this scales linearly with
$\rho_{\mathrm{cloud}}/\rho_{\mathrm{Roche}}$, so that if we choose a much less
``critical'' cloud density of, say, $\rho_{\mathrm{cloud}}=0.1\,\rho_{\mathrm{Roche}}$,
then the cloud should become 10$\times$ larger and 100$\times$ more massive.

If we want these clouds to be part of an unresolved clumpy medium, they should
be substantially smaller than the pressure scale height of the disk, and
substantially smaller than the ALMA band 6 resolution. Let us take the example
of the source AS 209, with $M_{*}=0.9\,M_{\odot}$, and its two high-contrast
rings at 74 au and 120 au, respectively. This gives minimal cloud
radii of $R\gtrsim 4\times 10^{-3}\,\mathrm{au}$ and $R\gtrsim 22\times
10^{-3}\,\mathrm{au}$, respectively. Clouds of size similar to these minimal
values are indeed much smaller than the beam width, and are therefore
unresolved.  This does not mean that there cannot be any smaller
clumps/clouds. It only means that any smaller clouds would be optically thin,
and we would thus not see any effect of these clumps on the unresolved
brightness: these clouds would have a spectral slope equal to that of an
optically thin non-clumpy model. At any rate, this estimate shows that, at least
in principle, it is possible that the rings we see in our data consist, in fact,
of numerous unresolved optically thick clouds interdispersed with optically thin
dust (or without any dust in between). Given that such an unresolved, but
optically thick, set of clumps has a lower spectral slope
$\alpha_{\mathrm{spec}}$ than an optically thin dusty medium, an analysis of the
measured spectral slope variations over the radial profile of the rings {\em
  may} give us some clues as to whether clumpiness occurs or not in these
rings, even if we cannot resolve the clumps themselves. We would expect that
within the ring the spectral slope is closer to $\alpha_{\mathrm{spec}}\simeq 2$
(modulo deviations from Rayleigh-Jeans) than in the inter-ring region.

However, the grain size segregation mechanism, that larger grains are more
concentrated in the ring than smaller grains, also tends to produce this
signature. This is particularly the case if the big grains (which are closest to
the ring peak) are larger than $\lambda/2\pi$ while the small grains are smaller
than $\lambda/2\pi$. To distinguish between the roles of size seggregation and
clumping, we need an independent way to probe the grain size. Perhaps the effects
of self-scattering induced polarization \citep{2015ApJ...809...78K} may provide
such an avenue.

Even if we can distinguish between size-sorting and clumping, we still have to
clarify whether the required total dust mass for the clumping scenario is
reasonable, and whether the required 'filling factor' is consistent with
simulations of the streaming instability. Both questions are related. Let us go
back to the two prominent isolated rings of AS 209. The peak linear brightness
temperature is 3.5 $\mathrm{K}$ for the inner one and 2.7 $\mathrm{K}$ for the
outer one. In our model the disk midplane temperature is 13 $\mathrm{K}$ for the
inner and 10 $\mathrm{K}$ for the outer ring, which translates (for
$\lambda=0.13\,\mathrm{cm}$) into a linear brightness temperature of
8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, respectively. If the disk were optically
thick, we would thus expect to observe a linear brightness temperature of
8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, which are only about 2 times higher.

{\bf [TODO: Handle the CMB and cloud background, which might reduce the
sub-Planckianness we observe.]}

This means that the filling factor of these clouds, as projected vertically
toward the observer, must be quite high, of the order of 50\%. {\bf [At this
    point we can calculate the minimum mass in dust needed for this]}

In principle the disk could be warmer than we assume, leading to a larger
saturation brightness, and thus a lower required filling factor. But given that
a 2$\times$ higher disk temperature requires a 16$\times$ higher irradiation
flux, we expect at most a factor of 2 uncertainty in the required filling
factor.

Checking whether this filling factor is consistent with the expectation of
simulations of the streaming instability is not easy. Such data are not
published for the models describing the streaming instability so far.  We can
only do a ``by eye'' check of the figures in those papers ({\bf give a few
  example papers here}). Such an exercise gives us the impression that the
typical fraction of the simulation box that is covered by the dense cluster
regions, as projected vertically, is typically substantially less than 50\%. If
this impression is true, then the ring emission seen in our sample cannot be
explained by a set of spatially unresolved optically thick clouds. That does not
imply, however, that the streaming instability does not operate in these
rings. It only implies that if it does, it will be hidden below an optically
thick shroud of more extended dust.

\subsection{Under which conditions should the streaming instability set in?}
While in Section \ref{sec-si-clump} we investigated if we can recognize the
ongoing streaming instability in our data if it is taking place in the disk, in
this Section we will compute, with our analytic dust trapping model, under which
condition we would {\em expect} that the streaming instability will operate in
the first place. This depends on the total dust mass locked in the trap, the
grain sizes and the turbulent strength. To make analytic estimates it is easiest
to assume a single grain size, instead of a grain size distribution.

{\bf FINISH THIS SECTION}

\section{Discussion}

\subsection{Gap depth estimates from 2-D hydrodynamic simulations}

{\bf Dong \& Fung (2017)} presented a systematic study of the gap
opening profiles for given planetary masses, and analyzed the
rings seen in scattered light of 5 sources (HD 97048, TW Hydra,
HD 169142, LkCa15 and RX J1615).

\subsection{Condition for the streaming instability}
In the literature it is often mentioned that the streaming instability requires
a dust-to-gas surface density ratio of
$\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ or higher to operate {\bf
  XXXXXXXX}. This can, however, not be directly compared to our models, because
this value of $0.02$ was found for models without any pre-determined
turbulence. The turbulence in those models was induced by the streaming
instability itself or, if the streaming instability does not operate, by the
Kelvin-Helmholtz instability ({\bf XXXX}). In the analytic model of this
Section, on the other hand, we set the turbulence strength by hand, by setting
$\alpha_{\mathrm{turb}}$ to some value. In essence, we assume that there is
another source of turbulence, such as the magnetorotational instability ({\bf
  XXXX}) or the vertical shear instability ({\bf XXXX}), that determines the
mixing of the dust in the disk.

According to \citet{2005ApJ...620..459Y} the true criterion for the onset of the
streaming instability is the ratio of dust and gas {\em volume} densities
$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$. For a given surface density
ratio $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}$, the midplane volume density
ratio, for a single grain species with midplane Stokes number $\mathrm{St}\ll
1$, depends on the turbulent strength as
\begin{equation}\label{eq-dtg-sig-vs-rho}
  \frac{\rho_{\mathrm{d}}}{\rho_{\mathrm{g}}}\simeq
  \left(1+\frac{\mathrm{St}}{\alpha_{\mathrm{turb}}}\right)
  \frac{\Sigma_{\mathrm{d}}}{\Sigma_{\mathrm{g}}}
\end{equation}
The criterion of $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ mentioned
in the literature thus relates to the criterion
$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$ via the turbulent strength and
the Stokes number. Given that we do not compute the turbulent strength
but prescribe it, we should rely on the more fundamental volume density
criterion of \citet{2005ApJ...620..459Y} to assess whether the dust in our
model triggers the streaming instability or not.

In our analytic setup of Section \ref{sec-anmodel-degree-trapping-alpha-series},
the dust ring is about 5 times narrower than the gas bump. Assuming that
initially the dust-to-gas ratio was $0.01$, this implies a dust-to-gas surface
density ratio at the peak of the ring of $0.05$, which is in excess of the usual
$0.02$ value. But as long as $\mathrm{St}\lesssim 19\alpha_{\mathrm{turb}}$,
Eq.~(\ref{eq-dtg-sig-vs-rho}) shows that the midplane volume density ratio still
stays below unity and no streaming instability sets in.

\subsection{Dust trapping in the horseshoe region and in resonances}
The 1-D approach of the {\sf DISKLAB} models make it impossible to treat the
trapping of particles in the horseshoe region. The low-$\alpha$ models of {\bf
  Liu et al.~2018} show that this may be a significant effect. However, this
tends to produce non-axisymmetric rings: there will clearly be a dearth of dust
near the location of the planet, and a slight enhancement of dust near the L4
and L5 Lagrange points. Given that the rings we see so far in the multi-ringed
protoplanetary disks are angularly symmetric, this appears to be ruled out.
That means that the 1-D approach is justified, but we do have to ensure that we
are not in the regime where such non-axisymmetric horseshoe-trapping is
important. Generally the horseshoe region tends to gradually deplete over time,
but the time scale depends on the turbulent viscosity, with low-$\alpha$ disks
depleting slower.

Another issue is the potential for large particles to get trapped in a
resonance of a planet instead of in a classical dust trap {\bf REFERENCE;
the Dong double-gap is different, though}.
This requires high Stokes number, generally $\mathrm{St}\gg 1$ ({\bf is
  this true?}). Given that grain growth models tend to keep grains below
this limit ({\bf see e.g.~Birnstiel et al.}), this is less likely to play
a role.

\subsection{Multiple gaps with a single planet}
{\bf Dong, Li \& Li (2017)} and 
{\bf Bae, Zhu \& Hartmann (2017)} show that for very low-$\alpha$ disks
a planet can open multiple gaps in the disk rather than just a single one.
The contrast of these rings and gaps in the gas can be large (factor of 2; see
Bae et al.),
and thus clearly constitute dust traps. Dong et al.\ show that these gaps
can even be observable for very low mass planets.

{\bf [TODO: We can do a simple model of dust trapping for their case.]}

{\bf Fung \& Chiang (2017)} have shown that low mass planets in nearly inviscid
disks can avoid migration and can be responsible for disk accretion in spite
of the low viscosity.

\subsection{Interpretation of the millimeter flux in terms of grain size}
A common rule of thumb for interpreting millimeter and sub-millimeter
fluxes and intensities from thermal dust emission is that you ``observe
the grain size equal to the wavelength you observe at''. This rule is
based on the tendency of the dust opacity derived from a Mie calculation
to peak at wavelength around $\lambda\sim 2\pi a$, with $a$ the grain
radius. For $\lambda \ll 2\pi a$ the opacity becomes nearly constant
while $\lambda \gg 2\pi a$ it drops. This effect is even stronger for
the scattering opacity. The validity of this rule of thumb relies,
however, on the grain size distribution itself. For the simple
{\bf Ivezic et al.~(1997)} opacity law, it is only valid for
grain size distribution powerlaws obeying {\bf XXXXXXXXXXX}.
{\bf [TODO: Maybe here do an experiment with full opacity laws.]}

\begin{acknowledgements}
  The authors acknowledge support
  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
  Datenverarbeitung of the University of T\"ubingen, the state of
  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
  through grant no INST 37/935-1 FUGG. This research was initiated at the
  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
  April 2018. Part of this work was also funded by the DFG Forschergruppe
  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
\end{acknowledgements}

\begingroup
\bibliographystyle{aa}
\bibliography{ms}
\endgroup

\end{document}

### 2018-05-22 reduction_scripts/doar25_contselfcal_ai.py
Update doar25_contselfcal_ai.py

@@ -636,31 +636,18 @@ os.system('cp -r '+combined_cont_ap+'_backup.ms DoAr25_cont_final.ms')
 
 mscont = 'DoAr25_cont_final.ms'
 
-# Make a final image with robust=0 cleaning deeper
-cont_final = prefix+'_cont_final_r0'
-os.system('rm -rf %s*' % cont_final)
-tclean_wrapper(vis=mscont, imagename=cont_final, cellsize='0.01arcsec', imsize=[1000], mask = LB1_mask, scales = LB_scales, threshold = '0.03mJy', robust=0, interactive=True)
-estimate_SNR(cont_final+'.image', disk_mask = LB1_mask, noise_mask = noise_annulus)
-#DoAr25_cont_final_r0.image
-#Beam 0.042 arcsec x 0.021 arcsec (74.31 deg)
-#Flux inside disk mask: 227.07 mJy
-#Peak intensity of source: 1.39 mJy/beam
-#rms: 1.73e-02 mJy/beam
-#Peak SNR: 80.27
-
 # Make a final image with robust=0.5 cleaning deeper
 cont_final = prefix+'_cont_final_r0.5'
 os.system('rm -rf %s*' % cont_final)
-tclean_wrapper(vis=mscont, imagename=cont_final, cellsize='0.01arcsec', imsize=[1000], mask = LB1_mask, scales = LB_scales, threshold = '0.02mJy', robust=0.5, interactive=True)
+tclean_wrapper(vis=mscont, imagename=cont_final, cellsize='0.003arcsec', imsize=[3000], mask = LB1_mask, scales = LB_scales, threshold = '0.02mJy', robust=0.5, interactive=True)
 estimate_SNR(cont_final+'.image', disk_mask = LB1_mask, noise_mask = noise_annulus)
-#DoAr25_cont_final_r0.image
-#Beam 0.042 arcsec x 0.021 arcsec (74.31 deg)
-#Flux inside disk mask: 227.07 mJy
-#Peak intensity of source: 1.39 mJy/beam
-#rms: 1.73e-02 mJy/beam
-#Peak SNR: 80.27
+#DoAr25_cont_final_r0.5.image
+#Beam 0.048 arcsec x 0.027 arcsec (79.38 deg)
+#Flux inside disk mask: 227.45 mJy
+#Peak intensity of source: 1.84 mJy/beam
+#rms: 1.34e-02 mJy/beam
+#Peak SNR: 137.41
 
-
[…]

### 2018-05-23 reduction_scripts/reduction_utils.py
corrected redundancy in reduction script

@@ -278,20 +278,18 @@ def image_each_obs(ms_dict, prefix, scales, smallscalebias = 0.6, mask = '', thr
 def fit_gaussian(imagename, region, dooff = False):
     """
     Wrapper for imfit in CASA to fit a single Gaussian component to a selected region of the image
-
     Parameters
     ==========
     imagename: Name of CASA image (ending in .image) (string)
     region: CASA region format, e.g., 'circle[[200pix, 200pix], 3arcsec]' (string)
     dooff: boolean option to allow for fitting a zero-level offset 
-
     """
     imfitdict = imfit(imagename = imagename, region = region, dooff = dooff)
+    # Check if the source was resolved
+    was_resolved = not imfitdict['deconvolved']['component0']['ispoint']
+    # Get the coordinate system
     coordsystem = imfitdict['deconvolved']['component0']['shape']['direction']['refer']
-    PA = imfitdict['deconvolved']['component0']['shape']['positionangle']['value']
-    majoraxis = imfitdict['deconvolved']['component0']['shape']['majoraxis']['value']
-    minoraxis = imfitdict['deconvolved']['component0']['shape']['minoraxis']['value']
-    
+    # Get the parameters
     headerlist = imhead(imagename)
     phasecenter_ra, phasecenter_dec = headerlist['refval'][:2]
     peak_ra = imfitdict['deconvolved']['component0']['shape']['direction']['m0']['value']
@@ -300,7 +298,7 @@ def fit_gaussian(imagename, region, dooff = False):
     deltax, deltay = headerlist['incr'][:2]
     peak_x = xcen+np.unwrap(np.array([0, peak_ra-phasecenter_ra]))[1]/deltax*np.cos(phasecenter_dec)
     peak_y = ycen+(peak_dec-phasecenter_dec)/deltay
-
+    # Print
     if coordsystem=='J2000':
         print '#Peak of Gaussian component identified with imfit: J2000 %s' % au.rad2radec(imfitdict = imfitdict, hmsdms = True, delimiter = ' ')
     elif coordsystem=='ICRS':
@@ -309,9 +307,15 @@ def fit_gaussian(imagename, region, dooff = False):
         print '#Peak in J2000 coordinates: %s' % J2000coords
     else:
[…]

### 2018-06-25 projects/dullemond/paper_dusttraps/ms.tex
New version of paper on dust traps by Kees et al.

@@
 \title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
-\authorrunning{Dullemond, Birnstiel, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
-\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
+\authorrunning{Dullemond, Birnstiel, Troncoso, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
+\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, N.~Troncoso$^{3}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
 \institute{
@@
 The concept of dust trapping in local pressure maxima has become a central theme
-in studies of planet formation and protoplanetary disk evolution, as it may
+in studies of planet formation and protoplanetary disk evolution, because it might
 provide an elegant solution to several problems in these fields of
 study. Theories of planet formation are plagued by the ``radial drift barrier'':
-the problem that as dust aggregates grow by coagulation, they tend to radially
+the problem that, as dust aggregates grow by coagulation, they tend to radially
 drift toward the star before they reach planetsimal size
@@
 \citep{1972fpp..conf..211W, 2007ApJ...664L..55K, 1995A&A...295L...1B,
-  1997Icar..128..213K}. From an observational perspective, the radial drift
+  1997Icar..128..213K}. Not only does this process prevent excessive radial
+drift of dust particles, it also tends to concentrate the dust into small
+volumes and high dust-to-gas ratios, which is beneficial to planet formation.
+From an observational perspective, the radial drift
 problem manifests itself by the presence of large grains in the outer regions of
@@
 previously suspected, leading to a higher gas friction for millimeter grains and
-thus longer drift times \citep{2017ApJ...840...93P}. But another explanation is
-to invoke, again, pressure traps. The most striking observational evidence for
+thus longer drift time scales \citep{2017ApJ...840...93P}. Another explanation is
+to invoke dust traps. The most striking observational evidence for
 dust trapping seems to come from large transitional disks, which feature giant
@@
 traps would have to be more subtle. \cite{2012A&A...538A.114P} explored the
-possibility that the disk features many axisymmetric local pressure maxima, and
+possibility that the disk contains many axisymmetric local pressure maxima, and
 calculated how the dust drift and growth would behave under such conditions. It
@@
 that the dust does not escape from the ring. This makes it possible to look for
-a steady-state dust trapping model in which the radial drift forces (that push
+a steady-state dust trapping solution in which the radial drift forces (that push
 the dust to the pressure peak) are balanced by turbulent mixing (that tends to
@@
 modeling, we review, in Section \ref{sec-data}, the key aspects of the data from
-our ALMA Large Programme that we compare our models to.
-
-\section{The high-contrast rings of AS 209, Elias 24, GW Lup, HD 143006 and HD 163296}
+our ALMA Large Programme that we compare to our models.
+
+\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
 \label{sec-data}
 %
-In this paper we discuss in particular those sources of the ALMA Large Programme
+In this paper we focus on those sources of the ALMA Large Programme
 that show high-contrast and radially thin rings that are separated by deep
 valleys, and that are sufficiently face-on to not have to worry about 3-D
-lie-of-sight aspects. These are AS 209, Elias 24, GW Lup, HD 143006 and HD
-163296. The high-contrast rings of these sources provide ``clean laboratories''
+line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
+HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|ccc|}
+\hline
+\hline
+Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
+\hline
+AS 209     & 121              & 0.912        &  0.631 \\
+Elias 24   & 136              & 0.776        &  6.03  \\
+HD 163296  & 101              & 2.14         &  25.7  \\
+GW Lup     & 155              & 0.457        &  0.331 \\
+HD 143006  & 166              & 0.813        &  0.871 \\
+\hline
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{tab-stellar-params}The stellar parameters for the
+  stars studied in this paper. Distance is in parsec and mass and luminosity
+  are in units of the solar values.}
+\end{table}
+
+A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.
+For an overview of the ALMA Large Programme we refer to {\bf Andrews et
+  al.~XXXX}, and for an in-depth discussion on the data of the individual
+sources we refer to {\bf [LIST HERE THE PAPERS]}.
+
+\begin{figure*}
+\centerline{\includegraphics[width=0.95\textwidth]{gallery_rings.pdf}}
+\caption{\label{fig-obs-images}The continuum maps in
+  band 6 of the five disks in our sample which have the most pronounced rings.
+  The eight highest contrast rings, which are the topic of this paper,
+  are marked in the images. For a detailed description of these
+  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
+\end{figure*}
+
+The high-contrast rings of these sources provide ``clean laboratories''
 for testing the theory of dust trapping, both in a ring-by-ring manner, as well
-as by comparing to global dust drift models. The most striking case is AS 209,
-which features two high-contrast radially thin rings which are isolated and
-large ({\bf see Guzm\'an et al.~(2018) for the detailed description of this
-  dataset}). Fig.\ref{fig-obs-as209} shows the radial profile (deprojected for
-inclination) of the thermal emission of the dust of the disk around AS 209. The
+as by comparing to global dust drift models. Fig.\ \ref{fig-obs-profiles}
+shows the radial profile (deprojected for
+inclination) of the thermal emission of the dust of the five disks. The
+procedure used to extract these radial profiles from the continuum maps
+is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
+HD 143006 we removed the blob]}. The
 brightness is expressed as linear brightness temperature
-($T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$), where $I_\nu$ is the
+$T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
 intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
-frequency. It should be noted that the full (non-linear) brightness temperature
-(defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
-Planck function) can have very different values, in particular at low
+frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
+defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
+Planck function, can have very different values, in particular at low
 brightness. As a result $T_{\mathrm{bright,full}}(r)$ tends to vastly
@@
 regions. Therefore, in this paper we plot the brightness using the linear
-brightness temperature.
-
-\begin{figure}
-\centerline{\includegraphics[width=0.5\textwidth]{as209_gauss_modelfit.pdf}}
-\caption{\label{fig-obs-as209}The linear brightness temperature in band 6 of the
-  disk around AS 209, as a function of deprojected distance to the star (see
-  {\bf Guzm\'an et al.~(2018)} for a detailed description and analysis of these data). The
-  vertical axis is logarithmic to better show the contrast. The two most
-  prominent rings are fitted by Gaussian rings. Ring 1 is fitted by
-  $T_{\mathrm{bright,linear}}(r)=
-  3.5\,\exp(-0.5(r-77\,\mathrm{au})^2/(4.0\,\mathrm{au})^2)$, while ring 2 is fitted by
-  $T_{\mathrm{bright,linear}}(r)=
-  2.8\,\exp(-0.5(r-124.7\,\mathrm{au})^2/(4.4\,\mathrm{au})^2)$.}
-\end{figure}
-
-The two most prominent rings (seen as bumps in Fig.\ref{fig-obs-as209}) are
-fitted by Gaussian profiles with a standard deviation width of
-$4.0\,\mathrm{au}$ and $4.4\,\mathrm{au}$ respectively. This corresponds to a
-full-width-at-half-maximum (FWHM) of $9.3\,\mathrm{au}$ and $10.3\,\mathrm{au}$.
-The ALMA beam FWHM for these observations is roughly $4.4\,\mathrm{au}$, meaning
-that the rings are 2.1 and 2.3 beams wide.
-
-These Gaussian profiles fit the observed data remarkably well, even in the wings
-of the profile (except the outer wing of the outer ring). This is somewhat
-surprising, given that the rings are resolved. The Gauss profiles therefore do
-not represent the beam profile, but tell something about the actual dust
-distribution within the ring.
-
-The source Elias 24 also shows a prominent high-contrast isolated ring,
-albeit flanked on the outside by a plateau of emission (see {\bf Andrews
-  et al.~(2018) for a detailed description of this dataset}).
-
-\begin{figure}
-\centerline{\includegraphics[width=0.5\textwidth]{elias24_gauss_modelfit.pdf}}
-\caption{\label{fig-obs-elias24}As Fig.~\ref{fig-obs-as209}, but now for the
-  source Elias 24 (see {\bf Andrews et al.~(2018)} for a detailed description
-  and analysis of these data). The most prominent ring is fitted by a Gaussian
-  profile $T_{\mathrm{bright,linear}}(r)=
-  5.3\,\exp(-0.5(r-79.9\,\mathrm{au})^2/(5.0\,\mathrm{au})^2)$.}
-\end{figure}
-
-The ring can be reasonably well fitted with a Gaussian profile with standard
-deviation width of $5.0\,\mathrm{au}$ corresponding to a FWHM of
-$11.7\,\mathrm{au}$. Given a beam size in band 6 of about $4.9\,\mathrm{au}$,
-the ring of Elias 24 is also resolved by about 2.4 beam widths. The Gaussian fit
-is, however, less convincing in the case of Elias 24 than in the case of AS 209,
-so that it cannot be clearly argued that the ring has a Gaussian shape in the
-case of Elias 24.
-
-{\bf [TODO: GW Lup, HD 143006 and HD 163296]}
+brightness temperature, which is proportional to the intensity in Jy/beam.}.
+
+\begin{figure*}
+\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
+\caption{\label{fig-obs-profiles}The linear brightness temperature in
+  band 6 of the five disks in our sample which have the most pronounced rings.
+  The vertical axis is logarithmic to better show the contrast.
+  The eight highest contrast rings are fitted by a Gaussian profile, shown
+  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
+  with the width of the ALMA beam. For a detailed description of these
+  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
+\end{figure*}
+
+The eight rings which we study in this paper (see annotations in
+Fig.~\ref{fig-obs-images}) are fitted with a Gaussian profile, shown as inverse
+parabolas in Fig.\ \ref{fig-obs-profiles}. The fitting procedure was done in the
+image plane, in order to make it possible to fit to a limited radial range,
+i.e.\ selecting a single ring. Details of this fitting procedure, and tests
+using the ALMA simulator, are described in Appendix {\bf XXXXX}. In
+Fig.~\ref{fig-obs-gaussfits} these gaussian fits are shown in close-up. The
+parameters of the best-fitting Gaussians are listed in Table
+\ref{tab-gauss-params}.
+
+\begin{figure*}
+\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
+\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
+  paper. The blue curves are the observations, the green curves are the best fit
+  Gaussian profiles. The ``fit range'' bar shows the radial range within which
+  the Gauss was fitted to the data. The fit range was chosen to fit the part of
+  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
+  FWHM beam size of the observations. The grey band around the blue curve shows
+  the estimated uncertainty of the data.}
+\end{figure*}
+
+\begin{figure*}
+  \centerline{
+    %\includegraphics[width=0.3\textwidth]{AS_209_doublegauss.pdf}
+    \includegraphics[width=0.42\textwidth]{HD_163296_doublegauss.pdf}
+    \includegraphics[width=0.42\textwidth]{HD_143006_doublegauss.pdf}
+  }
+  \caption{\label{fig-obs-doublegauss}The sum of the two Gaussian fits
+    for the two sources with two partly overlapping rings: HD 163296 and
+    HD 143006.}
+\end{figure*}
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{|cc|cc|ccccc|ccccc|}
+\hline
+\hline
+%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{model}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
+Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{model}}$ & $T_{\mathrm{model,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
+            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+\hline
+% NOTE: This table is made automatically with python_make_tables/make_latex_table.py
+AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  12.9 &   8.2 &   0.7 &  44.4 &   2.41\\
+AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  10.1 &   5.6 &   0.5 & 101.9 &   5.54\\
+Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &  41.1 &   2.23\\
+HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  34.2 &  29.0 &   1.6 &  50.4 &   2.74\\
+HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  28.1 &  22.9 &   0.8 &  42.5 &   2.31\\
+GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &  36.3 &   1.97\\
+HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  71.76 &   5.06 &   4.42 &  18.8 &  13.8 &   1.7 &  16.7 &   0.91\\
+HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 111.83 &   7.88 &   7.49 &  15.0 &  10.1 &   1.7 &  43.5 &   2.36\\
+\hline
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
+  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values $r_{\mathrm{in}}$
+  and $r_{\mathrm{out}}$ are the inner and outer boundary of the fitting domain. $A$ is the
+  peak brightness temperature of the ring, $r_0$ is its radius in $\mathrm{au}$,
+  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
+  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
+  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
+  dust emission profile, also expressed as standard deviation in units of
+  $\mathrm{au}$. The last five columns are derived from a simple disk model
+  assumption: $T_{\mathrm{model}}$ is the midplane temperature computed from
+  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
+  $\varphi=0.02$. From that one can derive the corresponding linear brightness
+  temerature $T_{\mathrm{model,br,lin}}=B_\nu(T_{\mathrm{model}})c^2/2\nu^2$ with
+  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of
+  0.13 $\mathrm{cm}$ of ALMA band 6.
+  The third-to-last column expresses the deconvolved dust ring
+  width in units of the disk pressure scale height $h_p$ computed from
+  $T_{\mathrm{model}}$. The last two columns are the computed dust mass in the
+  ring in units of Earth masses, assuming the small grain absorption opacity
+  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
+  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
+  value
+  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
+  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). The stellar parameter
+  are taken from Table \ref{tab-stellar-params}.}
+\end{table*}
+
+The observed rings are the result of the thermal emission of a dust ring convolved
+with the ALMA beam. The standard deviation width $\sigma$ of the fitted Gaussian
+profile is given in Table \ref{tab-gauss-params}.  To obtain the width of the
+underlying dust ring we have to deconvolve. Assuming a Gaussian beam and a
+Gaussian dust ring, we can use the rule of the convolution of two Gaussians, and
+obtain the width $w_d$ of the dust ring
+\begin{equation}
+w_d=\sqrt{\sigma^2-b^2}
+\end{equation}
+where $b$ is the beam width expressed as standard deviation in units of
+$\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
+$b_{\mathrm{fwhm,as}}$ is the beam full width at half maximum in units of
+arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
+parsec). The width of the dust rings
+in units of the local pressure scale height $h_p$ (last column of Table
+\ref{tab-gauss-params}) can only be roughly estimated, because we do not know
+the disk midplane temperature. Our estimate is made as follows:
+\begin{equation}\label{eq-disk-temperature-model}
+T_{\mathrm{model}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
+\end{equation}
+where $\sigma_{\mathrm{SB}}$ is the Stefan-Boltzmann constant and $\varphi$ is
+the so-called flaring angle \citep[e.g.][]{1997ApJ...490..368C,
+  1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
+follows from $h_p=\sqrt{k_BT_{\mathrm{model}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
+Boltzmann constant, $m_p$ the proton mass, $G$ the gravitational constant and
+$mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
+taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
+$\varphi=0.02$ which is an estimate based on typical values from models.  But
+this value can easily be wrong by a factor of four or so. The fact that the
+pressure scale height goes as $T_{\mathrm{model}}^{1/2}\propto \varphi^{1/8}$
+means, however, that such an error only affects $h_p$ mildly.
+
+We see from Table \ref{tab-gauss-params} that some rings are narrower than the
+(estimated) pressure scale height $h_p$, while others are broader. This
+comparison is important because long-lived pressure bumps in the gas cannot be
+radially narrower than about one pressure scale height. If a thermal emission
+ring produced by the dust is substantially narrower than $h_p$, then some kind
+of dust trapping must have taken place. If this were not the case, then the dust
+emission profile would have the same shape as the gas midplane density, and thus
+the gas midplane pressure bump would then be a narrower than $h_p$.
+
+We see that 5 out of 8 rings are narrower than $h_p$, but not by much. None are
+less than half the width of the pressure scale height of the gas. This is
+perhaps slightly disappointing, because it means that these results are no
+irrefutable proof that dust trapping is playing a role in the shaping of these
+rings. But it is strong evidence for that nonetheless.
+
+For these observations the fwhm beam size is of the order of 35 milliarcseconds
+{\bf [CHECK AND MAKE MORE ACCURATE]}. This number can be compared to the fwhm
+ring width in Table \ref{tab-gauss-params}, and it shows that the widths of all
+the rings are spatially resolved. Some are only marginally resolved (AS 209
+rings 1 and 2, GW Lup and the ring 1 of HD 143006), but the data are
+inconsistent with radially unresolved rings. This is confirmed by more detailed
+analysis in the uv-plane ({\bf REFS TO Guzm\'an et al.~(2018)}).
+
+These Gaussian profiles fit the observed data remarkably well, as can be
+best seen in Fig.~\ref{fig-obs-gaussfits}. 
+For the broad rings of HD 163296 the closeness to a Gaussian profile is
+particularly striking, given that these rings are resolved by many beam
+widths. In particular in this source there is therefore no doubt that the
+Gaussian profile is a reflection of a real phenomenon, and is not caused by
+unknown observational or data reduction issues.  But also the rings of AS 209,
+while not as well-resolved as those of HD 163296, are remarkably close to
+Gaussian.
+
+For Elias 24 the ring fits a Gaussian near the peak, but deviates from Gaussian
+already less half a width from the center. Likewise for the rings of HD
+143006. The ring around GW Lup is an intermediate case, which fits a Gauss well
+on the inner flank, but rather badly on the outer flank.
+
+In all ring profiles shown in Fig.~\ref{fig-obs-gaussfits} the observed profile
+rises above the Gaussian fit at some point in the wings. For the inner flanks of
+ring 1 of Elias 24 and ring 1 of HD 143006, as well as for the outer flanks of
+ring 2 of AS 209, ring 2 of HD 163296, the ring of GW Lup and ring 2 of HD
+143006 the excess above the Gaussian gradually increases away from the peak of
+the Gaussian. Put differently, the profiles appear to deviate from Gaussian
+shape to pyramid-shape (or Lorenzian shape) in the flanks. This cannot be simply
+the addition of a constant background, because that would preserve the Gaussian
+shape, and merely lift it higher. Instead, while the core of the profile can clearly
+be well-described by a Gaussian, in the flanks there is a clear transition
+to a different, wider, shape.
+
+A similar effect occurs between the two rings in HD 163296 and HD 143006, where
+in particular for HD 143006 it is clear from Fig.~\ref{fig-obs-doublegauss} that
+it cannot be just the effect of the sum of the two Gaussians.
+
+It is not immediately clear whether this is relevant or just chance. It lies in
+the nature of fitting a Gaussian profile to something non-Gaussian that there
+will be a region close to the peak where the curve fits the Gaussian reasonably
+well, while the deviation will increase the farther away from the peak one
+looks. This danger is particularly large in the present case, since the fitting
+range was chosen to maximize the similarity to the Gauss near the
+peak. Nevertheless, it clearly signals a deviation from Gaussian.
+
+Given the closeness of the profiles to a Gaussian, we can use the Gaussian fit
+to make estimates of the total mass of dust trapped in each of these rings. The
+procedure is described in Appendix \ref{sec-computing-dust-mass}, and the
+results, for the two opacity choices we make, are listed in Table
+\ref{tab-gauss-params}. In making these estimates we have assumed the dust to be
+made up of micron-sized dust, which is likely too small. The mass based on the
+Ivezic opacity, $M_{\mathrm{d,iv}}$, will increase for $a>\lambda/2\pi$ by a
+factor of $2\pi\,a/\lambda$ with $\lambda=0.13\,\mathrm{cm}$. For the mass based
+on the Ricci opacity the factor first decreases when $a$ approaches the opacity
+peak around $a=\lambda/2\pi$ and then increases for $a>\lambda/2\pi$. Both
+behaviors can be seen as the inverse behavior of the opacities shown in
+Fig.~\ref{fig-opacities}. For ever increasing grain size estimate, the mass
+estimate will increase indefinitely for both opacity choices.
+
+\section{The rings as dust traps}
+\label{sec-rings-as-dust-traps}
+%
+Given the well separated narrow rings in the above sources, we can analyze these
+rings on a one-by-one basis. The hypothesis we are going to test is that the
+rings are caused by dust trapping in axisymmetric pressure bumps. For simplicity
+we will assume that the radial gas pressure profile is fixed in time, and there
+is no back reaction of the dust onto the gas. The pressure bump is assumed to be
+so strong that the dust trapping in these rings is nearly perfect: no dust
+escapes. We then expect that the dust distribution finds an equilibrium between
+dust drift and turbulent spreading. Low turbulence will lead to thinner dust
+rings than high turbulence. For simplicity we shall assume the gas pressure bump
+to have a Gaussian radial profile with its peak at radius $r_0$ and width $w\ll
+r_0$. This problem can be solved analytically.
+
+\subsection{Model}
+\label{sec-model-dusttrap-rings}
+%
+Consider the following radial Gaussian profile for the
+pressure at the disk midplane:
+\begin{equation}\label{eq-gaussian-pressure-bump}
+p(r) = p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right)
+\end{equation}
+We will discuss the hydrodynamic stability of a gas
+distribution arranged in this way in Section \ref{sec-gas-ring-stability}. But
+for now we will not be concerned with that and we will assume the
+pressure-profile of Eq.~(\ref{eq-gaussian-pressure-bump}) to be given and fixed.
+
+In Appendix \ref{sec-steady-state-analytic-trap-model} we derive that, for our
+purposes, a good approximation of the stationary dust distribution within the
+dust trap is:
+\begin{equation}\label{eq-analytic-sol-radial-trapping-summary}
+\Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}^2}\right)
+\end{equation}
+where
+\begin{equation}\label{eq-wd-afo-w-psi}
+  w_{\mathrm{d}} = w\,\left(1+\psi^{-2}\right)^{-1/2}
+\end{equation}
+with $\psi$ given by
+\begin{equation}\label{eq-psi-afo-alpha-sc-st}
+\psi = \sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
+\end{equation}
+Here $\mathrm{St}$ is the Stokes number of the dust particles
+(Eq.~\ref{eq-definition-stokes-number}), $\mathrm{Sc}$ is the Schmidt number of
+the turbulence in the gas, and $\alpha_{\mathrm{turb}}$ is the usual turbulence
+parameter. Note that this solution is for a single grain size.
+
+For large grains and/or weak turbulence one finds $\psi\ll 1$, which leads to
+$w_{\mathrm{d}}\ll w$. In this case the dust is strongly trapped near the peak
+of the pressure bump. The opposite is the case for small grains and/or strong
+turbulence, for which one gets $\psi\gg 1$, which leads to
+$w_{\mathrm{d}}\rightarrow w$. In this case the trapping is very weak and the
+dust-to-gas ratio within the pressure bump stays constant. Only in the very
+wings of the Gaussian pressure bump is dust trapping significant, but this
+effect can only be studied using the more accurate non-gaussian solution of
+Eq.~(\ref{eq-analytic-sol-radial-trapping-better}). We will return to the dust
+trapping effect in the wings in Section {\bf XXXXX}.
+
+Eq.~(\ref{eq-analytic-sol-radial-trapping-summary}) has only two parameters:
+$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. Both can be fairly directly
+measured from the observations, though $\Sigma_{\mathrm{d0}}$ will depend on the
+opacity, and hence on the unknown grain size (see discussion in Appendix
+\ref{sec-computing-dust-mass}). The values of $w_{\mathrm{d}}$ for
+the rings in our sample can be directly taken from Table \ref{tab-gauss-params}.
+
+The width of the dust ring $w_{\mathrm{d}}$ is physically set by
+$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$ through the above
+equations. We therefore
+have one observational value for four unknown parameters. This is heavily
+degenerate. All we can do is to test if the measured value of
+$w_{\mathrm{d}}$ is consistent with expected values of $\alpha_{\mathrm{turb}}$,
+$\mathrm{Sc}$, $\mathrm{St}$ and $w$.
+
+\subsection{Ranges of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$}
+\label{sec-ranges-of-params}
+%
+First of all, the Schmidt number $\mathrm{Sc}$ is merely a way to relate the
+turbulent viscosity with the turbulent mixing. If we do not strive to learn
+about the turbulent viscosity and instead are satisfied with learning only about
+the turbulent mixing, then we are only interested in the combination
+$\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For simplicity we set $\mathrm{Sc}=1$.
+
+The value of the turbulence parameter $\alpha_{\mathrm{turb}}$ is usually
+considered to be between $10^{-6}\lesssim \alpha_{\mathrm{turb}}\lesssim
+10^{-2}$.
+
+The width of the pressure bump cannot be smaller than about a pressure
+scale height, but also not smaller than the width of the dust ring.
+Therefore $w_{\mathrm{min}}=\mathrm{max}(h_p,w_{\mathrm{d}})$.
+In the case of the double rings (AS 209, HD 163296 and HD 143006),
+the full-width-at-half-maximum $2.355\,w$ should not exceed the radial
+separation of the rings. For the two single ring sources we take the deepest
+point of the gap to the inside of the ring as the upper limit on the
+half-width-at-half-maximum $1.178\,w$. These lower and upper limits on $w$
+are listed in Table \ref{tab-ring-model-limits}.
+
+The Stokes number $\mathrm{St}$ can be any value. But it is
+directly related to the grain size $a$ and the gas density
+$\rho_{\mathrm{g}}$. If we have observational constraints on $a$ and a good
+estimate of the gas density $\rho_{\mathrm{g}}$, then we can eliminate this
+uncertainty, and we are left with two unknown parameters
+($\alpha_{\mathrm{turb}}$ and $w$) for one measurement ($w_{\mathrm{d}}$).
+Unfortunately, while estimating $a$ from observations may be doable, it is far
+more difficult to estimate $\rho_{\mathrm{g}}$. Standard disk gas mass estimates
+are of limited use, as they are based on measuring the dust mass and multiplying
+it by the estimated gas-to-dust ratio. Since we are testing the hypothesis of
+dust trapping, we cannot assume a standard gas-to-dust ratio.
+
+One can, however, set an upper bound on $\rho_{\mathrm{g}}$ by demanding that
+the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
+\begin{equation}
+Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
+\end{equation}
+Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
+Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
+weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
+$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
+surface density. The gas surface density and midplane density are related via
+the pressure scale height $h_p=c_s/\Omega_K$ via the identity
+$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. Taking the disk midplane
+temperature from Table \ref{tab-gauss-params}, which was calculated using
+Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
+$\rho_{\mathrm{g}}$ for all of the rings. The results are given in Table
+\ref{tab-ring-model-limits}.
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{|cc|cccccc|cc|}
+\hline
+\hline
+Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\rho_{\mathrm{g,min,iv}}$ & $\rho_{\mathrm{g,min,ri}}$ & $\rho_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $(\alpha/\mathrm{St})_{\mathrm{min}}$ & $(\alpha/\mathrm{St})_{\mathrm{max}}$\\
+           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & $\equiv\psi^{-2}_{\mathrm{max}}$ & $\equiv\psi^{-2}_{\mathrm{min}}$\\
+\hline
+AS 209     & 1 &  4.83 & 19.6 & 7.9e-17 & 1.4e-15 & 8.4e-14 & 11.1 & 3.2e-02 & 1.0e+00 \\
+AS 209     & 2 &  8.85 & 19.6 & 5.1e-17 & 9.3e-16 & 2.0e-14 &  4.0 & 5.0e-02 & 3.0e-01 \\
+Elias 24   & 1 &  7.26 & 17.1 & 3.6e-17 & 6.5e-16 & 6.4e-14 & 18.8 & 8.0e-02 & 7.0e-01 \\
+HD 163296  & 1 &  7.21$^{*}$ & 13.8 & 5.6e-17 & 1.0e-15 & 2.6e-13 & 48.7 & 3.8e-01 & -- \\
+HD 163296  & 2 &  7.27 & 13.8 & 2.5e-17 & 4.6e-16 & 8.1e-14 & 33.4 & 1.9e-01 & 1.4e+00 \\
+GW Lup     & 1 &  7.49 &  9.9 & 2.9e-17 & 5.4e-16 & 2.8e-14 &  9.8 & 2.1e-01 & 4.4e-01 \\
+HD 143006  & 1 &  4.42$^{*}$ & 10.1 & 7.8e-17 & 1.4e-15 & 4.4e-13 & 58.9 & 2.4e-01 & -- \\
+HD 143006  & 2 &  7.49$^{*}$ & 10.1 & 4.6e-17 & 8.5e-16 & 1.1e-13 & 25.1 & 1.2e+00 & -- \\
+\hline
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{tab-ring-model-limits}Limits on the free parameters of the dust
+  trapping model. The lower limit to the pressure bump width $w_{\mathrm{min}}$
+  is the pressure scale height $h_p$. If, however, the width of the dust ring
+  $w_d>h_p$, then the lower limit is $w_d$ (marked with the symbol $^{*}$).  The
+  upper limit $w_{\mathrm{max}}$ is derived from the separation between the
+  rings (for AS 209, HD 163296 and HD 143006) or from the separation of the ring
+  to the nearest minimum (for Elias 24 and GW Lup). The two lower limits on the
+  gas density $\rho_{\mathrm{g}}$ were derived by demanding
+  $\Sigma_{\mathrm{g}}\gtrsim \Sigma_{\mathrm{d}}$, where $\Sigma_{\mathrm{d}}$
+  was computed in Appendix \ref{sec-computing-dust-mass} using two different
+  opacities: that of Ivezic et al. (leading to $\rho_{\mathrm{g,min,iv}}$) and
+  that of Ricci et al.\ (leading to $\rho_{\mathrm{g,min,ri}}$). The column
+  marked with $\rho_{\mathrm{g,max}}$ is the upper limit on the midplane gas
+  density derived from demanding that the gas disk is gravitationally
+  stable. The $a_{\mathrm{max}}$ column is the maximum grain size for which the
+  derived dust surface density together with the gas surface density remain
+  gravitationally stable. The column $(\alpha/\mathrm{St})_{\mathrm{min}}\equiv
+  \psi^{-2}_{\mathrm{max}}$ is a result of the analysis: the lower limit to
+  $\alpha_{\mathrm{turb}}/\mathrm{St}$ derived from the ratio of
+  $w_{\mathrm{max}}/w_{\mathrm{d}}$.}
+\end{table*}
+
+One can estimate a lower limit to the gas density by demanding that the gas
+surface density must be at least as large as the dust surface density, since
+dust trapping is unlikely to achieve a larger concentration of dust than
+that. Depending on which of the two opacities we use (see Appendix
+\ref{sec-computing-dust-mass}) we arrive at different estimates of
+$\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
+\Sigma_{\mathrm{d}}$, and using the pressure scale height $h_p$
+and the formula $\rho_{\mathrm{g,midplane}}=\Sigma_{\mathrm{g}}/\sqrt{2\pi}h_p$,
+we arrive at two lower limits of the midplane gas density, given in
+Table \ref{tab-ring-model-limits}. It is likely that even for much
+larger values of the midplane gas density the dust-gas mixture becomes
+unstable to the streaming instability and other types of instabilities.
+If we can convince ourselves that such instabilities are not operating
+in these rings, then this puts a substantially stronger (i.e.~larger)
+lower limit on the midplane gas density.
+
+Also if the grains are much larger than $\lambda/(2\pi)\simeq 0.02\,\mathrm{cm}$,
+the opacity drops and the resulting dust surface density estimate increases,
+yielding larger lower limits to the gas density.
+
+Along this line of thinking one can compute the largest grain radius for which
+the Toomre parameter of the disk stays above 2. We use the Ivezic
+opacity for that, which, for large grains, has similar values as the Ricci
+opacity. This leads to values of several centimeters to half a meter (Table
+\ref{tab-ring-model-limits}). Again these are conservative limits, with real
+values likely to be substantially smaller.
+
+\subsection{Application to the observed rings}
+\label{sec-application-to-rings}
+%
+We now wish to apply the model of Subsection \ref{sec-model-dusttrap-rings}
+with the limits on the parameter ranges derived in Subsection
+\ref{sec-ranges-of-params} to the observed ring widths $w_d$ listed in
+Table \ref{tab-gauss-params}.
+
+Let us first choose the widest allowed pressure bump (i.e.\ taking
+$w=w_{\mathrm{max}}$ from Table \ref{tab-ring-model-limits}). In that
+case the dust rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$,
+which implies that the dust trapping is operational. We can derive the
+values of $\psi$ directly from Eq.~(\ref{eq-wd-afo-w-psi}). And together
+with Eq.~(\ref{eq-psi-afo-alpha-sc-st}) we can then directly derive the
+following constraint on the unknown parameters $\alpha_{\mathrm{turb}}$
+and $\mathrm{St}$, 
+\begin{equation}
+\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}} \equiv \psi^2 = \left(\frac{w}{w_{\mathrm{d}}}\right)^2-1
+\end{equation}
+where, we recall, set $\mathrm{Sc}=1$. Given that this is the largest
+possible value of $w$, and therefore the largest possible value
+of $\psi$, this gives a lower limit to the ratio $\alpha_{\mathrm{turb}}/\mathrm{St}$:
+\begin{equation}
+\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \gtrsim \frac{1}{\psi^2_{\mathrm{max}}} \qquad\hbox{($\psi_{\mathrm{max}}$ from largest $w$)}
+\end{equation}
+The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
+the first quantitative results of the application of the dust trapping model to
+the data.
+
+We can also derive an upper limit to the ratio
+$\alpha_{\mathrm{turb}}/\mathrm{St}$, but only for the rings for which
+$w_{\mathrm{d}}<h_p$. The reason is that for those rings with
+$w_{\mathrm{d}}>h_p$ (marked with a $^{*}$ in Table \ref{tab-ring-model-limits})
+the minimal pressure bump width is $w_{\mathrm{min}}=w_{\mathrm{d}}$, and the
+dust ring is as wide as the pressure bump, implying that dust trapping is weak.
+Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will keep
+$w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive an upper limit on
+$\alpha_{\mathrm{turb}}/\mathrm{St}$. For the rings with $w_{\mathrm{d}}<h_p$
+(and thus $w_{\mathrm{min}}=h_p$) we can write
+\begin{equation}
+\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \lesssim \frac{1}{\psi^2_{\mathrm{min}}} \qquad\hbox{($\psi_{\mathrm{min}}$ from smallest $w$)}
+\end{equation}
+The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
+the second quantitative results of the application of the dust trapping model to
+the data.
+
+If we will be able to measure the width of the gas pressure bump in an
+independent way, for instance with molecular line observations, then instead of
+the lower- and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$ we will be
+able to directly measure $\alpha_{\mathrm{turb}}/\mathrm{St}$. However, reliable
+measurements of the gas density are notoriously difficult. CO molecular lines
+are usually emitted in the warm sub-surface layers of the disk, not near the
+midplane (see e.g.~{\bf Weaver \& Isella companion paper as well as earlier
+  literature}), and CO appears to be depleted compared to chemistry models
+({\bf XXXX}). Other gas tracers are harder to measure and also not reliable
+indicators of the gas density. Therefore we will, at least for now, have to
+make do with the lower and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$
+from Table \ref{tab-ring-model-limits}.
+
+The next task is to convert from Stokes number $\mathrm{St}$ to grain size
+$a$.
+
+{\bf [TO BE DONE]}
+
+{\bf Make an exclusion diagram.}
+
+{\bf We can make predictions for pressure slopes in CO}
+
+{\bf Coincidence of the slope changes.}
+
+{\bf Can the height measurement help? There is still the degeneracy. Why did
+Pinte et al. not have this degeneracy? How could they pin down the turbulence?}
+
+\subsection{Including a grain size distribution}
+%
+
+{\bf [Here we will move away from a single grain size and see if we can understand
+the deviations from Gaussian by a grain size distribution]}
+
+\section{Streaming instability, clumping, and the spectral index}
+\label{sec-si-clump}
+%
+
+{\bf [This section is temporarily hidden, because a lot still has to be
+done on this.]}
+
+% Whenever the dust trapping becomes so effective that the volume density of the
+% dust at the midplane of the disk exceeds the gas volume density, the streaming
+% instability may set in \citep{2005ApJ...620..459Y}. This produces turbulence and
+% leads to clumping \citep{2007ApJ...662..627J, 2008A&A...479..883L,
+%   2010ApJ...722.1437B}. If conditions are right, some of these clumps may
+% gravitationally collapse and form planetesimals \citep{2007Natur.448.1022J,
+%   2017ApJ...847L..12S, 2017A&A...597A..69S}. It is believed that this mechanism
+% is the trigger that starts the process of the formation of planets
+% \citep[see][for a review]{2014prpl.conf..547J}. If one could find observational
+% evidence of this dust clumping occurring in protoplanetary disks, this would
+% be a major step forward for our understanding of planet formation.
+% 
+% The clumps that are formed in simulations of the streaming instability are,
+% however, typically substantially smaller than a pressure scale height of the
+% disk \citep{2007ApJ...662..627J, 2013MNRAS.434.1460K}. This makes it very hard
+% to spatially resolve these structures, even with ALMA at its highest resolution.
+% 
+% It might, however, be possible to find indications of particle clumping in an
+% indirect way, if it leads to the formation of optically thick clouds of
+% particles surrounded by optically thin regions. Even if we cannot spatially
+% resolve these clouds, the spatially averaged thermal emission from such a clumpy
+% medium has a different spectral slope than that of a homogeneous distribution of
+% dust. If all the dust is concentrated in a multitude of small and very optically
+% thick clumps, then the brightness of this unresolved clumpy medium would be the
+% Planck function times the covering fraction $C$ of the clumps:
+% \begin{equation}\label{eq-bright-clumpy}
+% I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,C
+% \end{equation}
+% where the limit $C\rightarrow 0$ means concentrating all the dust in infinitely
+% dense clumps so that the covering fraction is zero, and $C\rightarrow 1$ means
+% having so many clumps that they multiply overlap, when seen in projection from
+% above. In contrast, if all this dust is homogeneously distributed, then the
+% brightness is
+% \begin{equation}\label{eq-bright-homogeneous}
+% I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,(1-e^{-\tau_\nu})
+% \end{equation}
+% with $\tau_\nu$ being the vertical optical depth of the dust in the disk.
+% Reality will likely lie in between these two extreme scenarios, with
+% optically thin dust interdispersed between optically thick clouds.
+% 
+% There are at least two ways by which one could use Eqs.~(\ref{eq-bright-clumpy},
+% \ref{eq-bright-homogeneous}) to search for hints of clumpiness. The most direct
+% way is to make use of the different dependencies of these two equations on
+% wavelength $\lambda=c/\nu$. In the case of the clumpy medium with perfectly
+% optically thick clumps (Eq.~\ref{eq-bright-clumpy}) the brightness follows the
+% Planck curve, whereas for the homogeneous medium
+% (Eq.~\ref{eq-bright-homogeneous}), if it is optically thin
+% (i.e.~$1-e^{-\tau_\nu}\simeq \tau_\nu$), it follows the Planck curve times the
+% opacity law. If we are in the Rayleigh-Jeans domain of the Planck curve, then
+% these two limits are usually referred to as spectral slope
+% $\alpha_{\mathrm{spec}}=2$ and $\alpha_{\mathrm{spec}}=2+\beta$, respectively,
+% where $\beta$ is defined by $\kappa_\nu\propto \nu^\beta$. At ALMA band 6, for
+% the low temperatures we expect in these disks at large radii, we are no longer
+% strictly in the Rayleigh-Jeans domain, so the spectral slope analysis becomes
+% temperature-dependent, which complicates matters a bit, but the principle
+% remains the same. If the streaming instability is operating in the rings of our
+% sources, and if the numbers work out (see below), then we expect to see
+% $\alpha_{\mathrm{spec}}\simeq 2$ within the rings, and
+% $\alpha_{\mathrm{spec}}>2$ between the rings.
+% 
+% There is an important caveat to this idea: the grain size and the corresponding
+% opacity slope may mimick the same spectral signatures as the clumping. This is,
+% in fact, the same degeneracy of the interpretation of the millimeter spectral
+% slope as usual \citep[e.g.][]{2003A&A...403..323T}, just on a much smaller
+% spatial scale: a scale that remains unresolved even by ALMA.  Nevertheless it is
+% worthwhile to investigate if we can find clues that point to clumpiness.
+% 
+% In the scope of this paper we will only do a very simple analysis: we will
+% verify if, under the most benign conditions, the above mentioned optically
+% thick/thin clumpy medium is at all possible. This is not granted, because
+% these rings are several tens of au large, and therefore it requires a lot of
+% mass to make dust clumps optically thick at 1.3 millimeter wavelength.
+% Furthermore, these dust clouds cannot have a density in excess of the
+% Roche density, because then they would immediately gravitationally collapse.
+% Therefore, we cannot make the clumps arbitrary dense and compact.
+% 
+% So let us construct the most optimistic model of a dust cloud. The highest
+% possible density of a cloud is the Roche density, defined by
+% \begin{equation}
+% \rho_{\mathrm{Roche}}(r) = \frac{9}{4\pi}\frac{M_{*}}{r^3}
+% \end{equation}
+% where $M_{*}$ is the stellar mass and $r$ is the distance to the star. We take
+% $\rho_{\mathrm{cloud}}=\rho_{\mathrm{Roche}}$. The simplest opacity model for a
+% dust grain is that of \cite{1997MNRAS.291..121I}, where we omit the scattering
+% part:
+% \begin{equation}
+% \kappa(a) = \frac{3}{4}\frac{1}{\xi\,a}\left\{\begin{matrix}
+% 1 & \hbox{for} & \lambda<2\pi a \\
+% 2\pi a/\lambda & \hbox{for} & \lambda>2\pi a \\
+% \end{matrix}\right.
+% \end{equation}
+% where $a$ is the grain radius and $\xi$ is the material density of the dust
+% grain ($\xi=3.6$ for amorphous olivine, $\xi=2.0$ for silica {\bf CHECK}). For a
+% given wavelength $\lambda$ (in our case $\lambda=0.13\,\mathrm{cm}$ for ALMA
+% band 6), the most optimistic opacity is therefore $\kappa=(3\pi/2)/\xi\lambda$,
+% for all $a\le \lambda/2\pi$. For ALMA band 6 and $\xi=2\,\mathrm{g/cm}^3$ this
+% is $\kappa=18.1\,\mathrm{cm}^2/\mathrm{g}$. The largest grain size to still
+% have this opacity is therefore $a=\lambda/2\pi$, which for ALMA band 6 is
+% $a=0.02\,\mathrm{cm}$.  We make a spherical cloud of this dust with radius
+% $R$. The optical depth through the cloud center is then
+% \begin{equation}
+% \tau = 2R\rho_{\mathrm{cloud}}\kappa
+% \end{equation}
+% To obtain the effect of an optically thick/thin clumpy medium, these clouds
+% have to be sufficiently optically thick, not just marginally. So let us demand
+% a minimal optical depth of $\tau=4$. This gives a minimal cloud radius of
+% \begin{equation}
+%   R\gtrsim \frac{2}{\rho_{\mathrm{Roche}}\kappa} = 0.11\frac{1}{\rho_{\mathrm{Roche}}}
+%   = 0.15\frac{r^3}{M_{*}}
+% \end{equation}
+% in CGS units (for ALMA band 6). Note that this scales linearly with
+% $\rho_{\mathrm{cloud}}/\rho_{\mathrm{Roche}}$, so that if we choose a much less
+% ``critical'' cloud density of, say, $\rho_{\mathrm{cloud}}=0.1\,\rho_{\mathrm{Roche}}$,
+% then the cloud should become 10$\times$ larger and 100$\times$ more massive.
+% 
+% If we want these clouds to be part of an unresolved clumpy medium, they should
+% be substantially smaller than the pressure scale height of the disk, and
+% substantially smaller than the ALMA band 6 resolution. Let us take the example
+% of the source AS 209, with $M_{*}=0.9\,M_{\odot}$, and its two high-contrast
+% rings at 74 au and 120 au, respectively. This gives minimal cloud
+% radii of $R\gtrsim 4\times 10^{-3}\,\mathrm{au}$ and $R\gtrsim 22\times
+% 10^{-3}\,\mathrm{au}$, respectively. Clouds of size similar to these minimal
+% values are indeed much smaller than the beam width, and are therefore
+% unresolved.  This does not mean that there cannot be any smaller
+% clumps/clouds. It only means that any smaller clouds would be optically thin,
+% and we would thus not see any effect of these clumps on the unresolved
+% brightness: these clouds would have a spectral slope equal to that of an
+% optically thin non-clumpy model. At any rate, this estimate shows that, at least
+% in principle, it is possible that the rings we see in our data consist, in fact,
+% of numerous unresolved optically thick clouds interdispersed with optically thin
+% dust (or without any dust in between). Given that such an unresolved, but
+% optically thick, set of clumps has a lower spectral slope
+% $\alpha_{\mathrm{spec}}$ than an optically thin dusty medium, an analysis of the
+% measured spectral slope variations over the radial profile of the rings {\em
+%   may} give us some clues as to whether clumpiness occurs or not in these
+% rings, even if we cannot resolve the clumps themselves. We would expect that
+% within the ring the spectral slope is closer to $\alpha_{\mathrm{spec}}\simeq 2$
+% (modulo deviations from Rayleigh-Jeans) than in the inter-ring region.
+% 
+% However, the grain size segregation mechanism, that larger grains are more
+% concentrated in the ring than smaller grains, also tends to produce this
+% signature. This is particularly the case if the big grains (which are closest to
+% the ring peak) are larger than $\lambda/2\pi$ while the small grains are smaller
+% than $\lambda/2\pi$. To distinguish between the roles of size seggregation and
+% clumping, we need an independent way to probe the grain size. Perhaps the effects
+% of self-scattering induced polarization \citep{2015ApJ...809...78K} may provide
+% such an avenue. 
+% 
+% Even if we can distinguish between size-sorting and clumping, we still have to
+% clarify whether the required total dust mass for the clumping scenario is
+% reasonable, and whether the required 'filling factor' is consistent with
+% simulations of the streaming instability. Both questions are related. Let us go
+% back to the two prominent isolated rings of AS 209. The peak linear brightness
+% temperature is 3.5 $\mathrm{K}$ for the inner one and 2.7 $\mathrm{K}$ for the
+% outer one. In our model the disk midplane temperature is 13 $\mathrm{K}$ for the
+% inner and 10 $\mathrm{K}$ for the outer ring, which translates (for
+% $\lambda=0.13\,\mathrm{cm}$) into a linear brightness temperature of
+% 8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, respectively. If the disk were optically
+% thick, we would thus expect to observe a linear brightness temperature of
+% 8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, which are only about 2 times higher. 
+% 
+% 
+% This means that the filling factor of these clouds, as projected vertically
+% toward the observer, must be quite high, of the order of 50\%. {\bf [At this
+%     point we can calculate the minimum mass in dust needed for this]}
+% 
+% In principle the disk could be warmer than we assume, leading to a larger
+% saturation brightness, and thus a lower required filling factor. But given that
+% a 2$\times$ higher disk temperature requires a 16$\times$ higher irradiation
+% flux, we expect at most a factor of 2 uncertainty in the required filling
+% factor.
+% 
+% Checking whether this filling factor is consistent with the expectation of
+% simulations of the streaming instability is not easy. Such data are not
+% published for the models describing the streaming instability so far.  We can
+% only do a ``by eye'' check of the figures in those papers ({\bf give a few
+%   example papers here}). Such an exercise gives us the impression that the
+% typical fraction of the simulation box that is covered by the dense cluster
+% regions, as projected vertically, is typically substantially less than 50\%. If
+% this impression is true, then the ring emission seen in our sample cannot be
+% explained by a set of spatially unresolved optically thick clouds. That does not
+% imply, however, that the streaming instability does not operate in these
+% rings. It only implies that if it does, it will be hidden below an optically
+% thick shroud of more extended dust.
+% 
+% 
+% 
+% 
+% 
+% \subsection{Under which conditions should the streaming instability set in?}
+% While in Section \ref{sec-si-clump} we investigated if we can recognize the
+% ongoing streaming instability in our data if it is taking place in the disk, in
+% this Section we will compute, with our analytic dust trapping model, under which
+% condition we would {\em expect} that the streaming instability will operate in
+% the first place. This depends on the total dust mass locked in the trap, the
+% grain sizes and the turbulent strength. To make analytic estimates it is easiest
+% to assume a single grain size, instead of a grain size distribution.
+% 
+% {\bf FINISH THIS SECTION}
+% 
+%
+
+\section{Discussion}
+
+\subsection{Gap depth estimates from 2-D hydrodynamic simulations}
+
+{\bf Dong \& Fung (2017)} presented a systematic study of the gap
+opening profiles for given planetary masses, and analyzed the
+rings seen in scattered light of 5 sources (HD 97048, TW Hydra,
+HD 169142, LkCa15 and RX J1615).
+
+\subsection{Condition for the streaming instability}
+In the literature it is often mentioned that the streaming instability requires
+a dust-to-gas surface density ratio of
+$\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ or higher to operate {\bf
+  XXXXXXXX}. This can, however, not be directly compared to our models, because
+this value of $0.02$ was found for models without any pre-determined
+turbulence. The turbulence in those models was induced by the streaming
+instability itself or, if the streaming instability does not operate, by the
+Kelvin-Helmholtz instability ({\bf XXXX}). In the analytic model of this
+Section, on the other hand, we set the turbulence strength by hand, by setting
+$\alpha_{\mathrm{turb}}$ to some value. In essence, we assume that there is
+another source of turbulence, such as the magnetorotational instability ({\bf
+  XXXX}) or the vertical shear instability ({\bf XXXX}), that determines the
+mixing of the dust in the disk.
+
+According to \citet{2005ApJ...620..459Y} the true criterion for the onset of the
+streaming instability is the ratio of dust and gas {\em volume} densities
+$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$. For a given surface density
+ratio $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}$, the midplane volume density
+ratio, for a single grain species with midplane Stokes number $\mathrm{St}\ll
+1$, depends on the turbulent strength as
+\begin{equation}\label{eq-dtg-sig-vs-rho}
+  \frac{\rho_{\mathrm{d}}}{\rho_{\mathrm{g}}}\simeq
+  \left(1+\frac{\mathrm{St}}{\alpha_{\mathrm{turb}}}\right)
+  \frac{\Sigma_{\mathrm{d}}}{\Sigma_{\mathrm{g}}}
+\end{equation}
+The criterion of $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ mentioned
+in the literature thus relates to the criterion
+$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$ via the turbulent strength and
+the Stokes number. Given that we do not compute the turbulent strength
+but prescribe it, we should rely on the more fundamental volume density
+criterion of \citet{2005ApJ...620..459Y} to assess whether the dust in our
+model triggers the streaming instability or not.
+
+In our analytic setup of Section {\bf CHECK} %\ref{sec-anmodel-degree-trapping-alpha-series},
+the dust ring is about 5 times narrower than the gas bump. Assuming that
+initially the dust-to-gas ratio was $0.01$, this implies a dust-to-gas surface
+density ratio at the peak of the ring of $0.05$, which is in excess of the usual
+$0.02$ value. But as long as $\mathrm{St}\lesssim 19\alpha_{\mathrm{turb}}$,
+Eq.~(\ref{eq-dtg-sig-vs-rho}) shows that the midplane volume density ratio still
+stays below unity and no streaming instability sets in.
+
+\subsection{Can a resolved ring be in fact a blend of several unresolved rings?}
+Measuring the width of rings with finite-resolution observations is, of course,
+limited to ring widths that are larger than the ``beam''. However, even for
+cases where the ring width clearly exceeds the beam size by a factor of a few or
+more, one may ask the question: is the wide ring we see truly a single wide
+ring, or could it also be made up of a series of spatially unresolved rings that
+are blended into a single wide ring due to the beam convolution? This question
+is particularly relevant for the very wide rings of HD 163296. It is, of course,
+hard to answer in general, since we have no observational means to test this.
+
+But from the perspective of particle trapping by a pressure bump this question
+can be rigorously answered. A long-lived radial pressure perturbation in a
+protoplanetary disk cannot be much narrower than about a pressure scale height
+$h_p(r)$. A dust ring produced by dust trapping in this pressure bump may become
+rather narrow, dependent on a variety of parameters, as discussed in
+Section \ref{sec-steady-state-analytic-trap-model}. But there can not be
+more than a single such dust ring in each pressure bump.
+
+For the wide rings of HD 163296 the pressure scale height, even under the most
+optimistically low disk temperature (e.g.~10 K) the pressure scale height at
+rings 1 and 2 are 2.4 au and 4.3 au, respectively, which correspond to fwhm
+widths of 55 and 101 milliarcseconds, respectively (where we have multiplied by
+2.355 to obtain the full-width at half-maximum corresponding to a gaussian with
+$h_p$ as standard deviation width). Clearly the ALMA observations in band 6,
+with fwhm beam size of 35 mas, spatially resolve the pressure scale height.
+This means that the ring separation will be spatially resolved by ALMA,
+ruling out the possibility that the wide rings are made up of a multitude of
+narrow rings, at least in the dust trapping scenario.
+
+\subsection{Dust trapping in the horseshoe region and in resonances}
+The 1-D approach of the {\sf DISKLAB} models make it impossible to treat the
+trapping of particles in the horseshoe region. The low-$\alpha$ models of {\bf
+  Liu et al.~2018} show that this may be a significant effect. However, this
+tends to produce non-axisymmetric rings: there will clearly be a dearth of dust
+near the location of the planet, and a slight enhancement of dust near the L4
+and L5 Lagrange points. Given that the rings we see so far in the multi-ringed
+protoplanetary disks are angularly symmetric, this appears to be ruled out.
+That means that the 1-D approach is justified, but we do have to ensure that we
+are not in the regime where such non-axisymmetric horseshoe-trapping is
+important. Generally the horseshoe region tends to gradually deplete over time,
+but the time scale depends on the turbulent viscosity, with low-$\alpha$ disks
+depleting slower.
+
+Another issue is the potential for large particles to get trapped in a
+resonance of a planet instead of in a classical dust trap {\bf REFERENCE;
+the Dong double-gap is different, though}.
+This requires high Stokes number, generally $\mathrm{St}\gg 1$ ({\bf is
+  this true?}). Given that grain growth models tend to keep grains below
+this limit ({\bf see e.g.~Birnstiel et al.}), this is less likely to play
+a role.
+
+\subsection{Multiple gaps with a single planet}
+{\bf Dong, Li \& Li (2017)} and 
+{\bf Bae, Zhu \& Hartmann (2017)} show that for very low-$\alpha$ disks
+a planet can open multiple gaps in the disk rather than just a single one.
+The contrast of these rings and gaps in the gas can be large (factor of 2; see
+Bae et al.),
+and thus clearly constitute dust traps. Dong et al.\ show that these gaps
+can even be observable for very low mass planets.
+
+{\bf [TODO: We can do a simple model of dust trapping for their case.]}
+
+{\bf Fung \& Chiang (2017)} have shown that low mass planets in nearly inviscid
+disks can avoid migration and can be responsible for disk accretion in spite
+of the low viscosity.
+
+\subsection{Interpretation of the millimeter flux in terms of grain size}
+A common rule of thumb for interpreting millimeter and sub-millimeter
+fluxes and intensities from thermal dust emission is that you ``observe
+the grain size equal to the wavelength you observe at''. This rule is
+based on the tendency of the dust opacity derived from a Mie calculation
+to peak at wavelength around $\lambda\sim 2\pi a$, with $a$ the grain
+radius. For $\lambda \ll 2\pi a$ the opacity becomes nearly constant
+while $\lambda \gg 2\pi a$ it drops. This effect is even stronger for
+the scattering opacity. The validity of this rule of thumb relies,
+however, on the grain size distribution itself. For the simple
+{\bf Ivezic et al.~(1997)} opacity law, it is only valid for
+grain size distribution powerlaws obeying {\bf XXXXXXXXXXX}.
+{\bf [TODO: Maybe here do an experiment with full opacity laws.]}
+
+\begin{acknowledgements}
+  The authors acknowledge support
+  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
+  Datenverarbeitung of the University of T\"ubingen, the state of
+  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
+  through grant no INST 37/935-1 FUGG. This research was initiated at the
+  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
+  April 2018. Part of this work was also funded by the DFG Forschergruppe
+  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
+\end{acknowledgements}
+
+\begingroup
+\bibliographystyle{aa}
+\bibliography{ms}
+\endgroup
+
+\appendix
 
@@
 %
-Given the nicely separated narrow rings in the above sources, we can analyze
-these rings on a one-by-one basis. This is what we will do in this Section. We
-will assume that the dust trapping in these rings is absolute: no dust escapes.
-But increased turbulent mixing can spread the dust over a broader ring. Before
-we start constructing a model, we will discuss an important feature of the
-observed rings: that they appear to be radially resolved, Gaussian in shape, yet
-sub-Planckian in brightness. We will present a simple analytic dust trapping
-model, and show which radial brightness profiles it predicts, depending on the
-dust size distribution, the turbulent strength and the gas pressure bump width.
-This, when compared to the data, will allow us to draw some conclusions about
-the size distribution {\bf [and perhaps more?]}.
-
-\subsection{Resolved ring width and the brightness temperature}
-If our finding is correct that the rings in our sample are all spatially
-resolved in radial direction, then this poses an interesting question. In many
-cases the brightness of these rings is arguably sub-Planckian. This is, for
-instance, the case for the high-contrast rings of AS 209, as well as {\bf
-  XXXXXXX}. For the outer two rings of AS 209 (at 77 au and 125 au,
-respectively, the non-linear brightness temperatures are $3.5\,\mathrm{K}$ and
-$2.8\,\mathrm{K}$, respectively. If these rings would have been {\em
-  un}resolved, then this sub-Planckian emission is easily explained by the ring
-not filling the beam. But because these rings are radially resolved, this
-scenario should be rejected. The remaining possibilities are that the disk is
-colder than we think or that the rings are optically thin. In the latter case
-the spectral slope $\alpha_{\mathrm{spec}}$ at the location of these rings, once
-measured, will determine if the grains that are trapped in this ring are
-smaller, similar or larger than about $\lambda/2\pi$, where $\lambda$ is the
-wavelength around which the slope is measured. If we use
-$\lambda=0.13\,\mathrm{cm}$ and $\lambda=0.31\,\mathrm{cm}$ (representing ALMA
-band 6 and 3, respectively) to determine $\alpha_{\mathrm{spec}}$ this would
-mean we can distinguish between grains smaller, similar or larger than about 300
-$\mu$m. A low $\alpha_{\mathrm{spec}}$ would imply grains that are substantially
-larger than 300 $\mu$m, i.e.~of the order of millimeter or larger.
-
-In several recent observations of the spectral index across ringed disks
-\citep{2015ApJ...808L...3A,2018ApJ...852..122H} one clearly sees that
-$\alpha_{\mathrm{spec}}$ varies across these rings, being closer to $2$ at the
-ring center and substantially larger between the rings. This makes sense in
-terms of the dust trapping scenario in which we expect larger grains to be
-trapped more efficiently (and thus dominate the peak of the ring) than smaller
-grains, because the smaller grains will be more subject to turbulent
-mixing. This scenario requires a grain size distribution, so that the width of
-the dust ring is smaller for the bigger grains and bigger for the smaller ones.
-
-While this scenario is plausible and appealing, one may worry that it requires
-some fine-tuning. Why would the rings be dominated by a population of grains
-that are larger than 300 $\mu$m, but not too large, for else they would be so
-effectively trapped that they would produce an even narrower ring? For one or
-two sources that could be coincidence, but for a multitude of sources this is
-less plausible. One possible explanation is that millimeter grains are indeed
-the largest grains in the outer regions protoplanetary disks, because larger
-grains would easily get destroyed and/or simply bounce instead of stick when
-they collide, and thus not grow further \citep{2010A&A...513A..56G,
-  2010A&A...513A..57Z}. These models also predict larger grains to drift quickly
-inward, but that is only valid for a model without dust traps. At any rate:
-perhaps millimeter grains are a natural consequence of dust growth. If so, it
-then remains to be understood why the combination of turbulent strength and gas
-surface density conspire to trap these millimeter grains, but not too much, for
-else we might expect at least some rings to be narrower than a pressure scale
-height. We are aware that, even with a sample of 20 sources, we are still in the
-small-number-statistics regime, so these considerations are limited in their
-strength.
-
-In this section we will study this question using a very simple model of a dust
-trap: a parameterized gaussian pressure profile. The goal is to find out if the
-variations of $\alpha_{\mathrm{spec}}$ across the ring are naturally produced,
-while still keeping the dust ring radially resolved, without having to fine-tune
-the size distribution of the grains, the turbulence of the gas or other
-parameters. We will do this in several steps, starting with analytic solutions
-in Subsection \ref{sec-analytic-trap}, and proceeding with DISKLAB models in
-Subsection \ref{sec-disklab-trap}.
-
 \subsection{Analytic approximate solution of dust trapping}
@@
 midplane pressure given by
-\begin{equation}\label{eq-gaussian-pressure-bump}
+\begin{equation}\label{eq-gaussian-pressure-bump-repeat}
 p(r) = p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right)
 \end{equation}
-where $w\ll r_0$ is the parameter setting the width of this gaussian ring.
-We assume that the gas is weakly turbulent with turbulent diffusion coefficient
+where $w\ll r_0$ is the parameter setting the width of this gaussian gas ring.
+We assume that the gas is turbulent with turbulent diffusion coefficient
 $D$. Dust grains get trapped in this ring, and the dust will acquire a
@@
 pointing away from that position. The radial dust drift velocity is
-\citep[see e.g.][]{}:
+\citep[see e.g.][]{2010A&A...513A..79B}:
 \begin{equation}\label{eq-v-radial-drift}
-  v_{\mathrm{dr}} = \frac{1}{\mathrm{St}+\mathrm{St}^{-1}}
+  v_{\mathrm{dr}} = \frac{\mathrm{St}}{1+\mathrm{St}^{2}}
   \left(\frac{d\ln p}{d\ln r}\right)\frac{c_s^2}{\Omega_Kr}
@@
 defined as
-\begin{equation}
+\begin{equation}\label{eq-definition-stokes-number}
 \mathrm{St} = \Omega_Kt_{\mathrm{stop}}
@@
 width $w_d$ that is substantially smaller than the width of the gas ring
-$w$. Also the dust will all be settled vertically into a layer that has a
-vertical thickness less than the gas pressure scale height. In that case the
-solution to the drift-mixing equation for the dust becomes simple, because
-we can then assume everything to be constant with $r$ (within this limited
-domain), except for the gas pressure gradient. We will also ignore any terms
+$w$. In the following, we will ignore any terms
 arising from the curvature of the coordinates. The steady-state radial
@@
 \end{equation}
-From Eqs.(\ref{eq-v-radial-drift},\ref{eq-gaussian-pressure-bump}) we can express
+From Eqs.(\ref{eq-v-radial-drift},\ref{eq-gaussian-pressure-bump-repeat}) we can express
 $v_{\mathrm{dr}}$ as
@@
 With this expression we can solve Eq.~(\ref{eq-drift-mix-equil-eq}) for
-$\Sigma_{\mathrm{d}}$:
+$\Sigma_{\mathrm{d}}$, leading to the following simple analytic solution to the dust
+trapping problem:
 \begin{equation}\label{eq-analytic-sol-radial-trapping}
@@
 \begin{equation}
-  w_{\mathrm{d}} = w \sqrt{\frac{\Omega_KD_{\mathrm{d}}(\mathrm{St}+\mathrm{St}^{-1})}{c_s^2}}
-  = w\sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
-\end{equation}
-Note that this is only valid as long as $\alpha_{\mathrm{turb}}\ll \mathrm{Sc}\,\mathrm{St}$.
-This solution is, in fact, the radial version of the vertical settling-mixing
-equilibrium solutions of \citet{1995Icar..114..237D}.
-
-Now let us introduce a grain size distribution. Since the trapping width
-$w_{\mathrm{d}}$ for the different grain sizes is different, we can only
-impose the grain size distribution for the total (radially integrated)
-dust mass. We define the size distribution according to the following powerlaw
+  w_{\mathrm{d}} = w\, \sqrt{\frac{\Omega_KD_{\mathrm{d}}(\mathrm{St}+\mathrm{St}^{-1})}{c_s^2}}
+  = w\,\sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
+\end{equation}
+The normalization constant $\Sigma_{\mathrm{d0}}$ can be approximately expressed
+in terms of the total dust mass trapped in the pressure bump:
+\begin{equation}
+  M_d = 2\pi \int_0^\infty \Sigma_{\mathrm{d}}(r) rdr \simeq
+  2\pi r_0 \int_0^\infty \Sigma_{\mathrm{d}}(r)dr
+\end{equation}
+which leads to
+\begin{equation}
+\Sigma_{d0}\simeq \frac{M_d}{(2\pi)^{3/2}\, r_0\,w_{\mathrm{d}}}
+\end{equation}
+The approximation is best for narrow dust rings.
+
+Note that this analytic solution is only valid as long as
+$\alpha_{\mathrm{turb}}\ll \mathrm{Sc}\,\mathrm{St}$, or in other words as long
+as $w_{\mathrm{d}}$ is substantially smaller than $w$.  This solution is, in
+fact, the radial version of the vertical settling-mixing equilibrium solutions
+of \citet{1995Icar..114..237D}.
+
+Unfortunately, the condition that $\alpha_{\mathrm{turb}}\ll
+\mathrm{Sc}\,\mathrm{St}$ (and equivalently $w_d\ll w$) is easily broken for
+small grains and/or non-weak turbulence. This is similar to the case for
+vertical settling and mixing: Small grains tend not to settle below a few
+pressure scale heights. They settle down to some critical $z_{\mathrm{settle}}$,
+below which they are well mixed with the gas, and above which they are strongly
+depleted \citep{2004A&A...421.1075D}. \citet{2009A&A...496..597F} give an
+analytic solution to the settling-mixing problem (their Eq.~19) that reproduces
+the solutions of \citet{1995Icar..114..237D} for $\alpha_{\mathrm{turb}}\ll
+\mathrm{Sc}\,\mathrm{St}$, but at the same time remains valid also for
+$\alpha_{\mathrm{turb}}\gtrsim \mathrm{Sc}\,\mathrm{St}$, as long as
+$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant with
+height above the midplane.
+
+Given the similarity between the radial dust trapping problem and the vertical
+settling problem, one can show that the radial version of the solution of
+\citet{2009A&A...496..597F} reads:
+\begin{equation}\label{eq-analytic-sol-radial-trapping-better}
+  \Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\Bigg[
+    -\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}}
+     \left(\exp\left(\frac{\Delta r^2}{2w^2}\right)-1\right)
+    -\frac{\Delta r^2}{2w^2}\Bigg]
+\end{equation}
+where we defined $\Delta r$ as
+\begin{equation}
+\Delta r \equiv (r-r_0)
+\end{equation}
+The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) is valid for
+any value of $\alpha_{\mathrm{turb}}/\mathrm{Sc}\,\mathrm{St}$, as long as
+$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant along
+the radial width of the dust trap, and as long as $w\ll r_0$, to prevent
+geometric terms from the cilindrical coordinates from dominating.  One can
+easily verify that Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) reproduces
+the simpler Gaussian solution Eq.~(\ref{eq-analytic-sol-radial-trapping}) for
+$\Delta r\ll w$. One can also verify that for $\alpha_{\mathrm{turb}}\gg
+\mathrm{Sc}\,\mathrm{St}$ the shape of $\Sigma_{\mathrm{d}}(r)$ follows the
+shape of the gas pressure profile $p(r)$ (Eq.~\ref{eq-gaussian-pressure-bump-repeat}).
+
+The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better}), like its
+simplified version Eq.~(\ref{eq-analytic-sol-radial-trapping}), assumes that the
+dust is always near the midplane of the disk, so that one can ignore the
+vertical extent, and thereby ignore any variations in $\mathrm{St}$. In other
+words, we assume that the vertical scale height of the dust layer $h_d$ is much
+less than the gas pressure scale height $h_p$. As it turns out, this condition
+$h_d\ll h_p$ is satisfied if $\alpha_{\mathrm{turb}}\ll
+\mathrm{Sc}\,\mathrm{St}$, which is the same condition that leads to $w_d\ll w$
+(at least for $\mathrm{St}\ll 1$, which is what we assume).
+In fact, the radial dust ring width compared to the gas ring width is the
+same as the vertical dust layer thickness compared to the gas disk thickness:
+\begin{equation}
+\frac{w_d}{w} = \frac{h_d}{h_p}
+\end{equation}
+In other words: the dust behaves radially the same way as vertically. The
+vertical dust settling can be regarded as dust trapping in vertical direction,
+or the radial dust trapping can be seen as radial settling of dust toward
+the pressure peak.
+
+This also means that whenever the simplified Gaussian analytic solution
+(Eq.~\ref{eq-analytic-sol-radial-trapping}) ceases to be valid, so does the
+simplified vertical Gaussian solution of \citet{1995Icar..114..237D}. Or in
+other words, whenever the more sophisticated analytic solution
+Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) should be used instead of the
+simplified version, the assumption that the dust is settled near the midplane
+ceases to hold. Strictly speaking this invalidates
+Eq.~(\ref{eq-analytic-sol-radial-trapping-better}), since this was derived under
+the assumption that the dust is all located near the midplane. But given that
+most of the dust mass is still close to the midplane, even for
+$\alpha_{\mathrm{turb}}\gtrsim \mathrm{Sc}\,\mathrm{St}$, we are confident that
+Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) still remains approximately
+valid.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.47\textwidth]{analytic_trap_example.pdf}}
+\caption{\label{fig-analytic-model-example}The analytic solution shown for three
+  different grain sizes: small ($0.1\,\mu\mathrm{m}$), middle middle
+  ($10\,\mu\mathrm{m}$) and large ($1000\,\mu\mathrm{m}$). The normalizations of
+  the three curves are arbitrary, optimized for clarity. The solid (blue) curves
+  show the full analytic solution
+  (Eq.~\ref{eq-analytic-sol-radial-trapping-better}).  The dotted (green) curves
+  show the simplified analytic solution
+  (Eq.~\ref{eq-analytic-sol-radial-trapping}). The dashed (red) curves show the
+  dust profile {\em if} the dust-to-gas ratio would be constant, which shows the
+  shape of the gas pressure bump. The absense of the green dotted curve for the
+  case of $1000\,\mu\mathrm{m}$ grains (the narrowest peak) is because for these
+  large grains the simple analytic solution nearly perfectly overlaps with the
+  full analytic solution. The parameters are chosen for AS 209, ring 1,
+  with $M_{*}=0.912\,M_{\odot}$, $T_{\mathrm{gas}}=12.9\,\mathrm{K}$. The gas
+  surface density is taken to be $5\,\mathrm{g}/\mathrm{cm}^2$ at the position
+  of the ring, the turbulence parameter was set to
+  $\alpha_{\mathrm{turb}}=2\times 10^{-4}$, the dust material density was chosen
+  to be $2\,\mathrm{g}/\mathrm{cm}^3$ consistent with slightly porous silicate
+  dust aggregates. The standard-deviation width of the gas pressure bump was set
+  to $w=9.7\,\mathrm{au}$ which equals two times the gas pressure scale height.}
+\end{figure}
+
+The typical shapes of the analytic dust surface density profiles is shown in
+Fig.~\ref{fig-analytic-model-example} for three different grain sizes from small
+to large. For the largest grains, which are the most strongly trapped near the
+pressure peak, and hence have the narrowest peak in
+Fig.~\ref{fig-analytic-model-example}, the full analytic solution
+(Eq.~\ref{eq-analytic-sol-radial-trapping-better}) nearly perfectly overlaps
+with the simplified (Gaussian) analytic solution
+(Eq.~\ref{eq-analytic-sol-radial-trapping}). For large grains the simplified
+analytic solution is therefore sufficiently accurate. But in this parameter
+regime the dust rings are much narrower than the gas pressure bump ($w_d\ll w$).
+
+For the tiny grains, which are well coupled to the gas, and hence have the
+widest peak in Fig.~\ref{fig-analytic-model-example}, the dust trapping is very
+inefficient. Between 60 and 90 au the dust surface density follows nearly
+perfectly the shape of the gas bump, i.e.\ keeping the dust-to-gas ratio nearly
+constant. Only for $r\lesssim 60\,\mathrm{au}$ and $r\gtrsim 90\,\mathrm{au}$
+the full solution drops below a constant dust-to-gas ratio. To good approximation
+one can thus say that a dust ring made up of these tiny grains has a width
+that equals the width of the gas pressure bump. Note that in this regime the
+simplified analytic model (Eq.~\ref{eq-analytic-sol-radial-trapping}) fails
+miserably, as can be seen by the nearly horizontal green dotted curve in
+Fig.~\ref{fig-analytic-model-example}.
+
+The intermediate size grains, corresponding to the intermediate curve (marked
+with $100\,\mu\mathrm{m}$) in Fig.~\ref{fig-analytic-model-example}, have a
+behavior between the above two extremes. The simplified analytic solution is a
+bit wider than the more accurate solution. The shape of the more accurate
+solution deviates from Gaussian. However, the deviation from Gaussian shape of
+the model is opposite to the deviation from Gaussian shape in the observations
+(Fig.~\ref{fig-obs-gaussfits}). The observed data exceed the best-fit Gaussian
+in the flanks, while the model drops below the best-fit Gaussian in the flanks.
+
+Whether this deviation from Gaussian is detectable in the ALMA data is,
+however, not so clear. For intermediate size grains the non-Gaussian profile
+can be easily mistaken for a Gaussian profile with a narrower width than
+both the gas bump and the simple analytic solution.
+
+\subsection{Resulting profiles for a size distribution}
+\label{sec-anmodel-profiles}
+%
+
+{\bf [THIS SECTION NEEDS HEAVY EDITING]}
+
+The analytic model of the dust trapping for a single grain size speaks for
+itself. But how does it look for a grain size distribution? Since the trapping
+width $w_{\mathrm{d}}$ for the different grain sizes is different, we can only
+impose the grain size distribution for the total (radially integrated) dust
+mass. We define the size distribution according to the following powerlaw
 \begin{equation}
@@
 
-\subsection{Resulting profiles for a size distribution}
-\label{sec-anmodel-profiles}
-%
-The analytic model of the dust trapping for a single grain size speaks for
-itself. But how does it look for a grain size distribution? In
-Fig.~\ref{fig-anmodel-distribution-sigma} we show the radial profiles for an MRN size
-distribution between $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and
+In Fig.~\ref{fig-anmodel-distribution-sigma} we show the radial profiles for an
+MRN size distribution between $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and
 $a_{\mathrm{min}}=1\,\mathrm{cm}$ and a total dust mass of $M=1.3\times
@@
 distribution with 10 grain sizes, evenly distributed in $\ln(a)$. These are the
-10 curves in Fig.~\ref{fig-anmodel-distribution-sigma}. These surface densities are
-defined as being the local grain size distribution integrated over the width of
-the grain size bin:
+10 curves in Fig.~\ref{fig-anmodel-distribution-sigma}. These surface densities
+are defined as being the local grain size distribution integrated over the width
+of the grain size bin:
 $\Sigma_{\mathrm{bin}}=\int_{\mathrm{bin}}(d\Sigma_{\mathrm{dust}}/da)da$. The
@@
 
-\subsection{Effect of the degree of trapping on the appearance of the ring}
-\label{sec-anmodel-degree-trapping-alpha-series}
-%
-With this simple analytic dust trapping model we can investigate how the degree
-of trapping affects the appearance of the dust ring. For a model with a single
-grain size this is easy to predict: if the turbulence is weak enough that the
-gaussian profile of $\Sigma_{\mathrm{d}}(r)$ is narrower than the ALMA beam,
-then we observe a gaussian profile with the width of the beam. Even if the dust
-in this narrow ring would be optically thick, the observed ring would have
-sub-Planckian brightness due to beam-dilution. For stronger turbulence,
-$\Sigma_{\mathrm{d}}(r)$ is spread out wider, and we observe a gaussian profile
-with the width of the dust profile itself. As in the case of the strongly
-trapped dust, also the wide rust ring has a Gaussian profile, except that if it
-is optically thick, the observed profile is flattened at the top due to
-saturation to the Planck function.
-
-The situation gets more complex if we have a grain size distribution. For a
-given turbulence strength $\alpha_{\mathrm{turb}}$ some grains may be
-concentrated into a ring that is narrower than the ALMA beam, while others may
-be more spread out. The strongly concentrated grain population produces a cusp
-in the brightness, which will be partly smeared out by the beam convolution,
-while the weakly concentrated grain population produces a broad underlying
-brightness bump.
-
-To see how this looks in practice we set up a realization of our analytic
-model. The peak of the pressure is put at $r_{0}=77\,\mathrm{au}$ (corresponding
-to the inner of the two main rings of AS 209), and the width is chosen to be
-$w=30\,\mathrm{au}$ (corresponding to a FWHM of 70 au).  Given that the ALMA
-beam in band 6 has a FWHM of about 35 milliarcseconds, corresponding to 4.4 au
-for the distance to AS 209, the gas bump is much wider than the beam. The two
-main dust rings of AS 209 are about 3 beams wide, so that for our wide gas bump
-model the dust is strongly concentrated. This requires a low turbulent
-$\alpha_{\mathrm{turb}}$ or large grains. As in Section
-\ref{sec-anmodel-profiles} we choose a grain size distribution between
-$a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and $a_{\mathrm{min}}=1\,\mathrm{cm}$,
-following the usual MRN powerlaw of $q=1/2$ and a total dust mass of
-$M=1.3\times 10^{-5}\,M_\odot$. The dust material density is taken to be
-$\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
-distribution with 10 grain sizes, evenly distributed in $\ln(a)$. For the disk
-temperature we choose a simple powerlaw $T(r)=200\,(r/\mathrm{au})^{-1/2}$. The
-stellar mass is $M_{*}=0.9\,M_\odot$, and the source distance is taken to be
-$d=126\,\mathrm{pc}$. All these setting are taken to represent the inner of the
-two main high-contrast rings of the source AS 209. We set $\mathrm{Sc}=1$.
+% \subsection{Effect of the degree of trapping on the appearance of the ring}
+% \label{sec-anmodel-degree-trapping-alpha-series}
+% %
+% With this simple analytic dust trapping model we can investigate how the degree
+% of trapping affects the appearance of the dust ring. For a model with a single
+% grain size this is easy to predict: if the turbulence is weak enough that the
+% gaussian profile of $\Sigma_{\mathrm{d}}(r)$ is narrower than the ALMA beam,
+% then we observe a gaussian profile with the width of the beam. Even if the dust
+% in this narrow ring would be optically thick, the observed ring would have
+% sub-Planckian brightness due to beam-dilution. For stronger turbulence,
+% $\Sigma_{\mathrm{d}}(r)$ is spread out wider, and we observe a gaussian profile
+% with the width of the dust profile itself. As in the case of the strongly
+% trapped dust, also the wide rust ring has a Gaussian profile, except that if it
+% is optically thick, the observed profile is flattened at the top due to
+% saturation to the Planck function.
+% 
+% The situation gets more complex if we have a grain size distribution. For a
+% given turbulence strength $\alpha_{\mathrm{turb}}$ some grains may be
+% concentrated into a ring that is narrower than the ALMA beam, while others may
+% be more spread out. The strongly concentrated grain population produces a cusp
+% in the brightness, which will be partly smeared out by the beam convolution,
+% while the weakly concentrated grain population produces a broad underlying
+% brightness bump.
+% 
+% To see how this looks in practice we set up a realization of our analytic
+% model. The peak of the pressure is put at $r_{0}=77\,\mathrm{au}$ (corresponding
+% to the inner of the two main rings of AS 209), and the width is chosen to be
+% $w=30\,\mathrm{au}$ (corresponding to a FWHM of 70 au).  Given that the ALMA
+% beam in band 6 has a FWHM of about 35 milliarcseconds, corresponding to 4.4 au
+% for the distance to AS 209, the gas bump is much wider than the beam. The two
+% main dust rings of AS 209 are about 3 beams wide, so that for our wide gas bump
+% model the dust is strongly concentrated. This requires a low turbulent
+% $\alpha_{\mathrm{turb}}$ or large grains. As in Section
+% \ref{sec-anmodel-profiles} we choose a grain size distribution between
+% $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and $a_{\mathrm{min}}=1\,\mathrm{cm}$,
+% following the usual MRN powerlaw of $q=1/2$ and a total dust mass of
+% $M=1.3\times 10^{-5}\,M_\odot$. The dust material density is taken to be
+% $\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
+% distribution with 10 grain sizes, evenly distributed in $\ln(a)$. For the disk
+% temperature we choose a simple powerlaw $T(r)=200\,(r/\mathrm{au})^{-1/2}$. The
+% stellar mass is $M_{*}=0.9\,M_\odot$, and the source distance is taken to be
+% $d=126\,\mathrm{pc}$. All these setting are taken to represent the inner of the
+% two main high-contrast rings of the source AS 209. We set $\mathrm{Sc}=1$.
 
@@
 
-The results for several values of $\alpha_{\mathrm{turb}}$ are shown in
-Fig.~\ref{fig-anmodel-series-alpha}. In the top panel, representing weak
-turbulence, most of the dust is trapped into a ring more narrow than the
-beam. A clear cusp can be seen in the un-convolved brightness profile,
-which is due to the strongly trapped large grains of the size distribution.
-However, the convolution washes out this cusp for the most part. This
-situation is equivalent to the simple case of a single dust species being
-trapped to a narrow unresolved ring. The only difference is that for a single
-dust size the ring is either optically thick or optically thin, while in the
-multi-size case this can be a mix. That does not make a difference for the
-observed brightness profile, which is a Gaussian either way.
-
-For the case of the strongest turbulence, the bottom panel in
-Fig.~\ref{fig-anmodel-series-alpha}, the cusp is still seen in the un-convolved
-profile, but now the convolution does not wash it out completely. The resulting
-convolved brightness profile is clearly non-Gaussian as a result. It has a
-quasi-Gaussian cusp with broad wings. When seen with ALMA such a cuspy profile
-(narrow at the peak, broad in the wings) should be recognizable.
-
-These findings show that if the turbulent strength is weak, the predicted dust
-ring width equals the ALMA beam width, and the brightness shape should have the
-shape of the beam profile (here assumed to be Gaussian). For stronger
-turbulence, the dust ring becomes indeed broader than the ALMA beam, but then
-the shape of the observed brightness profile may teach us something about the
-grain size distribution. Roughly speaking, if the largest grains of the size
-distribution still get trapped in a narrow ring, this shows up as a cusp in the
-brightness profile. If the smallest grains of the grain size distribution are
-too small, the predicted brightness profile would be too broad compared to the
-observed ring width. In other words: if the observed brightness profile of a
-observed ring in given source is clearly radially resolved, and if it has a
-Gaussian shape, then the most likely fit to our analytic model is that of
-a narrow size distribution (quasi a single grain size) with the right turbulent
-strength to fit the observed ring width. A broader size distribution, with the
-appropriately chosen turbulent strength, would yield a cuspy profile, not a
-Gaussian one.
-
-The two high-contrast rings of AS 209 both appear to fit a Gaussian profile with
-2.5 times the width of the ALMA beam (Fig.~\ref{fig-as209-rings-gauss-fit}). In
-both cases it is therefore likely, according to this analysis, that the dust
-that is trapped in these rings has a narrow size distribution instead of a broad
-MRN powerlaw distribution.
+% The results for several values of $\alpha_{\mathrm{turb}}$ are shown in
+% Fig.~\ref{fig-anmodel-series-alpha}. In the top panel, representing weak
+% turbulence, most of the dust is trapped into a ring more narrow than the
+% beam. A clear cusp can be seen in the un-convolved brightness profile,
+% which is due to the strongly trapped large grains of the size distribution.
+% However, the convolution washes out this cusp for the most part. This
+% situation is equivalent to the simple case of a single dust species being
+% trapped to a narrow unresolved ring. The only difference is that for a single
+% dust size the ring is either optically thick or optically thin, while in the
+% multi-size case this can be a mix. That does not make a difference for the
+% observed brightness profile, which is a Gaussian either way.
+% 
+% For the case of the strongest turbulence, the bottom panel in
+% Fig.~\ref{fig-anmodel-series-alpha}, the cusp is still seen in the un-convolved
+% profile, but now the convolution does not wash it out completely. The resulting
+% convolved brightness profile is clearly non-Gaussian as a result. It has a
+% quasi-Gaussian cusp with broad wings. When seen with ALMA such a cuspy profile
+% (narrow at the peak, broad in the wings) should be recognizable.
+% 
+% These findings show that if the turbulent strength is weak, the predicted dust
+% ring width equals the ALMA beam width, and the brightness shape should have the
+% shape of the beam profile (here assumed to be Gaussian). For stronger
+% turbulence, the dust ring becomes indeed broader than the ALMA beam, but then
+% the shape of the observed brightness profile may teach us something about the
+% grain size distribution. Roughly speaking, if the largest grains of the size
+% distribution still get trapped in a narrow ring, this shows up as a cusp in the
+% brightness profile. If the smallest grains of the grain size distribution are
+% too small, the predicted brightness profile would be too broad compared to the
+% observed ring width. In other words: if the observed brightness profile of a
+% observed ring in given source is clearly radially resolved, and if it has a
+% Gaussian shape, then the most likely fit to our analytic model is that of
+% a narrow size distribution (quasi a single grain size) with the right turbulent
+% strength to fit the observed ring width. A broader size distribution, with the
+% appropriately chosen turbulent strength, would yield a cuspy profile, not a
+% Gaussian one.
+% 
+% The two high-contrast rings of AS 209 both appear to fit a Gaussian profile with
+% 2.5 times the width of the ALMA beam (Fig.~\ref{fig-as209-rings-gauss-fit}). In
+% both cases it is therefore likely, according to this analysis, that the dust
+% that is trapped in these rings has a narrow size distribution instead of a broad
+% MRN powerlaw distribution.
 
@@
 %
+
+{\bf [TO DO]}
+
+\subsection{Leakage and retention time of the dust in the traps}
+\label{sec-leakage}
+%
+In our analytic model we have assumed that the pressure bump has a perfect
+Gaussian shape, meaning that the further away one goes from the peak of the
+trap, the steeper the pressure gradient (and thereby the faster the dust drift)
+becomes. This leads to perfect trapping: no dust can leak out of the trap.
+In reality the pressure bump has a finite radial width, and smoothly
+connects to the global disk structure. There will be a maximum pressure
+gradient at some distance away from the peak, and beyond that the pressure
+gradient declines again. If the width of the trapped dust is much narrower
+than this distance, virtually no dust will get lost out of the trap. But
+for small grains and/or strong turbulence, the width of the dust profile may
+be large enough that at the point of largest pressure gradient there is
+still a substantial amount of dust. Then the dust will leak out of the
+trap.
+
+{\bf First make estimates based on current model. Then add the gas radial
+velocity by accretion.}
+
+\subsection{Stability of gas ring}
+\label{sec-gas-ring-stability}
+
+{\bf Here we must discuss how far into the wings of the gaussian is
+  stable against RWI and other instabilities. Solberg-Hoiland stability
+  criterion. A Gauss may not be very stabile in its flanks. This may limit
+the applicability of the full analytic solution.}
 
@@
 {\bf [CONTINUE THIS; SHOW THE FIT FIGURES; SHOW THE MCMC CORNER PLOTS]}
+
+{\bf For HD 163296 we have constraints from the sub/superkepler data of
+Teague. This limits the width of the gas bump.}
 
@@
 
-{\bf [TO BE DONE: So far the CMB or cloud background is not included in the model.
-    This has to be improved, because we are working at temperatures close to the
-    CMB. It might reduce the sub-Planckianness, I guess.]}
-
-{\bf [Question to the experts: The CMB and/or cloud background is automatically
-filtered out of the interferometric observations, I presume?]}
-
-\section{Streaming instability, clumping, and the spectral index}
-\label{sec-si-clump}
-%
-Whenever the dust trapping becomes so effective that the volume density of the
-dust at the midplane of the disk exceeds the gas volume density, the streaming
-instability may set in \citep{2005ApJ...620..459Y}. This produces turbulence and
-leads to clumping \citep{2007ApJ...662..627J, 2008A&A...479..883L,
-  2010ApJ...722.1437B}. If conditions are right, some of these clumps may
-gravitationally collapse and form planetesimals \citep{2007Natur.448.1022J,
-  2017ApJ...847L..12S, 2017A&A...597A..69S}. It is believed that this mechanism
-is the trigger that starts the process of the formation of planets
-\citep[see][for a review]{2014prpl.conf..547J}. If one could find observational
-evidence of this dust clumping occurring in protoplanetary disks, this would
-be a major step forward for our understanding of planet formation.
-
-The clumps that are formed in simulations of the streaming instability are,
-however, typically substantially smaller than a pressure scale height of the
-disk \citep{2007ApJ...662..627J, 2013MNRAS.434.1460K}. This makes it very hard
-to spatially resolve these structures, even with ALMA at its highest resolution.
-
-It might, however, be possible to find indications of particle clumping in an
-indirect way, if it leads to the formation of optically thick clouds of
-particles surrounded by optically thin regions. Even if we cannot spatially
-resolve these clouds, the spatially averaged thermal emission from such a clumpy
-medium has a different spectral slope than that of a homogeneous distribution of
-dust. If all the dust is concentrated in a multitude of small and very optically
-thick clumps, then the brightness of this unresolved clumpy medium would be the
-Planck function times the covering fraction $C$ of the clumps:
-\begin{equation}\label{eq-bright-clumpy}
-I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,C
-\end{equation}
-where the limit $C\rightarrow 0$ means concentrating all the dust in infinitely
-dense clumps so that the covering fraction is zero, and $C\rightarrow 1$ means
-having so many clumps that they multiply overlap, when seen in projection from
-above. In contrast, if all this dust is homogeneously distributed, then the
-brightness is
-\begin{equation}\label{eq-bright-homogeneous}
-I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,(1-e^{-\tau_\nu})
-\end{equation}
-with $\tau_\nu$ being the vertical optical depth of the dust in the disk.
-Reality will likely lie in between these two extreme scenarios, with
-optically thin dust interdispersed between optically thick clouds.
-
-There are at least two ways by which one could use Eqs.~(\ref{eq-bright-clumpy},
-\ref{eq-bright-homogeneous}) to search for hints of clumpiness. The most direct
-way is to make use of the different dependencies of these two equations on
-wavelength $\lambda=c/\nu$. In the case of the clumpy medium with perfectly
-optically thick clumps (Eq.~\ref{eq-bright-clumpy}) the brightness follows the
-Planck curve, whereas for the homogeneous medium
-(Eq.~\ref{eq-bright-homogeneous}), if it is optically thin
-(i.e.~$1-e^{-\tau_\nu}\simeq \tau_\nu$), it follows the Planck curve times the
-opacity law. If we are in the Rayleigh-Jeans domain of the Planck curve, then
-these two limits are usually referred to as spectral slope
-$\alpha_{\mathrm{spec}}=2$ and $\alpha_{\mathrm{spec}}=2+\beta$, respectively,
-where $\beta$ is defined by $\kappa_\nu\propto \nu^\beta$. At ALMA band 6, for
-the low temperatures we expect in these disks at large radii, we are no longer
-strictly in the Rayleigh-Jeans domain, so the spectral slope analysis becomes
-temperature-dependent, which complicates matters a bit, but the principle
-remains the same. If the streaming instability is operating in the rings of our
-sources, and if the numbers work out (see below), then we expect to see
-$\alpha_{\mathrm{spec}}\simeq 2$ within the rings, and
-$\alpha_{\mathrm{spec}}>2$ between the rings.
-
-There is an important caveat to this idea: the grain size and the corresponding
-opacity slope may mimick the same spectral signatures as the clumping. This is,
-in fact, the same degeneracy of the interpretation of the millimeter spectral
-slope as usual \citep[e.g.][]{2003A&A...403..323T}, just on a much smaller
-spatial scale: a scale that remains unresolved even by ALMA.  Nevertheless it is
-worthwhile to investigate if we can find clues that point to clumpiness.
-
-In the scope of this paper we will only do a very simple analysis: we will
-verify if, under the most benign conditions, the above mentioned optically
-thick/thin clumpy medium is at all possible. This is not granted, because
-these rings are several tens of au large, and therefore it requires a lot of
-mass to make dust clumps optically thick at 1.3 millimeter wavelength.
-Furthermore, these dust clouds cannot have a density in excess of the
-Roche density, because then they would immediately gravitationally collapse.
-Therefore, we cannot make the clumps arbitrary dense and compact.
-
-So let us construct the most optimistic model of a dust cloud. The highest
-possible density of a cloud is the Roche density, defined by
-\begin{equation}
-\rho_{\mathrm{Roche}}(r) = \frac{9}{4\pi}\frac{M_{*}}{r^3}
-\end{equation}
-where $M_{*}$ is the stellar mass and $r$ is the distance to the star. We take
-$\rho_{\mathrm{cloud}}=\rho_{\mathrm{Roche}}$. The simplest opacity model for a
-dust grain is that of \cite{1997MNRAS.291..121I}, where we omit the scattering
-part:
-\begin{equation}
-\kappa(a) = \frac{3}{4}\frac{1}{\xi\,a}\left\{\begin{matrix}
-1 & \hbox{for} & \lambda<2\pi a \\
-2\pi a/\lambda & \hbox{for} & \lambda>2\pi a \\
-\end{matrix}\right.
-\end{equation}
-where $a$ is the grain radius and $\xi$ is the material density of the dust
-grain ($\xi=3.6$ for amorphous olivine, $\xi=2.0$ for silica {\bf CHECK}). For a
-given wavelength $\lambda$ (in our case $\lambda=0.13\,\mathrm{cm}$ for ALMA
-band 6), the most optimistic opacity is therefore $\kappa=(3\pi/2)/\xi\lambda$,
-for all $a\le \lambda/2\pi$. For ALMA band 6 and $\xi=2\,\mathrm{g/cm}^3$ this
-is $\kappa=18.1\,\mathrm{cm}^2/\mathrm{g}$. The largest grain size to still
-have this opacity is therefore $a=\lambda/2\pi$, which for ALMA band 6 is
-$a=0.02\,\mathrm{cm}$.  We make a spherical cloud of this dust with radius
-$R$. The optical depth through the cloud center is then
-\begin{equation}
-\tau = 2R\rho_{\mathrm{cloud}}\kappa
-\end{equation}
-To obtain the effect of an optically thick/thin clumpy medium, these clouds
-have to be sufficiently optically thick, not just marginally. So let us demand
-a minimal optical depth of $\tau=4$. This gives a minimal cloud radius of
-\begin{equation}
-  R\gtrsim \frac{2}{\rho_{\mathrm{Roche}}\kappa} = 0.11\frac{1}{\rho_{\mathrm{Roche}}}
-  = 0.15\frac{r^3}{M_{*}}
-\end{equation}
-in CGS units (for ALMA band 6). Note that this scales linearly with
-$\rho_{\mathrm{cloud}}/\rho_{\mathrm{Roche}}$, so that if we choose a much less
-``critical'' cloud density of, say, $\rho_{\mathrm{cloud}}=0.1\,\rho_{\mathrm{Roche}}$,
-then the cloud should become 10$\times$ larger and 100$\times$ more massive.
-
-If we want these clouds to be part of an unresolved clumpy medium, they should
-be substantially smaller than the pressure scale height of the disk, and
-substantially smaller than the ALMA band 6 resolution. Let us take the example
-of the source AS 209, with $M_{*}=0.9\,M_{\odot}$, and its two high-contrast
-rings at 74 au and 120 au, respectively. This gives minimal cloud
-radii of $R\gtrsim 4\times 10^{-3}\,\mathrm{au}$ and $R\gtrsim 22\times
-10^{-3}\,\mathrm{au}$, respectively. Clouds of size similar to these minimal
-values are indeed much smaller than the beam width, and are therefore
-unresolved.  This does not mean that there cannot be any smaller
-clumps/clouds. It only means that any smaller clouds would be optically thin,
-and we would thus not see any effect of these clumps on the unresolved
-brightness: these clouds would have a spectral slope equal to that of an
-optically thin non-clumpy model. At any rate, this estimate shows that, at least
-in principle, it is possible that the rings we see in our data consist, in fact,
-of numerous unresolved optically thick clouds interdispersed with optically thin
-dust (or without any dust in between). Given that such an unresolved, but
-optically thick, set of clumps has a lower spectral slope
-$\alpha_{\mathrm{spec}}$ than an optically thin dusty medium, an analysis of the
-measured spectral slope variations over the radial profile of the rings {\em
-  may} give us some clues as to whether clumpiness occurs or not in these
-rings, even if we cannot resolve the clumps themselves. We would expect that
-within the ring the spectral slope is closer to $\alpha_{\mathrm{spec}}\simeq 2$
-(modulo deviations from Rayleigh-Jeans) than in the inter-ring region.
-
-However, the grain size segregation mechanism, that larger grains are more
-concentrated in the ring than smaller grains, also tends to produce this
-signature. This is particularly the case if the big grains (which are closest to
-the ring peak) are larger than $\lambda/2\pi$ while the small grains are smaller
-than $\lambda/2\pi$. To distinguish between the roles of size seggregation and
-clumping, we need an independent way to probe the grain size. Perhaps the effects
-of self-scattering induced polarization \citep{2015ApJ...809...78K} may provide
-such an avenue.
-
-Even if we can distinguish between size-sorting and clumping, we still have to
-clarify whether the required total dust mass for the clumping scenario is
-reasonable, and whether the required 'filling factor' is consistent with
-simulations of the streaming instability. Both questions are related. Let us go
-back to the two prominent isolated rings of AS 209. The peak linear brightness
-temperature is 3.5 $\mathrm{K}$ for the inner one and 2.7 $\mathrm{K}$ for the
-outer one. In our model the disk midplane temperature is 13 $\mathrm{K}$ for the
-inner and 10 $\mathrm{K}$ for the outer ring, which translates (for
-$\lambda=0.13\,\mathrm{cm}$) into a linear brightness temperature of
-8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, respectively. If the disk were optically
-thick, we would thus expect to observe a linear brightness temperature of
-8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, which are only about 2 times higher.
-
-{\bf [TODO: Handle the CMB and cloud background, which might reduce the
-sub-Planckianness we observe.]}
-
-This means that the filling factor of these clouds, as projected vertically
-toward the observer, must be quite high, of the order of 50\%. {\bf [At this
-    point we can calculate the minimum mass in dust needed for this]}
-
-In principle the disk could be warmer than we assume, leading to a larger
-saturation brightness, and thus a lower required filling factor. But given that
-a 2$\times$ higher disk temperature requires a 16$\times$ higher irradiation
-flux, we expect at most a factor of 2 uncertainty in the required filling
-factor.
-
-Checking whether this filling factor is consistent with the expectation of
-simulations of the streaming instability is not easy. Such data are not
-published for the models describing the streaming instability so far.  We can
-only do a ``by eye'' check of the figures in those papers ({\bf give a few
-  example papers here}). Such an exercise gives us the impression that the
-typical fraction of the simulation box that is covered by the dense cluster
-regions, as projected vertically, is typically substantially less than 50\%. If
-this impression is true, then the ring emission seen in our sample cannot be
-explained by a set of spatially unresolved optically thick clouds. That does not
-imply, however, that the streaming instability does not operate in these
-rings. It only implies that if it does, it will be hidden below an optically
-thick shroud of more extended dust.
-
-\subsection{Under which conditions should the streaming instability set in?}
-While in Section \ref{sec-si-clump} we investigated if we can recognize the
-ongoing streaming instability in our data if it is taking place in the disk, in
-this Section we will compute, with our analytic dust trapping model, under which
-condition we would {\em expect} that the streaming instability will operate in
-the first place. This depends on the total dust mass locked in the trap, the
-grain sizes and the turbulent strength. To make analytic estimates it is easiest
-to assume a single grain size, instead of a grain size distribution.
-
-{\bf FINISH THIS SECTION}
-
-\section{Discussion}
-
-\subsection{Gap depth estimates from 2-D hydrodynamic simulations}
-
-{\bf Dong \& Fung (2017)} presented a systematic study of the gap
-opening profiles for given planetary masses, and analyzed the
-rings seen in scattered light of 5 sources (HD 97048, TW Hydra,
-HD 169142, LkCa15 and RX J1615).
-
-\subsection{Condition for the streaming instability}
-In the literature it is often mentioned that the streaming instability requires
-a dust-to-gas surface density ratio of
-$\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ or higher to operate {\bf
-  XXXXXXXX}. This can, however, not be directly compared to our models, because
-this value of $0.02$ was found for models without any pre-determined
-turbulence. The turbulence in those models was induced by the streaming
-instability itself or, if the streaming instability does not operate, by the
-Kelvin-Helmholtz instability ({\bf XXXX}). In the analytic model of this
-Section, on the other hand, we set the turbulence strength by hand, by setting
-$\alpha_{\mathrm{turb}}$ to some value. In essence, we assume that there is
-another source of turbulence, such as the magnetorotational instability ({\bf
-  XXXX}) or the vertical shear instability ({\bf XXXX}), that determines the
-mixing of the dust in the disk.
-
-According to \citet{2005ApJ...620..459Y} the true criterion for the onset of the
-streaming instability is the ratio of dust and gas {\em volume} densities
-$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$. For a given surface density
-ratio $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}$, the midplane volume density
-ratio, for a single grain species with midplane Stokes number $\mathrm{St}\ll
-1$, depends on the turbulent strength as
-\begin{equation}\label{eq-dtg-sig-vs-rho}
-  \frac{\rho_{\mathrm{d}}}{\rho_{\mathrm{g}}}\simeq
-  \left(1+\frac{\mathrm{St}}{\alpha_{\mathrm{turb}}}\right)
-  \frac{\Sigma_{\mathrm{d}}}{\Sigma_{\mathrm{g}}}
-\end{equation}
-The criterion of $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ mentioned
-in the literature thus relates to the criterion
-$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$ via the turbulent strength and
-the Stokes number. Given that we do not compute the turbulent strength
-but prescribe it, we should rely on the more fundamental volume density
-criterion of \citet{2005ApJ...620..459Y} to assess whether the dust in our
-model triggers the streaming instability or not.
-
-In our analytic setup of Section \ref{sec-anmodel-degree-trapping-alpha-series},
-the dust ring is about 5 times narrower than the gas bump. Assuming that
-initially the dust-to-gas ratio was $0.01$, this implies a dust-to-gas surface
-density ratio at the peak of the ring of $0.05$, which is in excess of the usual
-$0.02$ value. But as long as $\mathrm{St}\lesssim 19\alpha_{\mathrm{turb}}$,
-Eq.~(\ref{eq-dtg-sig-vs-rho}) shows that the midplane volume density ratio still
-stays below unity and no streaming instability sets in.
-
-\subsection{Dust trapping in the horseshoe region and in resonances}
-The 1-D approach of the {\sf DISKLAB} models make it impossible to treat the
-trapping of particles in the horseshoe region. The low-$\alpha$ models of {\bf
-  Liu et al.~2018} show that this may be a significant effect. However, this
-tends to produce non-axisymmetric rings: there will clearly be a dearth of dust
-near the location of the planet, and a slight enhancement of dust near the L4
-and L5 Lagrange points. Given that the rings we see so far in the multi-ringed
-protoplanetary disks are angularly symmetric, this appears to be ruled out.
-That means that the 1-D approach is justified, but we do have to ensure that we
-are not in the regime where such non-axisymmetric horseshoe-trapping is
-important. Generally the horseshoe region tends to gradually deplete over time,
-but the time scale depends on the turbulent viscosity, with low-$\alpha$ disks
-depleting slower.
-
-Another issue is the potential for large particles to get trapped in a
-resonance of a planet instead of in a classical dust trap {\bf REFERENCE;
-the Dong double-gap is different, though}.
-This requires high Stokes number, generally $\mathrm{St}\gg 1$ ({\bf is
-  this true?}). Given that grain growth models tend to keep grains below
-this limit ({\bf see e.g.~Birnstiel et al.}), this is less likely to play
-a role.
-
-\subsection{Multiple gaps with a single planet}
-{\bf Dong, Li \& Li (2017)} and 
-{\bf Bae, Zhu \& Hartmann (2017)} show that for very low-$\alpha$ disks
-a planet can open multiple gaps in the disk rather than just a single one.
-The contrast of these rings and gaps in the gas can be large (factor of 2; see
-Bae et al.),
-and thus clearly constitute dust traps. Dong et al.\ show that these gaps
-can even be observable for very low mass planets.
-
-{\bf [TODO: We can do a simple model of dust trapping for their case.]}
-
-{\bf Fung \& Chiang (2017)} have shown that low mass planets in nearly inviscid
-disks can avoid migration and can be responsible for disk accretion in spite
-of the low viscosity.
-
-\subsection{Interpretation of the millimeter flux in terms of grain size}
-A common rule of thumb for interpreting millimeter and sub-millimeter
-fluxes and intensities from thermal dust emission is that you ``observe
-the grain size equal to the wavelength you observe at''. This rule is
-based on the tendency of the dust opacity derived from a Mie calculation
-to peak at wavelength around $\lambda\sim 2\pi a$, with $a$ the grain
-radius. For $\lambda \ll 2\pi a$ the opacity becomes nearly constant
-while $\lambda \gg 2\pi a$ it drops. This effect is even stronger for
-the scattering opacity. The validity of this rule of thumb relies,
-however, on the grain size distribution itself. For the simple
-{\bf Ivezic et al.~(1997)} opacity law, it is only valid for
-grain size distribution powerlaws obeying {\bf XXXXXXXXXXX}.
-{\bf [TODO: Maybe here do an experiment with full opacity laws.]}
-
-\begin{acknowledgements}
-  The authors acknowledge support
-  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
-  Datenverarbeitung of the University of T\"ubingen, the state of
-  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
-  through grant no INST 37/935-1 FUGG. This research was initiated at the
-  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
-  April 2018. Part of this work was also funded by the DFG Forschergruppe
-  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
-\end{acknowledgements}
-
-\begingroup
-\bibliographystyle{aa}
-\bibliography{ms}
-\endgroup
+\subsection{Computing the dust mass trapped in the rings}
+\label{sec-computing-dust-mass}
+%
+We can compute the dust mass trapped in each of the rings by integrating over
+the measured brightness temperature profiles. However, the profiles do not go
+down to zero far way from the peak, and so it is not well defined what belongs
+to the ring and what not. Instead, we will compute the dust mass from the
+Gaussian profile that is fitted to the ring emission. This will slightly
+underestimate the dust mass, because the excess emission in the flanks of the
+profiles (see Fig.~\ref{fig-obs-gaussfits}) will not be included. But there are
+far more serious issues to worry about, such as our limited knowledge of the
+disk temperature, the uncertainties of the dust opacities, as well as optical
+depth effects. So we will estimate the mass of the best fitting Gaussian model
+only.
+
+We will assume that the intrinsic dust temperature is as given in Table
+\ref{tab-gauss-params} (column $T_{\mathrm{model}}$), which was computed based on
+a simple flaring angle irradiation model. We consider the Gaussian profile with
+the deconvolved width, i.e.\ $w_d$ (not $\sigma$). The peak linear brightness
+temperature is taken to be
+\begin{equation}
+A_{\mathrm{deconv}} = \frac{\sigma}{w_d}\,A
+\end{equation}
+where $A$, $\sigma$ and $w_d$ are from Table \ref{tab-gauss-params}. We do not
+account for geometric terms in the deconvolution, which is justified as long as
+$\sigma\ll r$. The deconvolved Gaussian brightness profile is then
+\begin{equation}\label{eq-tbrlin-deconvolved}
+T_{\mathrm{br,lin,deconv}}(r) = A_{\mathrm{deconv}}\,\exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+\end{equation}
+
+Next we have to worry about optical depth effects. From Table
+\ref{tab-gauss-params} one can see that for most rings 
+$A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{model,br,lin}}$.
+However, for the rings of AS 209 the ratios
+$A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}$ are around 0.5, so optical depth
+effects start to play a role. Given a dust surface density profile
+$\Sigma_{\mathrm{d}}(r)$ as a function of radial coordinate $r$, and
+given an absorption opacity $\kappa_{\mathrm{a}}$ at the wavelength we
+are measuring, the absorption optical depth $\tau_{\mathrm{a}}(r)$ is
+\begin{equation}\label{eq-tau-afo-sig-kap}
+\tau_{\mathrm{a}}(r) = \Sigma_{\mathrm{d}}(r)\kappa_{\mathrm{a}}
+\end{equation}
+We assume that scattering does not play a role. The emission from this
+dusty ring is then, expressed as linear brightness temperature:
+\begin{equation}
+T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{model,br,lin}}
+\end{equation}
+If $\tau_{\mathrm{a}}(r)\ll 1$ for all $r$ this reduces to
+\begin{equation}\label{eq-int-afo-tau-bnu}
+T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{model,br,lin}}
+\end{equation}
+But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}
+=\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{model}}))$ can be as large as
+$\sim 0.5$, this condition is not always fulfilled.  On the other hand, these
+optical depth effects cannot be too strong, because then the Gaussian shape
+would flatten off at the top, which is not what we see in the observations. We
+will therefore continue our computation without accounting for these (mild)
+optical depth effects. This will lead to a slight underestimation of the
+mass. The reader can estimate the magnitude of this underestimation by comparing
+$(1-\exp(-\tau_{\mathrm{a}}))$ to $\tau_{\mathrm{a}}$.
+
+Using Eq.~(\ref{eq-int-afo-tau-bnu}) we can, from the deconvolved
+$T_{\mathrm{br,lin,deconv}}(r)$ (Eq.~\ref{eq-tbrlin-deconvolved}), solve for
+$\tau_{\mathrm{a}}(r)$ and obtain
+\begin{equation}
+  \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{model,br,lin}}}\,
+  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+\end{equation}
+leading with Eq.~(\ref{eq-tau-afo-sig-kap}) to
+\begin{equation}
+  \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\,
+  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+\end{equation}
+Integrating over $r$, ignoring geometric terms leads to
+\begin{equation}\label{eq-ringmass-from-obs}
+  \begin{split}
+    M_{\mathrm{d}} &\simeq 2\pi r_0\int_{r_{\mathrm{in}}}^{r_{\mathrm{out}}}\Sigma_{\mathrm{d}}(r)dr\\
+    &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\\
+    &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}
+  \end{split}
+\end{equation}
+
+\begin{figure}
+\centerline{\includegraphics[width=0.47\textwidth]{kappas_alma_band_6.pdf}}
+\caption{\label{fig-opacities}The dust opacity at $\lambda=0.13\,\mathrm{cm}$
+  as a function of the grain radius $a$. Plotted are the absorption opacity of
+  \citet{2010A&A...512A..15R}, the corresponding scattering opacity, as well as
+  the analytic absorption opacity of \citet{1997MNRAS.291..121I}.}
+\end{figure}
+
+As the final step we have to make an estimate of the absorption opacity
+$\kappa_{\mathrm{a}}$. This is the biggest uncertainty because it is not well
+understood what the composition of the dust is, what the grain sizes are and
+what the effect of the geometry is on the opacity. We will use two estimates: on
+the one hand we will employ the opacity used by \citet{2010A&A...512A..15R},
+which is a mixture of astronomical silicate and amorphous carbon (see
+Fig.~\ref{fig-opacities}). On the other hand we will use the dummy opacity model
+of \citet{1997MNRAS.291..121I}. These two opacities serve as two extremes that
+bracket the likely true value. It is known that at long wavelengths the
+opacities computed with the Mie procedure from laboratory measured optical
+constants are likely to be underestimates by up to a factor of 10
+\citep{2017A&A...600A.123D}. The more realistic opacities of
+\citep{2017A&A...600A.123D} are in between our two bracketing opacities.  For
+the mass estimate we use the opacity at $\lambda=0.13\,\mathrm{cm}$ for small
+grains ($a=1\mu\mathrm{m}$). For the Ricci opacity this is
+$\kappa_{\mathrm{a,ri}}=1.70\,\mathrm{cm}^2/\mathrm{g}$, while for the Ivezic
+opacity this is $\kappa_{\mathrm{a,ri}}=31.3\,\mathrm{cm}^2/\mathrm{g}$. The
+masses estimated from Eq.~(\ref{eq-ringmass-from-obs}) using these two
+bracketing opacities are listed in Table \ref{fig-obs-gaussfits}. Note that the
+opacity of \citet{2010A&A...512A..15R} has a peak around a grain size of 2
+$\mathrm{mm}$. That peak value happens to be very close to the Ivezic value of
+$31.3\,\mathrm{cm}^2/\mathrm{g}$. We can thus also regard the Ivezic mass
+estimate to be the mass if all grains have a radius of 2 mm.
 

### 2018-06-26 projects/huang/rings/continuumanalysis3.py
updated ring fitting code

@@ -13,7 +13,7 @@ import copy
 import matplotlib as mpl
 import matplotlib.colors as col
 import yaml
-from skimage import transform
+from skimage import transform, measure
 from scipy import ndimage
 
 
@@ -78,7 +78,11 @@ class Continuum:
         B_nu = Ivals*1.e-26*1/beamsize*(206264.806**2)
         return h*self.freq/(k*np.log(2*h*self.freq**3/c**2*1/B_nu+1.))
 
-    def azunwrap(self,radialbins, tbins = -179.+2*np.arange(180), yaxis = 'intensity'): 
+    def azunwrap(self,radialbins, tbins = -179.+2*np.arange(180), yaxis = 'intensity', theta_exclusion = np.array([])): 
+        """
+        Parameters
+        ==========
+        """
         rbins = radialbins
         rwidth = (rbins[1]-rbins[0])
         twidth = np.abs((tbins[1]-tbins[0]))
@@ -89,10 +93,16 @@ class Continuum:
         rad_prof_scatter = np.zeros_like(radial_profile)
 
         for i in range(len(rbins)):
-            annulus_intensities = self.image[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))]
             annulus_thetas = self.theta[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))]
+            if len(theta_exclusion)==2:
+                annulus_intensities = self.image[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))&(self.theta<=theta_exclusion[0] & (self.theta>=theta_exclusion[1]))]
+            else:
+                annulus_intensities = self.image[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))]
+
             radial_profile[i] = np.average(annulus_intensities)
             rad_prof_scatter[i] = np.std(annulus_intensities)
+            if len(theta_exclusion)==2:
+                annulus_intensities = self.image[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))]
             for j in range(len(tbins)):
                 wedge = annulus_intensities[(annulus_thetas>=(tbins[j]-0.5*twidth)) & (annulus_thetas<(tbins[j]+0.5*twidth))]
[…]

### 2018-06-27 projects/andrews/deredden.py
bunch of shit for intro paper

@@ -1,0 +1,30 @@
+import numpy as np
+from scipy.interpolate import interp1d
+
+def deredden(wl, Av, thresh=0.0):
+
+    # load composite extinction curves (Alambda / AK)
+    ext = np.loadtxt('ext_curves.dat')
+    awl = ext[:,0]
+    A1  = ext[:,1]
+    A2  = ext[:,2]
+    A3  = ext[:,3]
+
+    # set thresholds for different extinction curve regimes
+    Av_lo = thresh
+    Av_hi = 7.75	# McClure 2009 threshold: AK = 1.0
+    Av_me = 2.325	# McClure 2009 threshold: AK = 0.3
+    if (Av_lo >= Av_me): Av_lo = 0.
+    
+    # define AV / AK factors based on regime
+    if (Av >= Av_hi): AvAk, AA = 7.75, A3
+    if ((Av >= Av_me) & (Av < Av_hi)): AvAk, AA = 7.75, A2
+    if ((Av >= Av_lo) & (Av < Av_me)): AvAk, AA = 7.75, A2
+    if (Av < Av_lo): AvAk, AA = 9.03, A1
+    AK_AV = 1./AvAk
+
+    # interpolate extinction curve onto input wavelength grid
+    aint = interp1d(awl, AA, fill_value='extrapolate')
+    Alambda = Av * AK_AV * aint(wl)
+
+    return 10.**(0.4*Alambda)

### 2018-06-27 projects/andrews/hrd_inference.py

@@ -1,0 +1,169 @@
+import numpy as np
+import os
+import sys
+import time
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from astropy.io import ascii
+from scipy.interpolate import interp1d
+
+ndim = 6
+
+# plot layout
+xinch = 7.5
+yinch = 6.5
+dhspace, dwspace = 0.25, 0.25
+lbound, rbound, bbound, tbound = 0.08, 0.92, 0.08, 0.99
+fig = plt.figure(figsize=(xinch, yinch))
+gs = gridspec.GridSpec(2, 2)
+
+# plot ranges for each parameter
+Llims = [-1.0, 1.5]     # log Lstar / Lsun
+Tlims = [3.75, 3.5]      # log Teff / K
+Mlims = [-0.5, 0.4]     # log Mstar / Msun
+Alims = [5., 7.5]	
+
+# HRD
+ntest = 15000
+data_logT = np.random.normal(3.97, 0.03, ntest)
+data_logL = np.random.normal(1.23, 0.30, ntest)
+
+
+# HRD
+ax = fig.add_subplot(gs[0,0])
+
+
+# overplot models
+top_dir = '/pool/asha0/SCIENCE/Lupus_sizes/analysis/'
+iso_dir = top_dir+'MIST_isochrones/MIST_v1.1_vvcrit0.4_full_isos/'
+iso_file = iso_dir+'MIST_v1.1_feh_p0.00_afe_p0.0_vvcrit0.4_full.iso'
+iso_dat = ascii.read(iso_file)
+iso_AGE = iso_dat['col2']
+iso_M = iso_dat['col4']
+iso_L = iso_dat['col9']
+iso_T = iso_dat['col14']
+ax.plot(iso_T[iso_AGE == 5.6], iso_L[iso_AGE == 5.6], 'r')
+
+mt_dir = top_dir+'../SP/ScottiePippen/data/MIST/'
+mt_file = mt_dir+'MIST_v1.0_feh_p0.00_afe_p0.0_vvcrit0.4_EEPS/00070M.track.eep'
+mt_dat = ascii.read(mt_file)
+mt_AGE = np.log10(mt_dat['col1'])
+mt_M = mt_dat['col2']
+mt_L = mt_dat['col7']
+mt_T = mt_dat['col12']
[…]

### 2018-06-27 projects/andrews/specgrid.py

@@ -1,0 +1,26 @@
+import numpy as np
+import os
+import sys
+
+# set paths
+dir_nextgen = '/pool/asha1/HOLD/nextgen/'
+
+# make an appropriate spectral grid
+tgrid = np.loadtxt(dir_nextgen+'TEMP_GRID.dat')
+ggrid = np.loadtxt(dir_nextgen+'GRAV_GRID.dat')
+nwl = 1221
+tname = np.array([str(np.int(np.floor(i/100))) for i in tgrid])
+for i in range(len(tgrid)):
+    if (tgrid[i] < 10000.): tname[i] = '0'+tname[i]
+gname = np.array([str(i) for i in ggrid])
+spec = np.zeros((len(tgrid), len(ggrid), nwl))
+
+for i in range(len(tgrid)):
+    for j in range(len(ggrid)):
+        infile = dir_nextgen+'Fnu/NG_'+tname[i]+'_'+gname[j]+'_00.dat'
+        wl, cf = np.loadtxt(infile, usecols=(0, 2), unpack=True)
+        spec[i, j, :] = cf
+
+np.savez('nextgen_grid.npz', wl=wl, spec=spec, tgrid=tgrid, ggrid=ggrid)     
+    
+

### 2018-06-27 projects/andrews/starspectrum.py

@@ -1,0 +1,39 @@
+import numpy as np
+import sys
+from scipy.interpolate import interp1d
+
+def starspectrum(teff, lstar, dpc=1., type='Nextgen', grid=False, Jy=True):
+
+    # unit conversions
+    lstar = 3.826e33 * lstar
+    dcm = dpc * 3.0857e18
+
+    # HARD-CODE
+    logg = 4.0
+
+    # calculate other physical parameters
+    rstar = np.sqrt(lstar / (4.*np.pi*5.67051e-5*teff**4.))
+
+    # load spectrum grid
+    sgrid = np.load('nextgen_grid.npz')
+    wl, spec = sgrid['wl'], sgrid['spec']
+    tgrid, ggrid = sgrid['tgrid'], sgrid['ggrid']
+
+    # interpolate in Teff
+    ispec = np.zeros((len(ggrid), len(wl)))
+    for i in range(len(wl)):
+        for j in range(len(ggrid)):
+            tint = interp1d(tgrid, spec[:,j,i])
+            ispec[j, i] = tint(teff)
+
+    # interpolate in log g
+    fspec = np.zeros(len(wl))
+    for i in range(len(wl)):
+        gint = interp1d(ggrid, ispec[:,i])
+        fspec[i] = gint(logg)
+ 
+    # scale model spectrum to appropriate luminosity
+    Fnu = fspec * (rstar / dcm)**2
+    if (Jy == True): Fnu = 1e23 * Fnu
+
+    return 1e-3*wl, Fnu 

### 2018-07-25 projects/andrews/SEDgallery.py
figure generation code for intro paper

@@ -6,24 +6,25 @@ import matplotlib.gridspec as gridspec
 from astropy.io import ascii
 from starspectrum import starspectrum
 from deredden import deredden
+plt.rc('font', size=9)
 
-show_median = True
+show_median = False
 
 names = ['HTLup', 'GWLup', 'IMLup', 'HD142666', 'RULup',
          'HD143006', 'Sz129', 'MYLup', 'Sz114', 'AS205',
          'SR4', 'Elias20', 'DoAr25', 'Elias24', 'Elias27',
          'DoAr33', 'WSB52', 'WaOph6', 'AS209', 'HD163296']
 
-labels = [r'${\rm HT \;\: Lup}$', r'${\rm GW \;\: Lup}$', 
-          r'${\rm IM \;\: Lup}$', r'${\rm HD \;\: 142666}$', 
-          r'${\rm RU \;\: Lup}$', r'${\rm HD \;\: 143006}$', 
-          r'${\rm Sz \;\: 129}$', r'${\rm MY \;\: Lup}$', 
-          r'${\rm V908 \;\: Sco}$', r'${\rm AS \;\: 205}$',
-          r'${\rm SR \;\: 4}$', r'${\rm Elias \;\: 20}$', 
-          r'${\rm DoAr \;\: 25}$', r'${\rm Elias \;\: 24}$', 
-          r'${\rm Elias \;\: 27}$', r'${\rm DoAr \;\: 33}$', 
-          r'${\rm WSB \;\: 52}$', r'${\rm WaOph \;\: 6}$', 
-          r'${\rm AS \;\: 209}$', r'${\rm HD \;\: 163296}$']
+labels = ['HT Lup', 'GW Lup', 
+          'IM Lup', 'HD 142666', 
+          'RU Lup', 'HD 143006', 
+          'Sz 129', 'MY Lup', 
+          'Sz 114', 'AS 205',
+          'SR 4', 'Elias 20', 
+          'DoAr 25', 'Elias 24', 
+          'Elias 27', 'DoAr 33', 
+          'WSB 52', 'WaOph 6', 
+          'AS 209', 'HD 163296']
 
 dpc = np.array([154., 155., 158., 148., 159.,
                 165., 161., 156., 162., 128., 
@@ -48,7 +49,7 @@ Av = np.array([ 1.0,  0.5,  0.9,  0.5,  0.0,
 xlims = [0.2, 5000.]
[…]

### 2018-07-25 projects/andrews/casa_getfluxes.py

@@ -1,0 +1,39 @@
+import numpy as np
+import os
+import sys
+import time
+from datetime import date, timedelta
+
+cals = ['J1427-4206', 'J1517-2422', 'J1617-5848', 'J1733-1304']
+
+# set up dates and indexing
+d1, d2 = date(2017, 5, 1), date(2017, 12, 31) 
+delta = d2 - d1    
+
+# loop
+for j in range(len(cals)):
+
+    os.system('rm -rf '+cals[j]+'.fluxes.dat')
+
+    for i in range(delta.days + 1):
+
+        # compute a date
+        d = d1 + timedelta(i)
+
+        # measure the 240 GHz flux density for that date and calibrator
+        fetch = au.getALMAFlux(cals[0], frequency='238.74GHz', 
+                               date=d.strftime("20%y/%m/%d"))
+        Fnu, eFnu  = fetch['fluxDensity'], fetch['fluxDensityUncertainty']
+
+        # prepare for output
+        if (i < 10): ix = str(i) + '    '
+        if ((i >= 10) & (i < 100)): ix = str(i) + '   '
+        if (i >= 100): ix = str(i) + '  '
+        fstr = '%7.4f  %7.4f' % (Fnu, eFnu)
+
+        # write to file
+        f = open(cals[j]+'.fluxes.dat', 'a')
+        f.write(ix + str(d.year) + '  ' + str(d.month) + '  ' + str(d.day) + 
+                '  ' + fstr)
+        f.write('\n')
+        f.close()

### 2018-07-25 projects/andrews/flux_rescale_demo.py

@@ -1,0 +1,92 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import os
+import sys
+from deproject_vis import deproject_vis
+
+# plot setup
+plt.rc('font', size=8)
+fig = plt.figure(figsize=(3.5, 4.2))
+gs  = gridspec.GridSpec(2, 1)
+ax0 = fig.add_subplot(gs[0,0])
+ax1 = fig.add_subplot(gs[1,0])
+
+# deprojection parameters (from JH script)
+PA   = 41.5
+incl = 38.5
+offx = 0
+offy = 0
+
+# datafiles
+fle = ['SB1_initcont_exec0', 'SB1_initcont_exec1', 
+       'LB1_initcont_exec0', 'LB1_initcont_exec1']
+
+# load data and create visibility profiles
+all_rho = np.zeros((100, len(fle)))
+all_vre = np.zeros((100, len(fle)))
+all_sig = np.zeros((100, len(fle)))
+for i in range(len(fle)):
+    inpf = np.load('GWLup_'+fle[i]+'_shift.vis.npz')
+    u, v = inpf['u'], inpf['v']
+    vis  = inpf['Vis']
+    wgt  = inpf['Wgt']
+    uvbins = 10.+10.*np.arange(100)
+    vp = deproject_vis([u, v, vis, wgt], bins=uvbins, incl=incl, PA=PA,
+                       offx=offx, offy=offy)
+    vp_rho, vp_vis, vp_sig = vp
+    all_rho[:len(vp_rho),i] = vp_rho
+    all_vre[:len(vp_rho),i] = vp_vis.real
+    all_sig[:len(vp_rho),i] = vp_sig.real
+
+# plot the misaligned visibility profiles
+ax0.set_xlim([180, 520])
+ax0.set_ylim([0, 34])
+col = ['k', 'g', 'r', 'b']
+for i in range(len(fle)):
+    vp_rho = 1e-3*all_rho[:,i]
+    vp_vre = 1e3*all_vre[:,i]
+    vp_sig = 1e3*all_sig[:,i]
+    ax0.errorbar(vp_rho[vp_rho > 0], vp_vre[vp_rho > 0], 
+                 yerr=vp_sig[vp_rho > 0], fmt='o'+col[i], alpha=0.5)
+
+
+ax1.set_xlim([180, 520])
+ax1.set_ylim([0, 34])
+col = ['k', 'g', 'r', 'b']
+for i in range(len(fle)):
[…]

### 2018-07-25 projects/andrews/psf_demo.py

@@ -1,0 +1,201 @@
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from matplotlib.colorbar import Colorbar
+from matplotlib import mlab, cm
+from astropy.visualization import (AsinhStretch, LogStretch, LinearStretch, ImageNormalize)
+import sys
+
+plt.rc('font', size=9)
+
+# beam plotting function
+def mk_beam(hdr, npts=200):
+    semimajor = 0.5 * 3600. * hdr['BMAJ']
+    semiminor = 0.5 * 3600. * hdr['BMIN']
+    ang = np.radians(90.-hdr['BPA'])
+    phi = 2.*np.pi*np.arange(npts)/(npts-1.)
+    xp = semimajor*np.cos(phi)*np.cos(ang)-semiminor*np.sin(phi)*np.sin(ang)
+    yp = semimajor*np.cos(phi)*np.sin(ang)+semiminor*np.sin(phi)*np.cos(ang)
+    return [xp, yp]
+
+# generalities
+pdir = '/pool/asha0/SCIENCE/p484/sa_work/RULup/'
+cc = 2.9979e10
+kk = 1.381e-16
+
+# image sizes
+xlims = [0.5, -0.5]
+ylims = [-0.5, 0.5]
+aoff, doff = 3600.*-4.9074359961499946e-06, 3600.*2.3061111109563985e-5 
+
+# color scale 
+cm = 'inferno'
+vmin_c = 0.
+vmax_c = 70.
+norm_img = ImageNormalize(vmin=vmin_c, vmax=vmax_c, stretch=AsinhStretch())
+norm_psf = ImageNormalize(vmin=-0.05, vmax=1., stretch=AsinhStretch())
+
+# image lists
+sim_list = ['demo_SB', 'demo_LB1', 'demo_LB2', 
+            'demo_ALL_rob05', 'demo_ALL_rob00']
+
+# set up plot
+fig = plt.figure(figsize=(7.5, 4.65))
+gs  = gridspec.GridSpec(2, 6, height_ratios=(1, 1), 
+                              width_ratios=(1, 1, 1, 1, 1, 0.06))
+gs.update(bottom=0.075, top=0.61, wspace=0.05, hspace=0.05)
+
+# loop through images
+for i in range(len(sim_list)):
+
+    # grab image
[…]

### 2018-07-25 projects/andrews/selfcal_demo_combined.py

@@ -1,105 +1,0 @@
-import numpy as np
-from astropy.io import fits
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from matplotlib.colorbar import Colorbar
-from matplotlib import mlab, cm
-from astropy.visualization import (AsinhStretch, PowerStretch, LogStretch, ImageNormalize)
-import sys
-
-# beam plotting function
-def mk_beam(hdr, npts=200):
-    semimajor = 0.5 * 3600. * hdr['BMAJ']
-    semiminor = 0.5 * 3600. * hdr['BMIN']
-    ang = np.radians(90.-hdr['BPA'])
-    phi = 2.*np.pi*np.arange(npts)/(npts-1.)
-    xp = semimajor*np.cos(phi)*np.cos(ang)-semiminor*np.sin(phi)*np.sin(ang)
-    yp = semimajor*np.cos(phi)*np.sin(ang)+semiminor*np.sin(phi)*np.cos(ang)
-    return [xp, yp]
-
-# generalities
-pdir = '/pool/asha0/SCIENCE/p484/sa_work/RULup/'
-
-# image sizes
-xlims = [4., -4.]
-ylims = [-4., 4.]
-aoff, doff = 3600.*-4.9074359961499946e-06, 3600.*2.3061111109563985e-5 
-
-# color scale for compact observations
-cm = 'inferno'
-vmin_c = -0.025
-vmax_c = 5.
-norm = ImageNormalize(vmin=vmin_c, vmax=vmax_c, stretch=LogStretch())
-
-# image lists
-sim_list = ['combined_contp0', 'combined_contp2', 'combined_contp5', 'combined_contap']
-notes = [r'${\rm initial}$'+' '+r'${\rm (+}$'+' '+r'${\rm extended)}$',
-         r'${\rm phase}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (900,}$'+' '+r'${\rm 360s)}$',
-         r'${\rm phase}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (180,}$'+' '+r'${\rm 60,}$'+' '+r'${\rm 30s)}$',
-         r'${\rm amp}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (scan)}$',
-         ' ']
-
-
-# set up plot
-fig = plt.figure(figsize=(7.5, 1.9))
[…]

### 2018-07-25 projects/andrews/selfcal_demo_compact.py

@@ -1,105 +1,0 @@
-import numpy as np
-from astropy.io import fits
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-from matplotlib.colorbar import Colorbar
-from matplotlib import mlab, cm
-from astropy.visualization import (AsinhStretch, PowerStretch, LogStretch, ImageNormalize)
-import sys
-
-# beam plotting function
-def mk_beam(hdr, npts=200):
-    semimajor = 0.5 * 3600. * hdr['BMAJ']
-    semiminor = 0.5 * 3600. * hdr['BMIN']
-    ang = np.radians(90.-hdr['BPA'])
-    phi = 2.*np.pi*np.arange(npts)/(npts-1.)
-    xp = semimajor*np.cos(phi)*np.cos(ang)-semiminor*np.sin(phi)*np.sin(ang)
-    yp = semimajor*np.cos(phi)*np.sin(ang)+semiminor*np.sin(phi)*np.cos(ang)
-    return [xp, yp]
-
-# generalities
-pdir = '/pool/asha0/SCIENCE/p484/sa_work/RULup/'
-
-# image sizes
-xlims = [4., -4.]
-ylims = [-4., 4.]
-aoff, doff = 3600.*-4.9074359961499946e-06, 3600.*2.3061111109563985e-5 
-
-# color scale for compact observations
-cm = 'inferno'
-vmin_c = -0.065
-vmax_c = 30.
-norm = ImageNormalize(vmin=vmin_c, vmax=vmax_c, stretch=LogStretch())
-
-# image lists
-sim_list = ['SB_contp0', 'SB_contp1', 'SB_contp2', 'SB_contap']
-notes = [r'${\rm initial}$'+' '+r'${\rm (compact}$'+' '+r'${\rm only)}$',
-         r'${\rm phase}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (30s)}$',
-         r'${\rm phase}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (18s)}$',
-         r'${\rm amp}$'+' '+r'${\rm self}$'+'-'+r'${\rm cal}$'+' '+r'${\rm (scan)}$',
-         ' ']
-
-
-# set up plot
-fig = plt.figure(figsize=(7.5, 1.9))
[…]

### 2018-07-25 projects/andrews/astrom_demo.py
ht lup astrometry demo

@@ -1,0 +1,160 @@
+import numpy as np
+from astropy.io import fits
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+from matplotlib.colorbar import Colorbar
+from matplotlib import mlab, cm
+from astropy.visualization import (AsinhStretch, LinearStretch, ImageNormalize)
+import sys
+
+plt.rc('font', size=9)
+
+# beam plotting function
+def mk_beam(hdr, npts=200):
+    semimajor = 0.5 * 3600. * hdr['BMAJ']
+    semiminor = 0.5 * 3600. * hdr['BMIN']
+    ang = np.radians(90.-hdr['BPA'])
+    phi = 2.*np.pi*np.arange(npts)/(npts-1.)
+    xp = semimajor*np.cos(phi)*np.cos(ang)-semiminor*np.sin(phi)*np.sin(ang)
+    yp = semimajor*np.cos(phi)*np.sin(ang)+semiminor*np.sin(phi)*np.cos(ang)
+    return [xp, yp]
+
+# image sizes
+xlims = [0.5, -0.5]
+ylims = [-0.5, 0.5]
+aoff, doff = -0.045, -0.13
+
+# generalities
+cc = 2.9979e10
+kk = 1.381e-16
+
+# color scale for compact observations
+cm = 'inferno'
+norm = ImageNormalize(vmin=0, vmax=35, stretch=AsinhStretch())
+normc = ImageNormalize(vmin=0, vmax=5, stretch=AsinhStretch())
+
+# image lists
+sim_list = ['HTLup_SB_initcont_exec0', 'HTLup_SB_initcont_exec1', 
+            'HTLup_LB_initcont_exec0', 'HTLup_LB_initcont_exec1']
+notes = ['SB0', 'SB1', 'LB0', 'LB1']
+
+# set up plot
+fig = plt.figure(figsize=(7.5, 3.7))
+gs  = gridspec.GridSpec(2, 5, width_ratios=(1, 1, 1, 1, 0.06), 
+                              height_ratios=(1, 1))
+                              
+
+# loop through images
+for i in range(len(sim_list)):
+
+    # grab image
+    hdulist = fits.open(sim_list[i]+'.fits')
+    Inu = np.squeeze(hdulist[0].data)
+    hdr = hdulist[0].header
+    if (i == 0): ra0, dec0 = hdr['CRVAL1'], hdr['CRVAL2']
[…]

### 2018-07-27 projects/dullemond/paper_dusttraps/ms.tex
Added updated draft of dust trapping paper (still lots to do).

@@
 \abstract{A large fraction of the protoplanetary disks observed with the ALMA
-  Large Programme {\bf XXXX} display multiple well-defined and nearly perfectly
+  Large Programme DSHARP display multiple well-defined and nearly perfectly
   circular rings in the continuum, in many cases with substantial peak-to-valley
@@
 
-The data from the ALMA Large Programme 2016.1.00484.L offers an exciting new
+The data from the ALMA Large Programme DSHARP ({\bf REFERENCES}) offers an exciting new
 opportunity to put this concept to the test, and to put constraints on the
@@
 %
-In this paper we focus on those sources of the ALMA Large Programme
+In this paper we focus on those sources of the DSHARP Programme
 that show high-contrast and radially thin rings that are separated by deep
@@
 \hline
-AS 209     & 121              & 0.912        &  0.631 \\
-Elias 24   & 136              & 0.776        &  6.03  \\
-HD 163296  & 101              & 2.14         &  25.7  \\
-GW Lup     & 155              & 0.457        &  0.331 \\
-HD 143006  & 166              & 0.813        &  0.871 \\
+% AS 209     & 121              & 0.912        &  0.631 \\ % Data from 2018.05.13 Table
+AS 209     & 121              & 0.832        &  1.413 \\ % 2018.07.02
+Elias 24   & 136              & 0.776        &  6.03  \\ % 2018.07.02
+% HD 163296  & 101              & 2.14         &  25.7  \\ % Data from 2018.05.13 Table
+HD 163296  & 101              & 2.04         &  17.0  \\ % 2018.07.02
+GW Lup     & 155              & 0.457        &  0.331 \\ % 2018.07.02
+% HD 143006  & 166              & 0.813        &  0.871 \\ % Data estimated by Kees May/June 2018
+HD 143006  & 165              & 1.78        &  3.80 \\  % 2018.07.02 ** Big change **
 \hline
@@
 \end{center}
-\caption{\label{tab-stellar-params}The stellar parameters for the
+\caption{\label{tab-stellar-params}The stellar parameters assumed for the
   stars studied in this paper. Distance is in parsec and mass and luminosity
@@
 The high-contrast rings of these sources provide ``clean laboratories''
-for testing the theory of dust trapping, both in a ring-by-ring manner, as well
-as by comparing to global dust drift models. Fig.\ \ref{fig-obs-profiles}
+for testing the theory of dust trapping in a ring-by-ring manner.
+Fig.\ \ref{fig-obs-profiles}
 shows the radial profile (deprojected for
@@
 brightness is expressed as linear brightness temperature
-$T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
+$T_{\mathrm{br,lin}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
 intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
 frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
-defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
+defined such that $B_\nu(T_{\mathrm{br,full}})=I_\nu$, with $B_\nu(T)$ the
 Planck function, can have very different values, in particular at low
-brightness. As a result $T_{\mathrm{bright,full}}(r)$ tends to vastly
+brightness. As a result $T_{\mathrm{br,full}}(r)$ tends to vastly
 underestimate the contrast between the rings and the inter-ring
@@
 
-The eight rings which we study in this paper (see annotations in
-Fig.~\ref{fig-obs-images}) are fitted with a Gaussian profile, shown as inverse
-parabolas in Fig.\ \ref{fig-obs-profiles}. The fitting procedure was done in the
-image plane, in order to make it possible to fit to a limited radial range,
-i.e.\ selecting a single ring. Details of this fitting procedure, and tests
-using the ALMA simulator, are described in Appendix {\bf XXXXX}. In
-Fig.~\ref{fig-obs-gaussfits} these gaussian fits are shown in close-up. The
-parameters of the best-fitting Gaussians are listed in Table
-\ref{tab-gauss-params}.
+As is shown in Appendix \ref{sec-steady-state-analytic-trap-model}, for a
+Gaussian pressure bump the solution to the radial dust mixing and drift problem
+is, to first approximation, also a Gaussian, with a width smaller than, or equal
+to, that of the gas pressure bump. Our analysis of the eight rings of this paper
+therefore naturally starts with the fitting of the observed radial intensity
+profiles with a Gaussian function. For some {\bf [CHECK]} of the rings of this
+paper such Gaussian fits were done in the context of fitting the entire radial
+intensity profile in the uv-plane to the ALMA data ({\bf Guzman et al.XXXX;
+  More??}). Our fits may deviate slightly from those {\bf [CHECK]}, because here
+we put emphasis on the shape of the peak of the rings, and deemphasize the
+wings.
+
+The fitting procedure is described in Appendix \ref{sec-gauss-fitting}.  The
+fitting was done in the image plane, in order to make it possible to fit to a
+limited radial range, i.e.\ selecting a single ring. The Gaussian fits are shown
+as inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
+Fig.~\ref{fig-obs-gaussfits} these Gaussian fits are shown in close-up. The
+parameters of the best fits are listed in Table \ref{tab-gauss-params}.
 
@@
 \begin{center}
-\begin{tabular}{|cc|cc|ccccc|ccccc|}
+\begin{tabular}{|cc|cc|ccccc|cccccc|}
 \hline
 \hline
-%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{model}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
-Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{model}}$ & $T_{\mathrm{model,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
-            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{d}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
+%Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
+%            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
+            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
 \hline
 % NOTE: This table is made automatically with python_make_tables/make_latex_table.py
-AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  12.9 &   8.2 &   0.7 &  44.4 &   2.41\\
-AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  10.1 &   5.6 &   0.5 & 101.9 &   5.54\\
-Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &  41.1 &   2.23\\
-HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  34.2 &  29.0 &   1.6 &  50.4 &   2.74\\
-HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  28.1 &  22.9 &   0.8 &  42.5 &   2.31\\
-GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &  36.3 &   1.97\\
-HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  71.76 &   5.06 &   4.42 &  18.8 &  13.8 &   1.7 &  16.7 &   0.91\\
-HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 111.83 &   7.88 &   7.49 &  15.0 &  10.1 &   1.7 &  43.5 &   2.36\\
+% Note: New table made with new stellar parameters 2018.07.02
+AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  15.8 &  10.9 &   0.6 &   0.41 &  33.3 &   1.81\\
+AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  12.4 &   7.7 &   0.4 &   0.48 &  74.3 &   4.04\\
+Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.36 &  41.1 &   2.23\\
+HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  30.8 &  25.6 &   1.7 &   0.40 &  56.9 &   3.09\\
+HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  25.3 &  20.2 &   0.8 &   0.28 &  48.3 &   2.62\\
+GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &   0.30 &  36.3 &   1.97\\
+HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  72.20 &   5.06 &   4.42 &  27.2 &  22.0 &   2.1 &   0.16 &  10.5 &   0.57\\
+HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 112.51 &   7.88 &   7.49 &  21.6 &  16.6 &   2.0 &   0.16 &  26.5 &   1.44\\
 \hline
@@
 \caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
-  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values $r_{\mathrm{in}}$
-  and $r_{\mathrm{out}}$ are the inner and outer boundary of the fitting domain. $A$ is the
-  peak brightness temperature of the ring, $r_0$ is its radius in $\mathrm{au}$,
+  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values
+  $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$ are the inner and outer boundary of
+  the fitting domain. $A$ is the peak linear brightness temperature of the
+  best-fit Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$,
   $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
@@
   dust emission profile, also expressed as standard deviation in units of
-  $\mathrm{au}$. The last five columns are derived from a simple disk model
-  assumption: $T_{\mathrm{model}}$ is the midplane temperature computed from
+  $\mathrm{au}$. The last six columns are derived from a simple disk model
+  assumption: $T_{\mathrm{d}}$ is the midplane temperature of the disk (we
+  assume gas and dust temperature to be equal) computed from
   Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
   $\varphi=0.02$. From that one can derive the corresponding linear brightness
-  temerature $T_{\mathrm{model,br,lin}}=B_\nu(T_{\mathrm{model}})c^2/2\nu^2$ with
-  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of
-  0.13 $\mathrm{cm}$ of ALMA band 6.
-  The third-to-last column expresses the deconvolved dust ring
-  width in units of the disk pressure scale height $h_p$ computed from
-  $T_{\mathrm{model}}$. The last two columns are the computed dust mass in the
-  ring in units of Earth masses, assuming the small grain absorption opacity
+  temerature $T_{\mathrm{d,br,lin}}=B_\nu(T_{\mathrm{d}})c^2/2\nu^2$ where
+  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of 0.13
+  $\mathrm{cm}$ of ALMA band 6.  The fourth-to-last column expresses the
+  deconvolved dust ring width in units of the disk pressure scale height $h_p$
+  computed from $T_{\mathrm{d}}$. The $\tau$ column gives the estimated optical
+  depth at the peak of the ring, calculated from
+  $\tau=-\ln(1-A/T_{\mathrm{d,br,lin}})$.  The last two columns are the computed
+  dust mass in the ring in units of Earth masses, assuming the small grain
+  absorption opacity
   $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
@@
   $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
-  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). The stellar parameter
-  are taken from Table \ref{tab-stellar-params}.}
+  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). In making these
+  estimates, the optically thin approximation was used. See Appendix
+  \ref{sec-computing-dust-mass} for details. An estimate of the
+  optical-depth-corrected dust mass can be made by multiplying the mass values
+  in the table by $\tau/(1-e^{-\tau})$.}
 \end{table*}
@@
 parsec). The width of the dust rings
-in units of the local pressure scale height $h_p$ (last column of Table
+in units of the local pressure scale height $h_p$ (see Table
 \ref{tab-gauss-params}) can only be roughly estimated, because we do not know
@@
 \begin{equation}\label{eq-disk-temperature-model}
-T_{\mathrm{model}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
+T_{\mathrm{d}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
 \end{equation}
@@
   1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
-follows from $h_p=\sqrt{k_BT_{\mathrm{model}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
+follows from $h_p=\sqrt{k_BT_{\mathrm{d}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
 Boltzmann constant, $m_p$ the proton mass, $G$ the gravitational constant and
-$mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
+$\mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
 taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
@@
 this value can easily be wrong by a factor of four or so. The fact that the
-pressure scale height goes as $T_{\mathrm{model}}^{1/2}\propto \varphi^{1/8}$
+pressure scale height goes as $T_{\mathrm{d}}^{1/2}\propto \varphi^{1/8}$
 means, however, that such an error only affects $h_p$ mildly.
@@
 %
-First of all, the Schmidt number $\mathrm{Sc}$ is merely a way to relate the
-turbulent viscosity with the turbulent mixing. If we do not strive to learn
-about the turbulent viscosity and instead are satisfied with learning only about
-the turbulent mixing, then we are only interested in the combination
-$\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For simplicity we set $\mathrm{Sc}=1$.
+Reasonable values of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and
+$w$ obey certain restrictions. First of all, the Schmidt number $\mathrm{Sc}$
+is merely a way to relate the turbulent viscosity with the turbulent mixing. If
+we do not strive to learn about the turbulent viscosity, and instead are
+satisfied with learning only about the turbulent mixing, then we are only
+interested in the combination $\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For
+simplicity we set $\mathrm{Sc}=1$.
 
@@
 
-The Stokes number $\mathrm{St}$ can be any value. But it is
-directly related to the grain size $a$ and the gas density
-$\rho_{\mathrm{g}}$. If we have observational constraints on $a$ and a good
-estimate of the gas density $\rho_{\mathrm{g}}$, then we can eliminate this
-uncertainty, and we are left with two unknown parameters
-($\alpha_{\mathrm{turb}}$ and $w$) for one measurement ($w_{\mathrm{d}}$).
-Unfortunately, while estimating $a$ from observations may be doable, it is far
-more difficult to estimate $\rho_{\mathrm{g}}$. Standard disk gas mass estimates
-are of limited use, as they are based on measuring the dust mass and multiplying
-it by the estimated gas-to-dust ratio. Since we are testing the hypothesis of
-dust trapping, we cannot assume a standard gas-to-dust ratio.
-
-One can, however, set an upper bound on $\rho_{\mathrm{g}}$ by demanding that
+The Stokes number $\mathrm{St}$ can be any value. But it is directly related to
+the grain size $a$ and the gas density $\rho_{\mathrm{g}}$, where the gas
+density is directly related to the gas surface density $\Sigma_{\mathrm{g}}$ via
+$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. If we have observational
+constraints on $a$ and a good estimate of the gas surface density
+$\Sigma_{\mathrm{g}}$, then we can eliminate this uncertainty, and we are left
+with two unknown parameters ($\alpha_{\mathrm{turb}}$ and $w$) for one
+measurement ($w_{\mathrm{d}}$).  Unfortunately, while estimating $a$ from
+observations may be doable, it is far more difficult to estimate
+$\Sigma_{\mathrm{g}}$. Standard disk gas mass estimates are of limited use, as
+they are based on measuring the dust mass and multiplying it by the estimated
+gas-to-dust ratio. Since we are testing the hypothesis of dust trapping, we
+cannot assume a standard gas-to-dust ratio.
+
+One can, however, set an upper bound on $\Sigma_{\mathrm{g}}$ by demanding that
 the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
@@
 $G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
-surface density. The gas surface density and midplane density are related via
-the pressure scale height $h_p=c_s/\Omega_K$ via the identity
-$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. Taking the disk midplane
+surface density. Taking the disk midplane
 temperature from Table \ref{tab-gauss-params}, which was calculated using
 Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
-$\rho_{\mathrm{g}}$ for all of the rings. The results are given in Table
+$\Sigma_{\mathrm{g}}$ for all of the rings. The results are given in Table
 \ref{tab-ring-model-limits}.
@@
 \begin{center}
-\begin{tabular}{|cc|cccccc|cc|}
+\begin{tabular}{|cc|ccccccc|ccc|}
 \hline
 \hline
-Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\rho_{\mathrm{g,min,iv}}$ & $\rho_{\mathrm{g,min,ri}}$ & $\rho_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $(\alpha/\mathrm{St})_{\mathrm{min}}$ & $(\alpha/\mathrm{St})_{\mathrm{max}}$\\
-           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & $\equiv\psi^{-2}_{\mathrm{max}}$ & $\equiv\psi^{-2}_{\mathrm{min}}$\\
+Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\Sigma_{\mathrm{g,min,iv}}$ & $\Sigma_{\mathrm{g,min,ri}}$ & $\Sigma_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ & $\alpha/\mathrm{St}$ & $\alpha/\mathrm{St}$ & $\alpha_{\mathrm{exmp}}$\\
+           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & (for $\Sigma_{\mathrm{g,max}}$)& (for $w_\mathrm{max}$) & (for $w_\mathrm{min}$) &   \\
 \hline
-AS 209     & 1 &  4.83 & 19.6 & 7.9e-17 & 1.4e-15 & 8.4e-14 & 11.1 & 3.2e-02 & 1.0e+00 \\
-AS 209     & 2 &  8.85 & 19.6 & 5.1e-17 & 9.3e-16 & 2.0e-14 &  4.0 & 5.0e-02 & 3.0e-01 \\
-Elias 24   & 1 &  7.26 & 17.1 & 3.6e-17 & 6.5e-16 & 6.4e-14 & 18.8 & 8.0e-02 & 7.0e-01 \\
-HD 163296  & 1 &  7.21$^{*}$ & 13.8 & 5.6e-17 & 1.0e-15 & 2.6e-13 & 48.7 & 3.8e-01 & -- \\
-HD 163296  & 2 &  7.27 & 13.8 & 2.5e-17 & 4.6e-16 & 8.1e-14 & 33.4 & 1.9e-01 & 1.4e+00 \\
-GW Lup     & 1 &  7.49 &  9.9 & 2.9e-17 & 5.4e-16 & 2.8e-14 &  9.8 & 2.1e-01 & 4.4e-01 \\
-HD 143006  & 1 &  4.42$^{*}$ & 10.1 & 7.8e-17 & 1.4e-15 & 4.4e-13 & 58.9 & 2.4e-01 & -- \\
-HD 143006  & 2 &  7.49$^{*}$ & 10.1 & 4.6e-17 & 8.5e-16 & 1.1e-13 & 25.1 & 1.2e+00 & -- \\
+% Note: New table made with new stellar parameters 2018.07.02
+AS 209     & 1 &  5.59 & 19.6 & 1.1e-02 & 2.0e-01 & 1.6e+01 & 15.7 & 3.9e-03 & 3.2e-02 & 6.0e-01 & 1.2e-04\\
+AS 209     & 2 & 10.25 & 19.6 & 1.2e-02 & 2.3e-01 & 6.9e+00 &  5.8 & 9.1e-03 & 5.0e-02 & 2.1e-01 & 4.5e-04\\
+Elias 24   & 1 &  7.26 & 17.1 & 9.7e-03 & 1.8e-01 & 1.8e+01 & 18.8 & 3.6e-03 & 8.0e-02 & 7.0e-01 & 2.9e-04\\
+HD 163296  & 1 &  7.21$^{*}$ & 13.8 & 1.1e-02 & 1.9e-01 & 4.1e+01 & 40.0 & 1.5e-03 & 3.8e-01 & -- & 5.8e-04\\
+HD 163296  & 2 &  7.07 & 13.8 & 7.8e-03 & 1.4e-01 & 2.0e+01 & 27.3 & 3.1e-03 & 1.9e-01 & 1.6e+00 & 5.9e-04\\
+GW Lup     & 1 &  7.49 &  9.9 & 8.3e-03 & 1.5e-01 & 7.8e+00 &  9.8 & 8.0e-03 & 2.1e-01 & 4.4e-01 & 1.7e-03\\
+HD 143006  & 1 &  4.42$^{*}$ & 10.1 & 4.7e-03 & 8.6e-02 & 7.5e+01 & 167.2 & 8.4e-04 & 2.4e-01 & -- & 2.0e-04\\
+HD 143006  & 2 &  7.49$^{*}$ & 10.1 & 4.8e-03 & 8.8e-02 & 3.4e+01 & 73.2 & 1.9e-03 & 1.2e+00 & -- & 2.3e-03\\
 \hline
@@
   to the nearest minimum (for Elias 24 and GW Lup). The two lower limits on the
-  gas density $\rho_{\mathrm{g}}$ were derived by demanding
+  gas surface density $\Sigma_{\mathrm{g}}$ were derived by demanding
   $\Sigma_{\mathrm{g}}\gtrsim \Sigma_{\mathrm{d}}$, where $\Sigma_{\mathrm{d}}$
   was computed in Appendix \ref{sec-computing-dust-mass} using two different
-  opacities: that of Ivezic et al. (leading to $\rho_{\mathrm{g,min,iv}}$) and
-  that of Ricci et al.\ (leading to $\rho_{\mathrm{g,min,ri}}$). The column
-  marked with $\rho_{\mathrm{g,max}}$ is the upper limit on the midplane gas
+  opacities: that of Ivezic et al. (leading to $\Sigma_{\mathrm{g,min,iv}}$) and
+  that of Ricci et al.\ (leading to $\Sigma_{\mathrm{g,min,ri}}$). The column
+  marked with $\Sigma_{\mathrm{g,max}}$ is the upper limit on the gas surface
   density derived from demanding that the gas disk is gravitationally
@@
   derived dust surface density together with the gas surface density remain
-  gravitationally stable. The column $(\alpha/\mathrm{St})_{\mathrm{min}}\equiv
-  \psi^{-2}_{\mathrm{max}}$ is a result of the analysis: the lower limit to
-  $\alpha_{\mathrm{turb}}/\mathrm{St}$ derived from the ratio of
-  $w_{\mathrm{max}}/w_{\mathrm{d}}$.}
+  gravitationally stable. The $\mathrm{St}_{(a=0.02\,\mathrm{cm})}$ column gives
+  an example value of the Stokes number: it is the value of $\mathrm{St}$ if the
+  grains have a radius of $0.02\,\mathrm{cm}$, for the case of
+  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$. The two columns for
+  $\alpha/\mathrm{St}$ are derived for the widest and narrowest
+  gas bump, respectively. The column $\alpha_{\mathrm{exmp}}$ is the value of
+  $\alpha_{\mathrm{turb}}$ for an example choice of parameters: $w=w_{\mathrm{max}}$,
+  $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$.}
 \end{table*}
@@
 $\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
-\Sigma_{\mathrm{d}}$, and using the pressure scale height $h_p$
-and the formula $\rho_{\mathrm{g,midplane}}=\Sigma_{\mathrm{g}}/\sqrt{2\pi}h_p$,
-we arrive at two lower limits of the midplane gas density, given in
+\Sigma_{\mathrm{d}}$ we arrive at two lower limits of the gas surface density, given in
 Table \ref{tab-ring-model-limits}. It is likely that even for much
-larger values of the midplane gas density the dust-gas mixture becomes
+larger values of the gas surface density the dust-gas mixture becomes
 unstable to the streaming instability and other types of instabilities.
@@
 in these rings, then this puts a substantially stronger (i.e.~larger)
-lower limit on the midplane gas density.
+lower limit on the gas surface density.
 
@@
 
-Let us first choose the widest allowed pressure bump (i.e.\ taking
-$w=w_{\mathrm{max}}$ from Table \ref{tab-ring-model-limits}). In that
-case the dust rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$,
-which implies that the dust trapping is operational. We can derive the
-values of $\psi$ directly from Eq.~(\ref{eq-wd-afo-w-psi}). And together
-with Eq.~(\ref{eq-psi-afo-alpha-sc-st}) we can then directly derive the
-following constraint on the unknown parameters $\alpha_{\mathrm{turb}}$
-and $\mathrm{St}$, 
-\begin{equation}
-\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}} \equiv \psi^2 = \left(\frac{w}{w_{\mathrm{d}}}\right)^2-1
-\end{equation}
-where, we recall, set $\mathrm{Sc}=1$. Given that this is the largest
-possible value of $w$, and therefore the largest possible value
-of $\psi$, this gives a lower limit to the ratio $\alpha_{\mathrm{turb}}/\mathrm{St}$:
-\begin{equation}
-\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \gtrsim \frac{1}{\psi^2_{\mathrm{max}}} \qquad\hbox{($\psi_{\mathrm{max}}$ from largest $w$)}
-\end{equation}
-The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
-the first quantitative results of the application of the dust trapping model to
-the data.
-
-We can also derive an upper limit to the ratio
-$\alpha_{\mathrm{turb}}/\mathrm{St}$, but only for the rings for which
+From an assumed value of $w$ and the measured value $w_{\mathrm{d}}$ we can
+directly compute the ratio $\alpha_{\mathrm{turb}}/\mathrm{St}$
+\begin{equation}\label{eq-alphaSt-equation}
+\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \equiv \psi^{-2} = \left[\left(\frac{w}{w_{\mathrm{d}}}\right)^2-1\right]^{-1}
+\end{equation}
+where we used Eqs.~(\ref{eq-wd-afo-w-psi}, \ref{eq-psi-afo-alpha-sc-st}), and
+set $\mathrm{Sc}=1$. We will consider two choices of $w$: the $w_{\mathrm{min}}$ and
+$w_{\mathrm{max}}$ from Table \ref{tab-ring-model-limits}.
+
+For the choice $w=w_{\mathrm{max}}$ (the widest possible pressure bump) the dust
+rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$, which implies
+that the dust trapping is operational, and Eq.~(\ref{eq-alphaSt-equation}) gives
+information about the turbulent strength. For the choice $w=w_{\mathrm{min}}$
+(the narrowest possible pressure bump) we can only use
+Eq.~(\ref{eq-alphaSt-equation}) for the rings for which
 $w_{\mathrm{d}}<h_p$. The reason is that for those rings with
@@
 the minimal pressure bump width is $w_{\mathrm{min}}=w_{\mathrm{d}}$, and the
-dust ring is as wide as the pressure bump, implying that dust trapping is weak.
-Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will keep
-$w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive an upper limit on
-$\alpha_{\mathrm{turb}}/\mathrm{St}$. For the rings with $w_{\mathrm{d}}<h_p$
-(and thus $w_{\mathrm{min}}=h_p$) we can write
+dust ring is as wide as the pressure bump, implying that dust trapping is weak
+or non-operational. Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will
+keep $w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive any value for
+$\alpha_{\mathrm{turb}}/\mathrm{St}$. But for other rings we can compute
+$\alpha_{\mathrm{turb}}/\mathrm{St}$. The resulting values for both choices of
+pressure bump width are given in Table \ref{tab-ring-model-limits}. They can be
+understood as the lower and upper limit on $\alpha_{\mathrm{turb}}/\mathrm{St}$.
+
+% If we will be able to measure the width $w$ of the gas pressure bump in an
+% independent way, for instance with molecular line observations, then instead of
+% the lower- and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$ we will be
+% able to directly measure $\alpha_{\mathrm{turb}}/\mathrm{St}$. However, reliable
+% measurements of the gas density are notoriously difficult. CO molecular lines
+% are usually emitted in the warm sub-surface layers of the disk, not near the
+% midplane (see e.g.~{\bf Weaver \& Isella companion paper as well as earlier
+%   literature}), and CO appears to be depleted compared to chemistry models
+% ({\bf XXXX}). Other gas tracers are harder to measure and also not reliable
+% indicators of the gas density. Therefore we will, at least for now, have to
+% make do with the lower and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$
+% from Table \ref{tab-ring-model-limits}.
+
+The next task is to convert from Stokes number $\mathrm{St}$ to grain size
+$a$. The Epstein regime is valid for the typical grains we are interested in
+(less than a meter in size), in which case $a$ and $\mathrm{St}$ are related
+by 
 \begin{equation}
-\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \lesssim \frac{1}{\psi^2_{\mathrm{min}}} \qquad\hbox{($\psi_{\mathrm{min}}$ from smallest $w$)}
-\end{equation}
-The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
-the second quantitative results of the application of the dust trapping model to
-the data.
-
-If we will be able to measure the width of the gas pressure bump in an
-independent way, for instance with molecular line observations, then instead of
-the lower- and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$ we will be
-able to directly measure $\alpha_{\mathrm{turb}}/\mathrm{St}$. However, reliable
-measurements of the gas density are notoriously difficult. CO molecular lines
-are usually emitted in the warm sub-surface layers of the disk, not near the
-midplane (see e.g.~{\bf Weaver \& Isella companion paper as well as earlier
-  literature}), and CO appears to be depleted compared to chemistry models
-({\bf XXXX}). Other gas tracers are harder to measure and also not reliable
-indicators of the gas density. Therefore we will, at least for now, have to
-make do with the lower and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$
-from Table \ref{tab-ring-model-limits}.
-
-The next task is to convert from Stokes number $\mathrm{St}$ to grain size
-$a$.
-
-{\bf [TO BE DONE]}
-
-{\bf Make an exclusion diagram.}
+\mathrm{St} = \frac{\pi}{2}\frac{\xi_{\mathrm{dust}}a}{\Sigma_{\mathrm{g}}}
+\end{equation}
+where $\Sigma_{\mathrm{g}}$ is the gas surface density and $\xi_{\mathrm{dust}}$
+is the material density of the dust grains. Given that grains are expected
+to be a mixture of silicate, amorphous carbon and water ice, a value of
+$\xi_{\mathrm{dust}}\simeq 2\,\mathrm{g}/\mathrm{cm}^3$ is reasonable.
+
+To get a feeling for the results, let us choose the grain size to be
+$a=0.02\,\mathrm{cm}$, which corresponds to $\lambda/2\pi$ for
+$\lambda=0.13\,\mathrm{cm}$ (the wavelength of ALMA band 6).  The corresponding
+Stokes numbers, for the most massive possible gas disk
+($\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$), are listed in Table
+\ref{tab-ring-model-limits}. This then allows us to convert the value of
+$\alpha_{\mathrm{turb}}/\mathrm{St}$ into a value of
+$\alpha_{\mathrm{turb}}$. For the case $w=w_{\mathrm{max}}$ this leads to values
+$\alpha_{\mathrm{turb}}= 10^{-4}\cdots\mathrm{few}\,\times 10^{-3}$, listed in
+Table \ref{tab-ring-model-limits} (column $\alpha_{\mathrm{exmp}}$).
+
+These low values of $\alpha_{\mathrm{turb}}$ are consistent with the low values
+or upper limits reported recently \citep{2016ApJ...816...25P,
+  2018ApJ...856..117F} {\bf [more?]}. However, it has to be kept in mind that
+the values of $\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ were derived
+for an example choice of parameters: : $w=w_{\mathrm{max}}$,
+$\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$. For a
+smaller value of $w$, a lower value of $\Sigma_{\mathrm{g}}$, or larger grains,
+the computed value of $\alpha_{\mathrm{turb}}$ will increase. So that it is hard
+to set a true upper limit on $\alpha_{\mathrm{turb}}$ from these observations.
+
+This seems to be in conflict with the results found by
+\citet{2016ApJ...816...25P} who, from measuring an upper limit on the vertical
+extent of the dust, imply an upper limit on $\alpha_{\mathrm{turb}}$ of $3\times
+10^{-3}$. {\bf [IMPORTANT: We have to find out why Pinte et al. find this
+    limit. Do they have additional constraints? Why can't they increase the grain
+    size indefinitely? What are the constraints Eric and Andrea find for the
+similar analysis of HD 163296?]}
+
+Can we derive a {\em lower} limit to $\alpha_{\mathrm{turb}}$? This depends on
+whether we have information about the grain size. At present we have only the
+high resolution data for band 6, so we do not yet have information about the
+radial profile of the spectral slope. But in several recent observations of the
+spectral index across ringed disks
+\citep{2015ApJ...808L...3A,2018ApJ...852..122H} one clearly sees that
+$\alpha_{\mathrm{spec}}$ varies across these rings, being closer to $2$ at the
+ring center and substantially larger between the rings. This makes sense in
+terms of the dust trapping scenario in which we expect larger grains to be
+trapped more efficiently (and thus dominate the peak of the ring) than smaller
+grains, because the smaller grains will be more subject to turbulent
+mixing. This scenario requires a grain size distribution, so that the width of
+the dust ring is smaller for the bigger grains and bigger for the smaller ones.
+We will discuss our model in the context of grain size distributions in Section
+\ref{sec-model-with-grain-size-distribution}. But assuming that future data will
+also show similar spectral slope changes for our sources, we can already
+conclude that the grains trapped in the rings must have sizes larger than
+$\lambda/2\pi=0.02\,\mathrm{cm}$. The measured (resolved) width of the dust ring
+$w_d$ then implies a lower limit to $\alpha_{\mathrm{turb}}$. The value
+$\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ is then, in fact, this lower
+limit. We will, however, have to wait until high resolution ALMA data in another
+band becomes available to verify the minimal grain size of $0.02\,\mathrm{cm}$.
+
+Should the low spectral slope at the peak of the rings be confirmed, then the
+lower limits on $\alpha_{\mathrm{turb}}$ from the $\alpha_{\mathrm{exmp}}$
+column in Table \ref{tab-ring-model-limits} paint a different picture from the
+low-turbulence picture that has emerged in recent times. The values of a few
+times $10^{-4}$ to a few times $10^{-3}$ are low, but they are only lower
+bounds. It would be somewhat unlikely if all rings have parameters at this
+extreme size of parameter space, and it is therefore likely that
+$\alpha_{\mathrm{turb}}$ is easily a factor of ten higher or so. This would
+imply that protoplanetary disks are substantially turbulent after all. But to
+confirm this, we need to dig a bit deeper by including size distributions
+(Section \ref{sec-model-with-grain-size-distribution}) and non-perfect
+(``leaky'') dust traps (Section {\bf XXXX}).
 
@@
 
-{\bf Coincidence of the slope changes.}
-
-{\bf Can the height measurement help? There is still the degeneracy. Why did
-Pinte et al. not have this degeneracy? How could they pin down the turbulence?}
+{\bf CONSIDER PRESSURE OFFSET!}
 
@@
 %
+\label{sec-model-with-grain-size-distribution}
 
@@
 {\bf [This section is temporarily hidden, because a lot still has to be
-done on this.]}
+done on this. Maybe it should be removed and put into a future paper.]}
 
@@
 
-\subsection{Gap depth estimates from 2-D hydrodynamic simulations}
-
-{\bf Dong \& Fung (2017)} presented a systematic study of the gap
-opening profiles for given planetary masses, and analyzed the
-rings seen in scattered light of 5 sources (HD 97048, TW Hydra,
-HD 169142, LkCa15 and RX J1615).
-
 \subsection{Condition for the streaming instability}
@@
 
-\subsection{Dust trapping in the horseshoe region and in resonances}
-The 1-D approach of the {\sf DISKLAB} models make it impossible to treat the
-trapping of particles in the horseshoe region. The low-$\alpha$ models of {\bf
-  Liu et al.~2018} show that this may be a significant effect. However, this
-tends to produce non-axisymmetric rings: there will clearly be a dearth of dust
-near the location of the planet, and a slight enhancement of dust near the L4
-and L5 Lagrange points. Given that the rings we see so far in the multi-ringed
-protoplanetary disks are angularly symmetric, this appears to be ruled out.
-That means that the 1-D approach is justified, but we do have to ensure that we
-are not in the regime where such non-axisymmetric horseshoe-trapping is
-important. Generally the horseshoe region tends to gradually deplete over time,
-but the time scale depends on the turbulent viscosity, with low-$\alpha$ disks
-depleting slower.
-
-Another issue is the potential for large particles to get trapped in a
-resonance of a planet instead of in a classical dust trap {\bf REFERENCE;
-the Dong double-gap is different, though}.
-This requires high Stokes number, generally $\mathrm{St}\gg 1$ ({\bf is
-  this true?}). Given that grain growth models tend to keep grains below
-this limit ({\bf see e.g.~Birnstiel et al.}), this is less likely to play
-a role.
-
-\subsection{Multiple gaps with a single planet}
-{\bf Dong, Li \& Li (2017)} and 
-{\bf Bae, Zhu \& Hartmann (2017)} show that for very low-$\alpha$ disks
-a planet can open multiple gaps in the disk rather than just a single one.
-The contrast of these rings and gaps in the gas can be large (factor of 2; see
-Bae et al.),
-and thus clearly constitute dust traps. Dong et al.\ show that these gaps
-can even be observable for very low mass planets.
-
-{\bf [TODO: We can do a simple model of dust trapping for their case.]}
-
-{\bf Fung \& Chiang (2017)} have shown that low mass planets in nearly inviscid
-disks can avoid migration and can be responsible for disk accretion in spite
-of the low viscosity.
+{\bf [Add discussion of the plateau of Elias 24]}
 
@@
   to be $2\,\mathrm{g}/\mathrm{cm}^3$ consistent with slightly porous silicate
-  dust aggregates. The standard-deviation width of the gas pressure bump was set
+  dust aggregates.
+  {\bf [CHECK IF THIS IS CONSISTENT WITH THE OPACITY ESTIMATES, IN PARTICULAR
+  THE RICCI OPACITY.]}
+  The standard-deviation width of the gas pressure bump was set
   to $w=9.7\,\mathrm{au}$ which equals two times the gas pressure scale height.}
@@
 
-\subsection{Model fits to the ALMA data}
-\label{sec-anmodel-fits-to-data}
-%
-We now apply this model to the isolated rings in our ALMA sample (Section
-\ref{sec-data}). Given the degeneracies of the model in many of the parameters,
-we perform the fitting procedure in three parameters, while fixing the others
-to a particular choice of values.
-
-{\bf [CONTINUE THIS; SHOW THE FIT FIGURES; SHOW THE MCMC CORNER PLOTS]}
-
-{\bf For HD 163296 we have constraints from the sub/superkepler data of
-Teague. This limits the width of the gas bump.}
+%\subsection{Model fits to the ALMA data}
+%\label{sec-anmodel-fits-to-data}
+%%
+%We now apply this model to the isolated rings in our ALMA sample (Section
+%\ref{sec-data}). Given the degeneracies of the model in many of the parameters,
+%we perform the fitting procedure in three parameters, while fixing the others
+%to a particular choice of values. 
+%
+%
+%
+%{\bf [CONTINUE THIS; SHOW THE FIT FIGURES; SHOW THE MCMC CORNER PLOTS]}
+%
+%
+%
+%
+%{\bf For HD 163296 we have constraints from the sub/superkepler data of
+%Teague. This limits the width of the gas bump.}
 
@@
 
-\subsection{Computing the dust mass trapped in the rings}
+\section{Gauss fitting procedure}
+\label{sec-gauss-fitting}
+%
+In this paper we study each dust ring individually, and try to understand it in
+terms of the trapping of dust in a pressure trap. In Appendix
+\ref{sec-steady-state-analytic-trap-model} we find that for a Gaussian pressure
+bump the solution to the radial dust mixing and drift problem is, to first
+approximation, also a Gaussian, albeit a narrower one. Fitting the dust trapping
+model to the data can therefore be done in two stages: first fitting a Gaussian
+radial profile to the observed rings, then interpreting these Gaussian fits
+through the dust trapping model. In this appendix
+
+\section{Computing the dust mass trapped in the rings}
 \label{sec-computing-dust-mass}
@@
 We will assume that the intrinsic dust temperature is as given in Table
-\ref{tab-gauss-params} (column $T_{\mathrm{model}}$), which was computed based on
+\ref{tab-gauss-params} (column $T_{\mathrm{d}}$), which was computed based on
 a simple flaring angle irradiation model. We consider the Gaussian profile with
@@
 \ref{tab-gauss-params} one can see that for most rings 
-$A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{model,br,lin}}$.
+$A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{d,br,lin}}$.
 However, for the rings of AS 209 the ratios
-$A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}$ are around 0.5, so optical depth
+$A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}$ are around 0.5, so optical depth
 effects start to play a role. Given a dust surface density profile
@@
 \begin{equation}
-T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{model,br,lin}}
+T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{d,br,lin}}
 \end{equation}
@@
 \begin{equation}\label{eq-int-afo-tau-bnu}
-T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{model,br,lin}}
-\end{equation}
-But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}
-=\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{model}}))$ can be as large as
+T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{d,br,lin}}
+\end{equation}
+But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}
+=\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{d}}))$ can be as large as
 $\sim 0.5$, this condition is not always fulfilled.  On the other hand, these
@@
 \begin{equation}
-  \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{model,br,lin}}}\,
+  \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{d,br,lin}}}\,
   \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
@@
 \begin{equation}
-  \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\,
+  \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\,
   \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
@@
     M_{\mathrm{d}} &\simeq 2\pi r_0\int_{r_{\mathrm{in}}}^{r_{\mathrm{out}}}\Sigma_{\mathrm{d}}(r)dr\\
-    &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\\
-    &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}
+    &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\\
+    &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}
   \end{split}

### 2018-07-30 projects/andrews/SEDgallery.py
image gallery and 2018/07/30 paper draft

@@ -10,39 +10,33 @@ plt.rc('font', size=9)
 
 show_median = False
 
-names = ['HTLup', 'GWLup', 'IMLup', 'HD142666', 'RULup',
-         'HD143006', 'Sz129', 'MYLup', 'Sz114', 'AS205',
+names = ['HTLup', 'GWLup', 'IMLup', 'RULup', 'Sz114', 
+         'Sz129', 'MYLup', 'HD142666', 'HD143006', 'AS205',
          'SR4', 'Elias20', 'DoAr25', 'Elias24', 'Elias27',
          'DoAr33', 'WSB52', 'WaOph6', 'AS209', 'HD163296']
 
-labels = ['HT Lup', 'GW Lup', 
-          'IM Lup', 'HD 142666', 
-          'RU Lup', 'HD 143006', 
-          'Sz 129', 'MY Lup', 
-          'Sz 114', 'AS 205',
-          'SR 4', 'Elias 20', 
-          'DoAr 25', 'Elias 24', 
-          'Elias 27', 'DoAr 33', 
-          'WSB 52', 'WaOph 6', 
-          'AS 209', 'HD 163296']
+labels = ['HT Lup', 'GW Lup', 'IM Lup', 'RULup', 'Sz 114', 
+          'Sz 129', 'MY Lup', 'HD 142666', 'HD 143006', 'AS 205',
+          'SR 4', 'Elias 20', 'DoAr 25', 'Elias 24', 'Elias 27', 
+          'DoAr 33', 'WSB 52', 'WaOph 6', 'AS 209', 'HD 163296']
 
-dpc = np.array([154., 155., 158., 148., 159.,
-                165., 161., 156., 162., 128., 
+dpc = np.array([154., 155., 158., 159., 162., 
+                161., 156., 148., 165., 128., 
                 134., 138., 138., 136., 116.,
                 139., 136., 123., 131., 101.])
 
-Teff = 10.**(np.array([3.69, 3.56, 3.63, 3.88, 3.61,
-                       3.75, 3.61, 3.71, 3.50, 3.63,
+Teff = 10.**(np.array([3.69, 3.56, 3.63, 3.61, 3.50, 
+                       3.61, 3.71, 3.88, 3.75, 3.63,
                        3.61, 3.59, 3.63, 3.63, 3.59,
                        3.65, 3.57, 3.62, 3.63, 3.97]))
 
[…]

### 2018-07-30 projects/andrews/psf_demo.py

@@ -176,7 +176,8 @@ for i in range(len(uvsets)):
 
     # plot the (u,v) tracks
     ax = fig.add_subplot(gs1[0,i])
-    ax.plot(u, v, '.k', markersize=0.2, rasterized=True)
+    ax.plot(u, v, '.k', markersize=0.1, alpha=0.2, rasterized=True)
+    ax.plot(-u, -v, '.k', markersize=0.1, alpha=0.2, rasterized=True)
 
     # annotations
     ax.annotate(uvlbls[i], xy=(0.06, 0.86), xycoords='axes fraction',

### 2018-08-13 projects/dullemond/paper_dusttraps/ms.tex
New version of trapping paper.

@@
 \def\mugas{\ensuremath{\mu_\mathrm{gas}}\xspace}
-\def\mp{\ensuremath{m_\mathrm{p}}\xspace}
 \def\Hp{\ensuremath{H_\mathrm{p}}\xspace}
@@
 \def\inprep{in prep.\xspace}
+\def\paperdsharpandrews{Andrews et al.\ ({\bf 2018})}
+\def\paperdsharphuangrings{Huang et al.\ ({\bf 2018})}
+\def\paperdsharpisella{Isella et al.\ ({\bf 2018})}
+\def\paperdsharpguzmann{Guzmann et al.\ ({\bf 2018})}
+\def\paperdsharpbirnstiel{Birnstiel et al.\ ({\bf 2018})}
 %
@@
 \title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
-\authorrunning{Dullemond, Birnstiel, Troncoso, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
+\authorrunning{Dullemond et al.}
 \author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, N.~Troncoso$^{3}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
@@
 that show high-contrast and radially thin rings that are separated by deep
-valleys, and that are sufficiently face-on to not have to worry about 3-D
+valleys, and that are sufficiently face-on to not have to worry much about 3-D
 line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
@@
 \hline
-% AS 209     & 121              & 0.912        &  0.631 \\ % Data from 2018.05.13 Table
 AS 209     & 121              & 0.832        &  1.413 \\ % 2018.07.02
 Elias 24   & 136              & 0.776        &  6.03  \\ % 2018.07.02
-% HD 163296  & 101              & 2.14         &  25.7  \\ % Data from 2018.05.13 Table
 HD 163296  & 101              & 2.04         &  17.0  \\ % 2018.07.02
 GW Lup     & 155              & 0.457        &  0.331 \\ % 2018.07.02
-% HD 143006  & 166              & 0.813        &  0.871 \\ % Data estimated by Kees May/June 2018
 HD 143006  & 165              & 1.78         &  3.80 \\  % 2018.07.02 ** Big change **
@@
 \end{center}
-\caption{\label{tab-stellar-params}The stellar parameters assumed for the
-  stars studied in this paper. Distance is in parsec and mass and luminosity
-  are in units of the solar values.}
+\caption{\label{tab-stellar-params}The stellar parameters assumed for the stars
+  studied in this paper. Distance is in parsec and mass and luminosity are in
+  units of the solar values. More details, as well as references and uncertainty
+  estimates, can be found in \paperdsharpandrews{}.}
 \end{table}
 
-A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.
-For an overview of the ALMA Large Programme we refer to {\bf Andrews et
-  al.~XXXX}, and for an in-depth discussion on the data of the individual
-sources we refer to {\bf [LIST HERE THE PAPERS]}.
+A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.  For an
+overview of the ALMA Large Programme we refer to \paperdsharpandrews{}, and for
+an in-depth discussion on the data of the individual sources we refer to
+\paperdsharphuangrings{}, \paperdsharpisella{}, \paperdsharpguzmann{}.
 
@@
   are marked in the images. For a detailed description of these
-  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
+  data, see \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
+  for HD 163296, and \paperdsharphuangrings{} for the rest.}
 \end{figure*}
 
-The high-contrast rings of these sources provide ``clean laboratories''
-for testing the theory of dust trapping in a ring-by-ring manner.
-Fig.\ \ref{fig-obs-profiles}
-shows the radial profile (deprojected for
+The high-contrast rings of these sources provide ``clean laboratories'' for
+testing the theory of dust trapping in a ring-by-ring manner.
+Fig.\ \ref{fig-obs-profiles} shows the radial profile (deprojected for
 inclination) of the thermal emission of the dust of the five disks. The
-procedure used to extract these radial profiles from the continuum maps
-is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
-HD 143006 we removed the blob]}. The
-brightness is expressed as linear brightness temperature
-$T_{\mathrm{br,lin}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
-intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
-frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
-defined such that $B_\nu(T_{\mathrm{br,full}})=I_\nu$, with $B_\nu(T)$ the
-Planck function, can have very different values, in particular at low
-brightness. As a result $T_{\mathrm{br,full}}(r)$ tends to vastly
-underestimate the contrast between the rings and the inter-ring
-regions. Therefore, in this paper we plot the brightness using the linear
-brightness temperature, which is proportional to the intensity in Jy/beam.}.
+procedure used to extract these radial profiles from the continuum maps is
+described by \paperdsharphuangrings{}, though we removed the blob seen in HD
+163296 and HD 143006 for this work.
+
+The brightness profiles are expressed as linear brightness temperature
+$I_\nu^{\mathrm{lb}}$, which we define in Appendix \ref{sec-conventions}. The
+choice of linear brightness temperature over full brightness temperature is
+important, because the full brightness temperature tends to vastly underestimate
+the contrast between the rings and the inter-ring regions. The linear brightness
+temperature is instead linearly proportional to the intensity in Jy/beam, and
+thus represents the contrast appropriately.
 
@@
   with the width of the ALMA beam. For a detailed description of these
-  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
+  data, \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
+  for HD 163296, and \paperdsharphuangrings{} for the rest.}
 \end{figure*}
 
-As is shown in Appendix \ref{sec-steady-state-analytic-trap-model}, for a
+\section{Fitting a Gaussian profile to the ring emission}
+\label{sec-gauss-fits}
+%
+As we will discuss later (Section \ref{sec-rings-as-dust-traps} and Appendix
+\ref{sec-steady-state-analytic-trap-model}), for a
 Gaussian pressure bump the solution to the radial dust mixing and drift problem
@@
 we put emphasis on the shape of the peak of the rings, and deemphasize the
-wings.
-
-The fitting procedure is described in Appendix \ref{sec-gauss-fitting}.  The
-fitting was done in the image plane, in order to make it possible to fit to a
-limited radial range, i.e.\ selecting a single ring. The Gaussian fits are shown
-as inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
+wings. {\bf CHECK}.
+
+The aim is now to find, for each ring, a Gaussian linear brightness temperature
+profile
+\begin{equation}\label{eq-gauss-lin-br-temp}
+I_{\nu}^{\mathrm{lb,gauss}}(r) = A\,\exp\left(-\frac{(r-r_0)^2}{2\sigma^2}\right)
+\end{equation}
+that best describes the ring. To be more precise: We determine the values of
+$A$, $r_0$ and $\sigma$ for which Eq.~(\ref{eq-gauss-lin-br-temp}) best fits the
+observed linear brightness temperature profile $I_{\nu}^{\mathrm{lb,obs}}(r)$
+shown in Fig.~\ref{fig-obs-profiles} within a prescribed radial domain as given
+in Table \ref{tab-gauss-params}. Details of the fitting procedure are described
+in Appendix \ref{sec-gauss-fitting-procedure}. The Gaussian fits are shown as
+inverse parabolas in Fig.\ \ref{fig-obs-profiles}. In
 Fig.~\ref{fig-obs-gaussfits} these Gaussian fits are shown in close-up. The
-parameters of the best fits are listed in Table \ref{tab-gauss-params}.
+parameters of the best fits are listed in Table \ref{tab-gauss-params}.  For the
+error estimates on these values, see Appendix \ref{sec-gauss-fitting-procedure}.
+
+{\bf [ADD APPENDIX WITH DETAILS OF FITS AND ERROR ESTIMATES]}
 
@@
 \caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
-  paper. The blue curves are the observations, the green curves are the best fit
+  paper. The blue curves are the observations, the orange curves are the best fit
   Gaussian profiles. The ``fit range'' bar shows the radial range within which
@@
 \begin{center}
-\begin{tabular}{|cc|cc|ccccc|cccccc|}
+\begin{tabular}{|cc|cc|cccccc|ccccc|}
 \hline
@@
 %            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
-Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
-            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+%Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
+%            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+%Source & Ring & beam & domain & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $T_{\mathrm{d,br,lin}}$ & $w_d/h_p$ & $\tau$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
+%            & & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
+Source & Ring & beam & domain & $A$ & $A_{\mathrm{dec}}$ & $r_0$ & $\mathrm{FWHM}$ & $\sigma$ & $w_d$ & $T_{\mathrm{d}}$ & $B_{\mathrm{d}\nu}^{\mathrm{lb}}$ & $w_d/h_p$ & $\tau_\nu^{\mathrm{peak}}$ & $M_{\mathrm{d}}$\\
+            & & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  &  & $[M_{\oplus}]$\\
 \hline
@@
 % Note: New table made with new stellar parameters 2018.07.02
-AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  15.8 &  10.9 &   0.6 &   0.41 &  33.3 &   1.81\\
-AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  12.4 &   7.7 &   0.4 &   0.48 &  74.3 &   4.04\\
-Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.36 &  41.1 &   2.23\\
-HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  30.8 &  25.6 &   1.7 &   0.40 &  56.9 &   3.09\\
-HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  25.3 &  20.2 &   0.8 &   0.28 &  48.3 &   2.62\\
-GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &   0.30 &  36.3 &   1.97\\
-HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  72.20 &   5.06 &   4.42 &  27.2 &  22.0 &   2.1 &   0.16 &  10.5 &   0.57\\
-HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 112.51 &   7.88 &   7.49 &  21.6 &  16.6 &   2.0 &   0.16 &  26.5 &   1.44\\
+% Note: New table with new beam size estimates 2018.08.06
+% Note: Added beam size to table, changed a bit the layout, and used new opacities for mass estimates (though still needs improving) 2018.08.07
+% Note: Removed the Ivezic mass column
+% Note: Added A_dec
+AS 209     & 1 & 37 &  69 --  79 &   3.64 &   4.18 &  74.1 &  75.4 &   3.87 &   3.38 &  15.8 &  10.9 &   0.6 &   0.48 &  28.5\\
+AS 209     & 2 & 37 & 115 -- 125 &   2.95 &   3.23 & 120.3 &  90.3 &   4.64 &   4.23 &  12.4 &   7.7 &   0.4 &   0.55 &  60.7\\
+Elias 24   & 1 & 35 &  72 --  82 &   5.22 &   5.69 &  76.9 &  88.0 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.40 &  33.4\\
+HD 163296  & 1 & 43 &  60 --  76 &   8.48 &   8.76 &  67.4 & 171.7 &   7.36 &   7.13 &  30.8 &  25.6 &   1.6 &   0.42 &  43.8\\
+HD 163296  & 2 & 43 &  94 -- 104 &   4.92 &   5.20 &  99.9 & 133.6 &   5.73 &   5.42 &  25.3 &  20.2 &   0.8 &   0.30 &  38.0\\
+GW Lup     & 1 & 44 &  79 --  89 &   1.48 &   1.87 &  85.1 &  72.1 &   4.74 &   3.76 &  10.3 &   5.7 &   0.5 &   0.40 &  34.1\\
+HD 143006  & 1 & 45 &  35 --  45 &   3.21 &   4.11 &  41.0 &  72.2 &   5.06 &   3.95 &  27.2 &  22.0 &   1.9 &   0.21 &  10.0\\
+HD 143006  & 2 & 45 &  59 --  72 &   2.48 &   2.71 &  64.8 & 112.5 &   7.88 &   7.22 &  21.6 &  16.6 &   2.0 &   0.18 &  21.6\\
 \hline
@@
 \caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
-  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values
-  $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$ are the inner and outer boundary of
-  the fitting domain. $A$ is the peak linear brightness temperature of the
-  best-fit Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$,
-  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
-  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
-  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
-  dust emission profile, also expressed as standard deviation in units of
-  $\mathrm{au}$. The last six columns are derived from a simple disk model
-  assumption: $T_{\mathrm{d}}$ is the midplane temperature of the disk (we
-  assume gas and dust temperature to be equal) computed from
-  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
-  $\varphi=0.02$. From that one can derive the corresponding linear brightness
-  temerature $T_{\mathrm{d,br,lin}}=B_\nu(T_{\mathrm{d}})c^2/2\nu^2$ where
-  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of 0.13
-  $\mathrm{cm}$ of ALMA band 6.  The fourth-to-last column expresses the
-  deconvolved dust ring width in units of the disk pressure scale height $h_p$
-  computed from $T_{\mathrm{d}}$. The $\tau$ column gives the estimated optical
-  depth at the peak of the ring, calculated from
-  $\tau=-\ln(1-A/T_{\mathrm{d,br,lin}})$.  The last two columns are the computed
-  dust mass in the ring in units of Earth masses, assuming the small grain
-  absorption opacity
-  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
-  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
-  value
-  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
-  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). In making these
-  estimates, the optically thin approximation was used. See Appendix
-  \ref{sec-computing-dust-mass} for details. An estimate of the
-  optical-depth-corrected dust mass can be made by multiplying the mass values
-  in the table by $\tau/(1-e^{-\tau})$.}
+  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The ``beam''
+  column gives the beam size (full-width-at-half-max). Given that in reality the
+  beam is elliptic, and the disk has an inclination, this is merely an estimate.
+  The ``domain'' column gives the radial domain within which the Gaussian fit
+  was made. $A$ is the peak linear brightness temperature of the best-fit
+  Gaussian ring model, $r_0$ is its radius in $\mathrm{au}$, $\mathrm{FWHM}$ is
+  the width expressed as full-width-at-half-maximum in units of milliarcsecond,
+  $\sigma$ is its width expressed as standard deviation in units of
+  $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved) dust
+  emission profile, also expressed as standard deviation in units of
+  $\mathrm{au}$. The $A_{\mathrm{dec}}$ column is like $A$, but corrected for
+  the deconvolution (see Eq.~\ref{eq-a-deconv}). The last five columns are
+  derived from a simple disk model assumption: $T_{\mathrm{d}}$ is the midplane
+  temperature of the disk (we assume gas and dust temperature to be equal)
+  computed from Eq.~(\ref{eq-disk-temperature-model}), assuming a flaring angle
+  of $\varphi=0.02$. From that one can derive the corresponding linear
+  brightness temperature $B_{\mathrm{d}\nu}^{\mathrm{lb}}$ using
+  Eq.~(\ref{eq-app-bnu-bright}) for $\nu=c/(0.13\,\mathrm{cm})$ (ALMA band 6).
+  The $w_d/h_p$ column expresses the deconvolved dust ring width in units of the
+  disk pressure scale height $h_p$ computed from $T_{\mathrm{d}}$. The
+  $\tau_\nu^{\mathrm{peak}}$ column gives the estimated optical depth at the peak of the ring,
+  calculated from Eq.~(\ref{eq-tau-estimate}). The last column is the computed dust mass in the
+  ring in units of Earth masses. See Appendix \ref{sec-computing-dust-mass} for
+  details. In making these mass estimates we use the DSHARP dust opacity model
+  (\paperdsharpbirnstiel{}) for a grain radius of $a=0.1\,\mathrm{cm}$, which
+  yields an absorption opacity
+  $\kappa_\nu^{\mathrm{abs}}(\lambda=0.13\,\mathrm{cm})=1.93\,\mathrm{cm}^2/\mathrm{g}$.
+  The uncertainty of the opacity value is, however, very large, since we do not
+  know apriori the grain size distribution.}
 \end{table*}
@@
 The observed rings are the result of the thermal emission of a dust ring convolved
-with the ALMA beam. The standard deviation width $\sigma$ of the fitted Gaussian
-profile is given in Table \ref{tab-gauss-params}.  To obtain the width of the
+with the ALMA beam. To obtain the width of the
 underlying dust ring we have to deconvolve. Assuming a Gaussian beam and a
@@
 obtain the width $w_d$ of the dust ring
-\begin{equation}
+\begin{equation}\label{eq-simple-deconvolve-gauss}
 w_d=\sqrt{\sigma^2-b^2}
@@
 $\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
-$b_{\mathrm{fwhm,as}}$ is the beam full width at half maximum in units of
+$b_{\mathrm{fwhm,as}}$ is the beam full-width at half-maximum (FWHM) in units of
 arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
-parsec). The width of the dust rings
+parsec). In using Eq.~(\ref{eq-simple-deconvolve-gauss}) we have ignored the
+elliptical shape of the beam as well as the inclination of the disk.
+
+The slightly narrower deconvolved ring should also have a correspondingly
+higher amplitude $A_{\mathrm{dec}}$ given by
+\begin{equation}\label{eq-a-deconv}
+A_{\mathrm{dec}} = \frac{\sigma}{w_d} \, A
+\end{equation}
+to conserve luminosity, where we ignore the geometric effects due to the
+circular coordinates. The values of $A_{\mathrm{dec}}$ are listed in Table
+\ref{tab-gauss-params} as well.
+
+For completeness, let us note that the deconvolved Gaussian model then becomes
+\begin{equation}\label{eq-gauss-lin-br-temp-deconv}
+I_{\nu}^{\mathrm{lb,gauss,dec}}(r) = A_{\mathrm{dec}}\,\exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+\end{equation}
+
+The width of the dust rings
 in units of the local pressure scale height $h_p$ (see Table
 \ref{tab-gauss-params}) can only be roughly estimated, because we do not know
-the disk midplane temperature. Our estimate is made as follows:
+the disk midplane temperature. Our estimate of the midplane disk temperature
+is made using the following simple irradiated flaring disk recipe:
 \begin{equation}\label{eq-disk-temperature-model}
@@
 the so-called flaring angle \citep[e.g.][]{1997ApJ...490..368C,
-  1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
+  1998ApJ...500..411D, 2001ApJ...560..957D}. The factor of $1/2$ in front of the
+flaring angle $\varphi$ originates from the consideration that the stellar
+radiation heats the surface layer of the disk, which then radiates half of that
+energy away and half into the deeper regions of the disk. Only the latter half
+goes into the energy balance equation between irradiation and radiative cooling
+of the disk interior. We take the flaring angle to be $\varphi=0.02$ which is an
+estimate based on typical values from models. In using this simple recipe we
+assume that the optical and infrared dust opacity is dominated by a population
+of small grains that is vertically more extended than the larger dust grains
+seen in the ALMA images. The two dust populations may
+be largely decoupled (see e.g.\ the BL-series of models of
+\citet{2004A&A...417..159D} as an illustration). The value of $\varphi=0.02$ is
+nothing more than an educated guess, so its value can easily be wrong by a
+factor of four or so. But the fact that the pressure scale height goes as
+$T_{\mathrm{d}}^{1/2}\propto \varphi^{1/8}$ means that such an error only
+affects $h_p$ mildly.
+
+Assuming that the gas temperature is equal to the dust temperature, the
+pressure scale height of the disk now
 follows from $h_p=\sqrt{k_BT_{\mathrm{d}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
@@
 $\mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
-taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
-$\varphi=0.02$ which is an estimate based on typical values from models.  But
-this value can easily be wrong by a factor of four or so. The fact that the
-pressure scale height goes as $T_{\mathrm{d}}^{1/2}\propto \varphi^{1/8}$
-means, however, that such an error only affects $h_p$ mildly.
+taken from Table \ref{tab-stellar-params}.
 
@@
 ring produced by the dust is substantially narrower than $h_p$, then some kind
-of dust trapping must have taken place. If this were not the case, then the dust
-emission profile would have the same shape as the gas midplane density, and thus
-the gas midplane pressure bump would then be a narrower than $h_p$.
-
-We see that 5 out of 8 rings are narrower than $h_p$, but not by much. None are
-less than half the width of the pressure scale height of the gas. This is
-perhaps slightly disappointing, because it means that these results are no
-irrefutable proof that dust trapping is playing a role in the shaping of these
-rings. But it is strong evidence for that nonetheless.
-
-For these observations the fwhm beam size is of the order of 35 milliarcseconds
-{\bf [CHECK AND MAKE MORE ACCURATE]}. This number can be compared to the fwhm
-ring width in Table \ref{tab-gauss-params}, and it shows that the widths of all
-the rings are spatially resolved. Some are only marginally resolved (AS 209
-rings 1 and 2, GW Lup and the ring 1 of HD 143006), but the data are
-inconsistent with radially unresolved rings. This is confirmed by more detailed
-analysis in the uv-plane ({\bf REFS TO Guzm\'an et al.~(2018)}).
-
-These Gaussian profiles fit the observed data remarkably well, as can be
-best seen in Fig.~\ref{fig-obs-gaussfits}. 
-For the broad rings of HD 163296 the closeness to a Gaussian profile is
-particularly striking, given that these rings are resolved by many beam
-widths. In particular in this source there is therefore no doubt that the
-Gaussian profile is a reflection of a real phenomenon, and is not caused by
-unknown observational or data reduction issues.  But also the rings of AS 209,
-while not as well-resolved as those of HD 163296, are remarkably close to
-Gaussian.
-
-For Elias 24 the ring fits a Gaussian near the peak, but deviates from Gaussian
-already less half a width from the center. Likewise for the rings of HD
-143006. The ring around GW Lup is an intermediate case, which fits a Gauss well
-on the inner flank, but rather badly on the outer flank.
+of dust trapping must have taken place. We can thus conclude that we have strong
+evidence of dust trapping operating in the rings of AS 209, Elias 24 and GW Lup.
+For the other rings dust trapping is certainly not ruled out either, but would
+require further evidence.
+
+For these observations the FWHM beam size for each ring is listed in Table
+\ref{tab-gauss-params}. These are azimuthal average beam size values according
+to Eq.~(\ref{eq-def-average-beam}), based on the values listed in
+\paperdsharphuangrings{}. This number can be compared to the FWHM ring width in
+Table \ref{tab-gauss-params}, and it shows that the widths of all the rings are
+spatially resolved. Some are only marginally resolved (AS 209 rings 1 and 2, GW
+Lup and the ring 1 of HD 143006), but the data are inconsistent with radially
+unresolved rings. For AS 209 This is confirmed by more detailed analysis in the
+uv-plane by \paperdsharpguzmann{}. {\bf CHECK IF HUANG DOES THIS, TOO.}
+
+These Gaussian profiles fit the observed data remarkably well, as can be best
+seen in Fig.~\ref{fig-obs-gaussfits}.  For the broad rings of HD 163296 the
+closeness to a Gaussian profile is particularly striking, given that these rings
+are resolved by many beam widths.  For Elias 24 the ring fits a Gaussian near
+the peak, but deviates from Gaussian already less half a width from the
+center. Likewise for the rings of HD 143006. The ring around GW Lup is an
+intermediate case, which fits a Gauss well on the inner flank, but rather badly
+on the outer flank.
 
@@
 
-Given the closeness of the profiles to a Gaussian, we can use the Gaussian fit
-to make estimates of the total mass of dust trapped in each of these rings. The
-procedure is described in Appendix \ref{sec-computing-dust-mass}, and the
-results, for the two opacity choices we make, are listed in Table
-\ref{tab-gauss-params}. In making these estimates we have assumed the dust to be
-made up of micron-sized dust, which is likely too small. The mass based on the
-Ivezic opacity, $M_{\mathrm{d,iv}}$, will increase for $a>\lambda/2\pi$ by a
-factor of $2\pi\,a/\lambda$ with $\lambda=0.13\,\mathrm{cm}$. For the mass based
-on the Ricci opacity the factor first decreases when $a$ approaches the opacity
-peak around $a=\lambda/2\pi$ and then increases for $a>\lambda/2\pi$. Both
-behaviors can be seen as the inverse behavior of the opacities shown in
-Fig.~\ref{fig-opacities}. For ever increasing grain size estimate, the mass
-estimate will increase indefinitely for both opacity choices.
+\section{Optically thin analysis}
+\label{sec-opt-thin-analysis}
+%
+As a first guess, we will assume that we can ignore optical depth effects,
+i.e.\ that the thermal emission of the dust is optically thin. The linear
+brightness temperature profiles shown in Section \ref{sec-data}, after
+deconvolution with the beam, are then linear maps of the spatial distribution of
+dust, if we ignore any temperature gradients across these rings.  In other
+words: the deconvolved versions of the Gaussian fits of Section
+\ref{sec-gauss-fits} (Eq.~\ref{eq-gauss-lin-br-temp-deconv}) will then directly
+translate in Gaussian models of the radial dust distribution.
+
+The conversion between the deconvolved observed linear brightness temperature
+profile $I_\nu^{\mathrm{lb,dec}}(r)$ and the dust surface density profile
+$\Sigma_d(r)$ is then
+\begin{equation}\label{eq-optthin-conversion-linbright-sigmad}
+  %I_\nu^{\mathrm{lb}} = \Sigma_d\,\kappa_\nu^{\mathrm{abs}}\, B_\nu^{\mathrm{lb}}(T_d)
+  \Sigma_d(r) =  \frac{I_\nu^{\mathrm{lb,dec}}(r)}{\kappa_\nu^{\mathrm{abs}}\, B_\nu^{\mathrm{lb}}(T_d)} 
+\end{equation}
+where $T_d$ is the temperature of the dust, $\kappa_\nu^{\mathrm{abs}}$ is the
+absorption opacity, and $B_\nu^{\mathrm{lb}}(T_d)$ is the Planck function
+expressed as linear brightness temperature (Eq.~\ref{eq-app-bnu-bright}).
+
+By replacing $I_\nu^{\mathrm{lb,dec}}(r)$ with the Gaussian fit
+$I_\nu^{\mathrm{lb,gauss,dec}}(r)$ given by
+Eq.~(\ref{eq-gauss-lin-br-temp-deconv}) we obtain the corresponding
+$\Sigma_d^{\mathrm{gauss}}(r)$ from
+Eq.~(\ref{eq-optthin-conversion-linbright-sigmad}). From this Gaussian
+model we can derive the total dust mass trapped in the ring:
+\begin{equation}
+  M_d = \int_0^\infty 2\pi r \Sigma_d(r) dr \simeq
+  \frac{(2\pi)^{3/2} r_0\, A\,\sigma}{\kappa_\nu^{\mathrm{abs}}\, B_\nu^{\mathrm{lb}}(T_d)}
+\end{equation}
+where we used the identity $A\,\sigma=A_{\mathrm{dec}}\,w_d$.
+
+We use the DSHARP opacity model (\paperdsharpbirnstiel{}) which, for a grain
+radius of $a=0.1\,\mathrm{cm}$ yields a dust opacity of
+$\kappa_\nu^{\mathrm{abs}}(\lambda=0.13\,\mathrm{cm})=1.93\,\mathrm{cm}^2/\mathrm{g}$.
+The resulting dust mass estimates are listed in Table \ref{tab-gauss-params}.
+
+The main uncertainty lies in the uncertainty in the opacity value
+$\kappa_\nu^{\mathrm{abs}}$. This value depends on the grain size (or grain size
+distribution) as well as many other factors including composition, grain shape
+and uncertainties in the method of computation of the opacity. We refer to
+\paperdsharpbirnstiel{} for a discussion of the opacity model used for the
+DSHARP campaign.
+
+{\bf [Discuss the minimum and maximum mass values]}
+
+The other uncertainty, though much less severe, is the dust temperature
+$T_d$. From the continuum images we have no information about $T_d(r)$. From the
+CO line emission one can estimate the temperature in the disk surface layers,
+but not in the midplane \citep[see e.g.][]{2018ApJ...853..113W}. Instead, we
+will use the flaring angle recipe given by Eq.~(\ref{eq-disk-temperature-model})
+with $\varphi=0.02$. The resulting values of $T_d$ at the peak of the rings are
+given in Table \ref{tab-gauss-params}. For
+Eq.~(\ref{eq-optthin-conversion-linbright-sigmad}) we need the corresponding
+linear brightness temperature $B_{\mathrm{d}\nu}^{\mathrm{lb}}\equiv B_\nu^{\mathrm{lb}}(T_d)$, which is listed in the
+next column in Table \ref{tab-gauss-params}.
+
+Given the amplitude of the deconvolved Gauss fit $A_{\mathrm{dec}}$ (see
+Eq.~\ref{eq-gauss-lin-br-temp-deconv}), we can derive the optical depth
+$\tau_\nu^{\mathrm{peak}}$ of the ring at its peak at $r=r_0$:
+\begin{equation}\label{eq-tau-estimate}
+\tau_\nu^{\mathrm{peak}} = -\ln\left(1-\frac{A_{\mathrm{dec}}}{B_{\nu}^{\mathrm{lb}}(T_d)}\right)
+\end{equation}
+This allows us to verify if our assumption of optical thinness is justified. The
+results are listed in Table \ref{tab-gauss-params}.  We conclude that the
+optically thin assumption is not entirely wrong, but not quite right either. The
+estimated dust surface density $\Sigma_d$ and dust mass $M_d$ derived from the
+optically thin assumption ( Eqs.~\ref{eq-optthin-conversion-linbright-sigmad}
+and {\bf XXX}, respectively) are only mildly affected by the optical depth
+effects.
+
+A much stronger effect on the conversion of the observations to dust mass
+remains the uncertainty of the opacity. As shown in \paperdsharpbirnstiel{}
+the value of $\kappa_\nu^{\mathrm{abs}}=1.93\,\mathrm{cm}^2/\mathrm{g}$ that
+we use here can easily be wrong by a factor of $10$ upward or a factor of
+$0.1$ downward, with correspondingly large changes in the derived dust mass.
+
+\section{Optical depth effects}
+\label{sec-optical-depth-effects}
+%
+As we have seen in Section \ref{sec-opt-thin-analysis}, the optical depth of
+these dust rings is not very low: about 0.2 and upward, usually around 0.4. For
+the case of HD 163296 there is evidence from the absorption of CO line emission
+from the back side of the disk that the optical depth in the two prominent rings
+is even close to unity, as shown by \paperdsharpisella{}. For simplicity we will
+perform most of our analysis of the later Sections using the optically thin
+approximation. But before doing that, it is important to investigate how strong
+the effect of optical depth is, and how it would affect our results. This is the
+topic of this Section.
+
+\subsection{Simple optical depth correction}
+Let us assume that the dust has zero albedo. Then the inclusion of the effect of
+optical depth is simple. We replace
+Eq.~(\ref{eq-optthin-conversion-linbright-sigmad}) with the formal transfer
+equation:
+\begin{equation}\label{eq-simple-formal-rt}
+  I_\nu^{\mathrm{lb,dec}}(r) = e^{-\tau_\nu(r)}I_\nu^{\mathrm{lb,bg}} +
+  \left(1-e^{-\tau_\nu(r)}\right)B_\nu^{\mathrm{lb}}(T_d)
+\end{equation}
+where $\tau_\nu(r)$ is the optical depth profile across the ring, and
+$I_\nu^{\mathrm{lb,bg}}$ is the background brightness, either from background
+clouds or from the cosmic microwave background. In interferometric observations
+a flat background is Fourier-filtered out, but in the radiative transfer it
+plays a role in the non-linear optical depth regime ($\tau_\nu\gtrsim 0.3$),
+even if we are only interested in the excess emission above the background
+$I_\nu-I_\nu^{\mathrm{bg}}$. We assume, however, that the cloud background
+emission is negligible. The linear brightness temperature of the cosmic
+background at $\lambda=0.13\,\mathrm{cm}$ is
+$I_\nu^{\mathrm{lb,cmb}}=0.195\,\mathrm{K}$. This is not negligible compared to
+the brightness temperatures in the inter-ring regions, but those regions are low
+optical depth regions, where the radiative transfer equation for for
+$I_\nu-I_\nu^{\mathrm{bg}}$ is linear and the background drops out. Only near
+the peak of the rings the optical depth becomes non-negligible, but the value of
+$B_{\mathrm{d}\nu}^{\mathrm{lb}}\equiv B_\nu^{\mathrm{lb}}(T_d)$ is always much
+larger than the cosmic background value of $0.195\,\mathrm{K}$ (see Table
+\ref{tab-gauss-params}). We will therefore ignore the background and set
+$I_\nu^{\mathrm{lb,bg}}=0$ in Eq.~(\ref{eq-simple-formal-rt}) for simplicity.
+
+To obtain the dust distribution we first compute $\tau_\nu(r)$
+\begin{equation}\label{eq-tau-profile}
+\tau_\nu(r) = -\ln\left(1-\frac{I_\nu^{\mathrm{lb,dec}}(r)}{B_\nu^{\mathrm{lb}}(T_d)}\right)
+\end{equation}
+which is identical to Eq.~(\ref{eq-tau-estimate}), but now for the entire
+profile instead of just for the peak. The profile for $\Sigma_d(r)$ now follows
+from
+\begin{equation}
+\Sigma_d(r) = \frac{\tau_\nu(r)}{\kappa_\nu^{\mathrm{abs}}}
+\end{equation}
+
+\subsection{Saturation and flat-topped ring profiles}
+\label{sec-saturation-flat-top-profiles}
+%
+Given the non-linearity of Eq.~(\ref{eq-tau-profile}), in particular near
+the peak of the ring profile, the resulting $\Sigma_d(r)$ profile becomes
+non-Gaussian even if the measured $I_\nu^{\mathrm{lb,dec}}(r)$ has a Gaussian
+shape. Or put the other way: if $\Sigma_d(r)$ is a Gaussian profile, the
+resulting $I_\nu^{\mathrm{lb,dec}}(r)$ will not: it will lead to a flattened
+top, where $I_\nu^{\mathrm{lb,dec}}(r)$ saturates to $B_\nu^{\mathrm{lb}}(T_d)$.
+See Fig.~\ref{fig-flat-topped} for a quantitative example.
+
+\begin{figure}
+  \centerline{\includegraphics[width=0.47\textwidth]{saturation_example.pdf}}
+  \caption{\label{fig-flat-topped}The saturation effect of a gaussian emissivity
+    profile at various peak optical depths. Solid curves: the result of the
+    formal radiative transfer equation (Eq.~\ref{eq-simple-formal-rt} with
+    background set to zero). Dotted curves: the unsaturated case, normalized to
+    have the same total integral as the solid curves.}
+\end{figure}
+
+The Gaussian fits of Section \ref{sec-gauss-fits} suggest, however, that the
+$I_\nu^{\mathrm{lb,dec}}(r)$ profiles do not show clear signs of being
+flat-topped. A simple comparison to Fig.~\ref{fig-flat-topped} shows that peak
+optical depths of $\gg 1$ can be clearly excluded.
+
+However, the deviation from true Gaussian becomes small for $\tau_\nu\lesssim
+1$. It is therefore an important question how strongly we can exclude the
+flattening of the top of the peak, or in other words: how strong an upper limit
+we can set on $\tau_\nu^{\mathrm{peak}}$. Can the flattening of the underlying
+Gaussian emissivity due to self-absorption be mistaken for a different Gaussian
+profile that is wider and lower than the underlying dust emissivity, because the
+deviations vanish within the errorbars on the measurements? We investigate this
+in detail in Appendix \ref{sec-mimicry-flat-topped}. We find that up to
+$\tau_\nu\lesssim 1$ the flattened-out profile can still be reasonably well
+fitted by a Gaussian function. This Gauss, of course, has a strongly reduced
+amplitude compared to the optically thin approximation: by a factor
+$(1-e^{-\tau_\nu^{\mathrm{peak}}})/\tau_\nu^{\mathrm{peak}}$, before beam
+smearing, in order to mimick the self-absorption. Also, it has a slightly wider
+shape, but this is a small effect.
+
+This means that, just from the analysis of the shape of the radial profiles of
+the rings, we can only exclude large optical depth ($\tau_\nu^{\mathrm{peak}}\gg
+1$), but not mild optical depth $\tau_\nu^{\mathrm{peak}}\simeq 0.3\cdots 1$.
+The optically thin treatment, as we will adopt in Section
+\ref{sec-rings-as-dust-traps} and onward, could therefore lead to ignoring
+optical depth effects of the order of
+$(1-e^{-\tau_\nu^{\mathrm{peak}}})/\tau_\nu^{\mathrm{peak}}$, which amounts to
+an underestimation of dust mass of up to 37\% for $\tau_\nu^{\mathrm{peak}}$ up
+to 1.
+
+There is, however, another way to constrain these optical depth effects.  As we
+have already seen in Section \ref{sec-gauss-fits}: given the measured linear
+brightness temperature $I_\nu^{\mathrm{lb,obs}}$, and given a dust temperature
+$T_d$, the optical depth follows immediately. In other words: a larger optical
+depth than given in Table \ref{tab-gauss-params} requires a lower dust
+temperature than given in that table. These temperature estimates are already on
+the low side, by assuming the somewhat low value of the flaring angle
+$\varphi=0.02$ in Eq.~(\ref{eq-disk-temperature-model}). It requires a quite
+strong reduction of $\varphi$ to reduce the temperature sufficiently to increase
+the $\tau_\nu^{\mathrm{peak}}$ from Table \ref{tab-gauss-params} to unity. As an
+example let us take ring 1 of HD 163296 with $T_d=30.8\,\mathrm{K}$ and
+$\tau_\nu^{\mathrm{peak}}=0.42$ according to Table \ref{tab-gauss-params}.
+According to Eq.~(\ref{eq-tau-estimate}), if we set
+$\tau_\nu^{\mathrm{peak}}=1$, we obtain from $A_{\mathrm{dec}}=8.76$ the value
+$B_\nu^{\mathrm{lb}}(T_d)=13.9\,\mathrm{K}$, corresponding, via
+Eq.~(\ref{eq-app-bnu-bright}) to $T_d=18.9\,\mathrm{K}$. This temperature is
+almost twice as low as the original estimate, and would require, according to
+Eq.~(\ref{eq-disk-temperature-model}), a flaring angle of
+$\varphi=0.00082$. Such a flaring angle is unreasonably low and it invalidates
+the simple flaring angle recipe in the first place. Instead we would need to
+perform detailed 2-D/3-D radiative transfer calculations of the disk to
+investigate if such a low temperature is physically attainable. Or equivalently,
+such a study may allow us to find a theoretical lower limit to the dust
+temperature, which then translates into an upper limit to the
+$\tau_\nu^{\mathrm{peak}}$. Alternatively, and possibly more reliably, if we
+obtain a high-resolution map in a second band, we can investigate the behavior
+of the spectral index over the radial profile of the rings, which also provides
+information about the optical depth.
+
+\subsection{The effect of scattering albedo}
+If the dust grains have a radius $a$ comparable to the wavelength of our
+observations, the scattering albedo can be quite high. This means that the
+absorption optical depth can be substantially lower than unity, even if the full
+extinction optical depth (absorption plus scattering) is unity or larger.  The
+near-unity extinction found in HD 163296 by \paperdsharpisella{} from the CO
+maps could thus be compatible with the sub-unity absorption optical depth we
+derived in out Gaussian fitting procedure (see Table \ref{tab-gauss-params}).
+
+In fact, staying with the case of HD 163296, if we would assume that the albedo
+is zero, i.e.\ that the measured extinction optical depth from the CO maps
+equals the absorption optical depth, then we would find rather low dust
+temperatures at the location of the rings, which may be hard to explain
+theoretically, as shown in Section \ref{sec-saturation-flat-top-profiles}.
+If, however, part of the extinction is due to scattering, then it is easier
+to remain consistent with the dust temperature estimated from the flaring
+angle recipe.
+
+However, when scattering is included, the radiative transfer becomes more
+complex than a simple use of a factor $1-e^{-\tau_\nu}$. For completeness we
+describe, in Appendix \ref{sec-oned-radtrans-model}, an approximate solution to
+this problem for a thin slab model.  In principle one would have to replace, in
+the above sections, as well as in Appendix \ref{sec-mimicry-flat-topped}, all
+instances of $1-e^{-\tau_\nu}$ with the more detailed radiative transfer model
+of Appendix \ref{sec-oned-radtrans-model}. But to avoid too much complexity in
+this paper, we leave this to an interested reader.
+
+\subsection{Why are most rings so ``fine-tuned''?}
+It is rather striking that in the analysis of the rings up to this point we have
+found several rather ``fine-tuned'' properties. For instance, the rings of AS
+209, Elias 24, GW Lup and the inner ring of HD 143006 have a width that is only
+roughly twice the beam size (between $1.5$ and $2.4$ times, to be precise), but
+none are unresolved. Given the small sample it is very well possible that this
+is just coincidence. The fact that some rings (in particular those of HD 163296)
+are clearly much wider, lends some support to this.
+
+The derived peak optical depths for most sources (except HD 143006), assuming
+our model of the dust temperature is correct, hover around 0.4, i.e.\ just in
+between the optically thin and optically thick regime. Is this coincidence as
+well, or is there a deeper reason behind this? In this paper we have focussed on
+the strongest, most isolated rings in the DSHARP sample, which could mean that
+this is a selection effect.
+
+Finally, all ring profiles are remarkably similar to a Gaussian shape. On the
+one hand, this may be something similar to the effect discussed in Section
+\ref{sec-saturation-flat-top-profiles} and Appendix
+\ref{sec-mimicry-flat-topped}. And with some rings being only about two beams
+wide, any `peaklike' shape will tend to look similar to a Gaussian. On the other
+hand, the two rings or HD 163296 are resolved by more than 3 beams, yet they
+show a Gaussian shape well into the wings.
+
+Could it be that some observational or data reduction effect may have smeared
+out our data more than expected? We have, to the best of our abilities, been
+unable to identify such an effect. And, as described in Appendix {\bf XXXX}, we
+have performed a full self-test cycle of a mock ring being processed through our
+pipeline, and found no excessive smearing-out, nor effects that would skew our
+conclusion that the rings indeed have an intrinsically Gaussian profile. It seems
+that these ``fine-tuned'' properties of the rings are real.
 
@@
 % Note: New table made with new stellar parameters 2018.07.02
-AS 209     & 1 &  5.59 & 19.6 & 1.1e-02 & 2.0e-01 & 1.6e+01 & 15.7 & 3.9e-03 & 3.2e-02 & 6.0e-01 & 1.2e-04\\
-AS 209     & 2 & 10.25 & 19.6 & 1.2e-02 & 2.3e-01 & 6.9e+00 &  5.8 & 9.1e-03 & 5.0e-02 & 2.1e-01 & 4.5e-04\\
+% Note: New table with new beam size estimates 2018.08.06
+AS 209     & 1 &  5.59 & 19.6 & 1.1e-02 & 2.0e-01 & 1.6e+01 & 15.7 & 3.9e-03 & 3.1e-02 & 5.7e-01 & 1.2e-04\\
+AS 209     & 2 & 10.25 & 19.6 & 1.2e-02 & 2.3e-01 & 6.9e+00 &  5.8 & 9.1e-03 & 4.9e-02 & 2.1e-01 & 4.4e-04\\
 Elias 24   & 1 &  7.26 & 17.1 & 9.7e-03 & 1.8e-01 & 1.8e+01 & 18.8 & 3.6e-03 & 8.0e-02 & 7.0e-01 & 2.9e-04\\
-HD 163296  & 1 &  7.21$^{*}$ & 13.8 & 1.1e-02 & 1.9e-01 & 4.1e+01 & 40.0 & 1.5e-03 & 3.8e-01 & -- & 5.8e-04\\
-HD 163296  & 2 &  7.07 & 13.8 & 7.8e-03 & 1.4e-01 & 2.0e+01 & 27.3 & 3.1e-03 & 1.9e-01 & 1.6e+00 & 5.9e-04\\
-GW Lup     & 1 &  7.49 &  9.9 & 8.3e-03 & 1.5e-01 & 7.8e+00 &  9.8 & 8.0e-03 & 2.1e-01 & 4.4e-01 & 1.7e-03\\
-HD 143006  & 1 &  4.42$^{*}$ & 10.1 & 4.7e-03 & 8.6e-02 & 7.5e+01 & 167.2 & 8.4e-04 & 2.4e-01 & -- & 2.0e-04\\
-HD 143006  & 2 &  7.49$^{*}$ & 10.1 & 4.8e-03 & 8.8e-02 & 3.4e+01 & 73.2 & 1.9e-03 & 1.2e+00 & -- & 2.3e-03\\
+HD 163296  & 1 &  7.13$^{*}$ & 13.8 & 1.1e-02 & 1.9e-01 & 4.1e+01 & 40.0 & 1.5e-03 & 3.6e-01 & -- & 5.6e-04\\
+HD 163296  & 2 &  7.07 & 13.8 & 7.8e-03 & 1.4e-01 & 2.0e+01 & 27.3 & 3.1e-03 & 1.8e-01 & 1.4e+00 & 5.6e-04\\
+GW Lup     & 1 &  7.49 &  9.9 & 8.3e-03 & 1.5e-01 & 7.8e+00 &  9.8 & 8.0e-03 & 1.7e-01 & 3.4e-01 & 1.3e-03\\
+HD 143006  & 1 &  3.95$^{*}$ & 10.1 & 4.7e-03 & 8.6e-02 & 7.5e+01 & 167.1 & 8.4e-04 & 1.8e-01 & -- & 1.5e-04\\
+HD 143006  & 2 &  7.22$^{*}$ & 10.1 & 4.8e-03 & 8.8e-02 & 3.4e+01 & 73.1 & 1.9e-03 & 1.0e+00 & -- & 1.9e-03\\
 \hline
@@
 \ref{tab-ring-model-limits}). Again these are conservative limits, with real
-values likely to be substantially smaller.
+values likely to be substantially smaller. {\bf UPDATE THESE VALUES AND RICCI
+NAME SINCE WE NOW USE DIFFERENT OPACITY MODEL.}
 
@@
 protoplanetary disk cannot be much narrower than about a pressure scale height
-$h_p(r)$. A dust ring produced by dust trapping in this pressure bump may become
+$h_p(r)$. {\bf [Perhaps we should add an appendix where we use the stability
+    criteria to rigorously put limits on the width and depth of the rings (even
+    a wide ring will, deep in the gaussian wings, become unstable), and
+    link this to a leakage of dust out of the trap.]}
+
+A dust ring produced by dust trapping in this pressure bump may become
 rather narrow, dependent on a variety of parameters, as discussed in
@@
 optimistically low disk temperature (e.g.~10 K) the pressure scale height at
-rings 1 and 2 are 2.4 au and 4.3 au, respectively, which correspond to fwhm
+rings 1 and 2 are 2.4 au and 4.3 au, respectively, which correspond to FWHM
 widths of 55 and 101 milliarcseconds, respectively (where we have multiplied by
@@
 $h_p$ as standard deviation width). Clearly the ALMA observations in band 6,
-with fwhm beam size of 35 mas, spatially resolve the pressure scale height.
+with FWHM beam size of 35 mas, spatially resolve the pressure scale height.
 This means that the ring separation will be spatially resolved by ALMA,
@@
 \appendix
+
+\section{Conventions on brightness}
+\label{sec-conventions}
+%
+The intensity of an interferometrically produced image at (sub-)millimeter
+wavelengths can be expressed in numerous ways. We will define our symbols here
+to avoid confusion. Expressed in CGS units, the intensity $I_\nu$ at frequency
+$\nu$ has dimension
+$\mathrm{erg}\,\mathrm{cm}^{-2}\,\mathrm{s}^{-1}\,\mathrm{Hz}^{-1}\,\mathrm{ster}^{-1}$.
+From an observational point of view it is convenient to express it instead in
+Jansky/beam. We shall write this as $I_\nu^{\mathrm{jpb}}$. The conversion is
+\begin{equation}
+  I_\nu = 10^{-23}\,\frac{\pi(\mathrm{au}/\mathrm{pc})^2 b_{\mathrm{as}}^2}{4\ln(2)}\frac{1}{b^2_{\mathrm{as}}}I_\nu^{\mathrm{jpb}}
+  = 3.755\times 10^{-13}\frac{1}{b^2_{\mathrm{as}}}I_\nu^{\mathrm{jpb}}
+\end{equation}
+where
+\begin{equation}\label{eq-def-average-beam}
+b_{\mathrm{as}}=\sqrt{b_{\mathrm{as}}^{\parallel}b_{\mathrm{as}}^{\bot}}
+\end{equation}
+is the position-angle-averaged beam full-width-at-half-maximum in units of
+arcseconds. Alternatively it can be expressed as brightness temperature
+$T_\nu^{\mathrm{b}}$ in Kelvin by solving the equation
+\begin{equation}
+I_\nu = B_\nu(T_\nu^{\mathrm{b}})
+\end{equation}
+where $B_\nu(T)=2h\nu^3c^{-2}/(\exp(h\nu/k_BT)-1)$ is the Planck function with $h$
+Planck's constant, $k_B$ Boltzmann's constant and $c$ the light speed. The
+relation between brightness temperature and intensity is non-linear, which is
+inconvenient. As an alternative to the brightness temperature (henceforth called
+``full brightness tempearture'') we therefore use the {\em linear} brightness
+temperature $I_\nu^{\mathrm{lb}}$, also expressed in Kelvin, defined as
+\begin{equation}
+I_\nu^{\mathrm{lb}} = \frac{c^2}{2\nu^2k_B}I_\nu
+  = 3.255\times 10^{36}\, \frac{1}{\nu^2}I_\nu
+\end{equation}
+The linear brightness temperature is thus a linear representation of the
+intensity, which is why we use the symbol $I_\nu^{\mathrm{lb}}$ instead of
+$T_\nu^{\mathrm{lb}}$. The full brightness temperature and the linear brightness
+temperature become equal in the Rayleigh-Jeans regime. However, at the
+wavelength we are working in (ALMA band 6 at $\lambda\simeq 0.13\,\mathrm{cm}$)
+and the temperatures we are dealing with (several Kelvins to tens of Kelvins) we
+are not in the Rayleigh-Jeans regime anymore, so the distinction between linear
+and non-linear brightness temperature becomes important.
+
+The conversion between the full brightness temperature $T_\nu^{\mathrm{b}}$ and
+the linear brightness temperature $I_\nu^{\mathrm{lb}}$ is given by
+\begin{equation}\label{eq-convert-lin-to-full-bright}
+\frac{h\nu}{k_BT_\nu^{\mathrm{b}}} = \ln\left(1+\frac{h\nu}{k_BI_\nu^{\mathrm{lb}}}\right)
+\end{equation}
+where in our case, with $\lambda=0.13\,\mathrm{cm}$, we have
+$h\nu/k_B=11\,\mathrm{K}$.
+
+The Planck function in units of Kelvin, to be used in conjunction with the
+brightness, is then
+\begin{equation}\label{eq-app-bnu-bright}
+B_\nu^{\mathrm{lb}}(T) = \frac{h\nu}{k_B}\frac{1}{e^{h\nu/k_BT}-1}
+\end{equation}
+so that at high optical depth the brightness saturates to
+$I_\nu^{\mathrm{lb}}=B_\nu^{\mathrm{lb}}(T)$. Throughout this paper we 
+express the intensity as $I_\nu^{\mathrm{lb}}$.
+
+\section{Gauss fitting procedure}
+\label{sec-gauss-fitting-procedure}
+%
+In this paper we study each dust ring individually, and try to understand it in
+terms of the trapping of dust in a pressure trap. In Appendix
+\ref{sec-steady-state-analytic-trap-model} we find that for a Gaussian pressure
+bump the solution to the radial dust mixing and drift problem is, to first
+approximation, also a Gaussian, albeit a narrower one. Fitting the dust trapping
+model to the data can therefore be done in two stages: first fitting a Gaussian
+radial profile to the observed rings, then interpreting these Gaussian fits
+through the dust trapping model. In this Appendix we detail the procedure we
+used to fit the radial intensity profiles of the rings with Gaussian profiles.
+
+{\bf Complete this section. Add an example of a corner plot (only one is
+presumably enough) and provide a more detailed table of error estimates.}
+
+The fitting was done in the
+image plane, in order to make it possible to fit to a limited radial range,
+i.e.\ selecting a single ring.
+
+\section{Gaussian mimicry: Detectability of mild optical depth effects}
+\label{sec-mimicry-flat-topped}
+%
+As was discussed in Section \ref{sec-saturation-flat-top-profiles}, the
+saturation effect of radiative transfer leads to Gaussian radial profiles of the
+rings to be flattened at the peak by the $1-e^{-\tau_\nu}$ self-absorption
+effect. The resulting flat-topped profile deviates from the optically thin
+approximation (which is proportional to $\tau_\nu$) by a factor
+$(1-e^{-\tau_\nu})/\tau_\nu$. This leads, for $\tau_\nu^{\mathrm{peak}}\gtrsim
+0.6$, to a substantial reduction near the peak. However, for
+$\tau_\nu^{\mathrm{peak}}\lesssim 1$, the {\em shape} of the profile stays
+remarkably close to a Gaussian profile, albeit one that is wider and weaker than
+the underlying $\tau_\nu(r)$. If the differences lie within the errorbars of the
+observations, there is no way to tell the difference.
+
+This `mimicry' (a non-Gaussian function posing as a Gaussian one) makes it hard
+to exclude, from looking at the shape of the profile (it being not
+`flat-topped'), that optical depth effects play a role. Given that much of our
+analysis of the rings in terms of dust trapping relies on the optically thin
+approximation (``what you see is what is there''), it is important to
+quantify this. This is the topic of this appendix.
+
+We do the following experiment: We start from a simple Gaussian function
+$g(x)$ given by
+\begin{equation}\label{eq-mimicry-underlying-gauss}
+g(x) = \exp\left(-\frac{1}{2}x^2\right)
+\end{equation}
+where $x=(r-r_0)/w_d$. From this we generate the optical depth profile
+\begin{equation}
+\tau_\nu(x) = \tau_\nu^{\mathrm{peak}}\,g(x)
+\end{equation}
+for a given value of $\tau_\nu^{\mathrm{peak}}$. We then compute the function
+\begin{equation}
+  f(x) \equiv \frac{I_\nu^{\mathrm{lb}}(x)}{B_\nu^{\mathrm{lb}}(T_d)}
+  = 1-e^{-\tau_\nu(x)}
+\end{equation}
+The results, for different values of $\tau_\nu^{\mathrm{peak}}$, are shown in
+the left panel of Fig.~\ref{fig-mimicry-profiles}. As one can see, the
+differences between $\tau_\nu(x)$ and $f(x)$ become quite substantial for
+$\tau_\nu^{\mathrm{peak}}\simeq 0.7$.
+
+\begin{figure*}
+\centerline{\includegraphics[width=0.95\textwidth]{saturation_example_combined.pdf}}
+\caption{\label{fig-mimicry-profiles}Demonstration of the Gaussian 'mimicry'
+  effect discussed in Appendix \ref{sec-mimicry-flat-topped}. Left panel:
+  Profiles before beam-convolution. Solid curves show the radiative transfer
+  result $f(x)=1-e^{-\tau_\nu(x)}$, dotted curves show the underlying
+  $\tau_\nu(x)$.  Middle panel is the same, but convolved with a beam that is
+  half the width of the Gaussian profile of $\tau_\nu(x)$. Right panel: Solid
+  lines are the same as for the middle panel, but the dashed lines are the best
+  fit Gaussian curves (also convolved with the beam).}
+\end{figure*}
+
+For both $\tau_\nu(x)$ and $f(x)$ we now compute the beam-convolved versions:
+\begin{equation}
+  \tau_\nu^{\mathrm{conv}}(x) = \frac{1}{\sqrt{2\pi}\sigma_{\mathrm{beam}}}\int_{-\infty}^{+\infty}
+  \tau_\nu(x')\,\exp\left(-\frac{(x-x')^2}{2\sigma_{\mathrm{beam}}^2}\right)dx'
+\end{equation}
+and likewise for $f(x)$, where $\sigma_{\mathrm{beam}}$ is the
+standard-deviation beam width in $x$-units. In this experiment the width of the
+underlying Gaussian `dust distribution' (Eq.~\ref{eq-mimicry-underlying-gauss})
+stays always fixed at $\sigma_{\mathrm{gauss}}=1$, and the beam size is
+expressed in relation to that. For the example of a beam that is half the width
+of the Gaussian profile of $\tau_\nu(x)$ (corresponding roughly to ring 1 of AS
+209, for instance) the results are shown in the middle panel of
+Fig.~\ref{fig-mimicry-profiles}. This panel shows that the beam convolution
+changes the results only by a little, at least for beams smaller than, or equal
+to, the original width of the dust distribution. Given that all the rings of
+this paper appear to be spatially resolved, we do not consider larger beams
+here.
+
+Now we try to find the best-fit Gaussian profile for $f(x)$. We do that by
+constructing an alternative Gaussian function $\tilde g(x)$
+\begin{equation}\label{eq-mimicry-mimicked-gauss}
+\tilde g(x) = \tilde a\exp\left(-\frac{x^2}{2\tilde\sigma^2}\right)
+\end{equation}
+where, in contrast to the original $g(x)$ of
+Eq.~(\ref{eq-mimicry-underlying-gauss}), the $\tilde a$ and $\tilde\sigma$ are
+free parameters to fit. We subsequently obtain $\tilde \tau_\nu(x)$ and $\tilde
+\tau_\nu^{\mathrm{conv}}(x)$ by the same procedure as above, then then use a
+minimization procedure to find the values of $\tilde a$ and $\tilde \sigma$ for
+which the difference to the function $f^{\mathrm{conv}}(x)$ is minimal.  We then
+plot the resulting $\tilde \tau_\nu^{\mathrm{conv}}(x)$ as dashed lines in the
+right panel of Fig.~\ref{fig-mimicry-profiles}. One can see that even for the
+case of $\tau_\nu^{\mathrm{peak}}=3.0$ the Gaussian fit mimicks the real curve
+remarkably (or perhaps disturbingly) well. Given the high signal to noise we
+have in our azimuthally-averaged radial profiles (see
+Fig.~\ref{fig-obs-gaussfits}), we can nevertheless exclude the case
+$\tau_\nu^{\mathrm{peak}}=3.0$. But it is much harder to exclude the case
+$\tau_\nu^{\mathrm{peak}}=1.2$, and impossible to exclude the case
+$\tau_\nu^{\mathrm{peak}}=0.7$.
+
+\begin{figure*}
+\centerline{\includegraphics[width=0.95\textwidth]{gausssatur_ratios.pdf}}
+\caption{\label{fig-mimicry-bestfitgauss-ratios}A quantitative analysis of
+  the mimicry effect of Appendix \ref{sec-mimicry-flat-topped}, scanning
+  the two parameters of the model: $(\sigma_{\mathrm{beam}},\tau_\nu^{\mathrm{peak}})$.
+  The four panels show the four derived parameters explained in the text.
+  Upper left shows the degree of overestimation of the width, upper right
+  the degree of underestimation of the dust mass, bottom left the underestimation
+  of the temperature, bottom right an estimate of the deviation of the Gauss fit
+  to the real curve (the closer to unity, the better the mimicry, and therefore
+  the worse the interpretation of the observations in terms of an optically thin
+  model).
+}
+\end{figure*}
+
+To quantify this a bit better, we have run the above models for a 2-D grid of
+parameters $(\sigma_{\mathrm{beam}},\tau_\nu^{\mathrm{peak}})$, and computed, for
+each pair of parameters, the following quantities:
+\begin{eqnarray}
+  {\cal W} &=& \tilde\sigma \\
+  {\cal M} &=& \tilde a \\
+  {\cal T} &=& f^{\mathrm{conv}}(x=0)\\
+  {\cal F} &=& \tilde \tau_\nu^{\mathrm{conv}}(0)/f^{\mathrm{conv}}(x=0)
+\end{eqnarray}
+They are plotted in Fig.~\ref{fig-mimicry-bestfitgauss-ratios}. The ${\cal W}$
+parameter tells us how much the mimicry will lead to an overestimation of the
+true width of the dust distribution. The ${\cal M}$ parameter tells us how much we
+would underestimate the dust mass by this effect. The ${\cal T}$ parameter tells
+us the degree to which we underestimate the true linear brightness temperature
+$B_\nu^{\mathrm{lb}}(T_d)$ (and by inference, the true dust temperature $T_d$)
+if we would interpret the observed linear brightness temperature as if it were
+optically thick (i.e.~by setting
+$B_\nu^{\mathrm{lb}}(T_d)=I_{\nu}^{\mathrm{lb,obs}}$). For the case of no
+convolution this becomes ${\cal T}=1-e^{\tau_\nu^{\mathrm{peak}}}$. Finally the
+${\cal F}$ parameter gives the ratio of the mimicking Gauss function to the real
+$f^{\mathrm{conv}}$ function, at $x=0$, as a representation of `how bad is the
+mimicking'. The closer ${\cal F}$ is to 1, the better is the mimicking, meaning
+the more difficult it is to distinguish the real from the mimicked curve.
+If ${\cal F}-1$ is less than the relative error of the observations at the
+peak of the curve, then they are indistinguishable.
+
+Thes results show that up to $\tau_\nu^{\mathrm{peak}}\lesssim 1$ the
+differences between the best-fit Gauss and the real curve are at most $\sim
+1\cdots 2$ \%, depending on the beam size. The beam size has most influence
+on ${\cal F}$, but less on the other parameters. That is because many of the
+effects of beam smearing are cancelled out because we know the beam size and
+can deconvolve. But the beam smearing does make a non-Gaussian profile more
+Gaussian-shaped, hence the stronger effect on ${\cal F}$.
+
+Although the mimicry effect is bad news for the interpretation of the observed
+radial ring profiles, the upper-left panel of
+Fig.~\ref{fig-mimicry-bestfitgauss-ratios} shows that the effect on the derived
+width of the dust ring remains modest. The fact that the widths of the rings in
+our sample all seem to be spatially resolved by at least 2 beam widths can
+therefore not be due to an unresolved dust ring being broadened by the optical
+depth effect. To broaden by a factor of 2 (i.e.~${\cal W}=2$) you would need
+an optical depth so large that the mimicry will clearly fail, i.e.\ we would
+easily recognize the deviations from a Gaussian shape.
+
+\section{Thermal emission from a thin dust layer with scattering}
+\label{sec-oned-radtrans-model}
+%
+So far we have assumed that only the absorption opacity
+$\kappa_\nu^{\mathrm{abs}}$, not the scattering opacity, matters. For optically
+thin dust layers this is indeed appropriate. But we have seen in Sections
+\ref{sec-opt-thin-analysis} and \ref{sec-optical-depth-effects} that the
+optical depth are not that low. Furthermore, the CO line extinction data
+of HD 163296 discussed by \paperdsharpisella{} suggest that the dust
+layer has an extinction optical depth close to unity. Even if the absorption
+optical depth is substantially below 1, the total extinction (absorption +
+scattering) can easily exceed unity, if the grains are of similar size to
+the wavelength. For $a\simeq \lambda/2\pi=0.13/2\pi = 0.02\,\mathrm{cm}$ the
+albedo of the grain can, in fact, be as high as 0.8 {\bf CHECK}.
+
+The inclusion of scattering complicates the radiative transfer equation
+enormously. Strictly speaking a full radiative transfer calculation, for
+instance with a Monte Carlo code, is necessary. However, in the spirit of
+this paper we wish to find a simple approximation to handle this without
+resorting to numerical simulation.
+
+We will outline here a simple two-stream radiative transfer approach to the
+problem. We will assume that the dust seen in the ALMA observations is located
+in a geometrically thin layer at the midplane, so we can use the 1-D slab
+geometry approach. We will assume that the scattering is isotropic. This
+may be a bad approximation, especially for $2\pi a\gg \lambda$. To reduce
+the impact of this approximation we replace the scattering opacity
+$\kappa_\nu^{\mathrm{scat}}$
+with
+\begin{equation}
+\kappa_\nu^{\mathrm{scat,eff}} = (1-g_\nu)\, \kappa_\nu^{\mathrm{scat}}
+\end{equation}
+where $g_\nu$ is the usual forward-scattering parameter (the expectation
+value of $\cos\theta$, where $\theta$ is the scattering angle). According
+to \citet{1978wpsr.book.....I}, this approximation works well in
+optically thick media.
+
+We will now follow the two-steam / moment method approach from
+\citet{1986rpa..book.....R} to derive the solution to the
+emission/absorption/scattering problem in this slab. The slab is put between
+$z=-\tfrac{1}{2}\Delta z$ and $z=+\tfrac{1}{2}\Delta z$ and we assume a constant
+density of dust between these two boundaries. The mean intensity $J_\nu(z)$
+of the radiation field then obeys the equation
+\begin{equation}
+\frac{1}{3}\frac{d^2J_\nu(\tau)}{d\tau_\nu^2} = \epsilon_\nu \big(J_\nu-B_\nu(T_d)\big)
+\end{equation}
+where
+\begin{equation}
+  \tau_\nu=\rho_d\,(\kappa_\nu^{\mathrm{abs}}+\kappa_\nu^{\mathrm{scat,eff}}) z
+  \equiv \rho_d\, \kappa_\nu^{\mathrm{tot}} z
+\end{equation}
+with $\rho_d$ being the dust density, and
+\begin{equation}
+\epsilon_\nu = \frac{\kappa_\nu^{\mathrm{abs}}}{\kappa_\nu^{\mathrm{abs}}+\kappa_\nu^{\mathrm{scat,eff}}}
+\end{equation}
+The boundary conditions at $z=\pm\tfrac{1}{2}\Delta z$ are
+\begin{equation}
+\frac{dJ_\nu}{d\tau_\nu} = \mp \sqrt{3}J_\nu
+\end{equation}
+This leads to the following solution:
+\begin{equation}
+  \frac{J_\nu(\tau_\nu)}{B_\nu(T_d)} =  1-b\,
+    \left(e^{-\sqrt{3\epsilon_\nu}\left(\tfrac{1}{2}\Delta\tau-\tau_\nu\right)}+e^{-\sqrt{3\epsilon_\nu}\left(\tfrac{1}{2}\Delta\tau+\tau_\nu\right)}\right)
+\end{equation}
+where $\Delta\tau = \rho_d\, \kappa_\nu^{\mathrm{tot}} \Delta z$ and $b$ is
+\begin{equation}
+b = \left[(1-\sqrt{\epsilon_\nu})e^{-\sqrt{3\epsilon_\nu}\Delta\tau} + 1 + \sqrt{\epsilon_\nu}\right]^{-1}
+\end{equation}
+Given this solution for the mean intensity $J_\nu(\tau_\nu)$ we can now numerically integrate
+the formal transfer equation along a single line of sight passing through the slab at an
+angle $\theta$:
+\begin{equation}\label{eq-fte-slab-with-scat}
+  \mu\frac{dI_\nu(\tau_\nu)}{d\tau_\nu} = \epsilon_\nu B_\nu(T_d) + (1-\epsilon_\nu) J_\nu(\tau_\nu)
+  -I_\nu(\tau_\nu)
+\end{equation}
+where $\mu=\cos\theta$. We start at $\tau_\nu=-\tfrac{1}{2}\Delta \tau$ with
+$I_\nu=0$ and integrate to $\tau_\nu=+\tfrac{1}{2}\Delta \tau$. The resulting
+$I_\nu^{\mathrm{out}}=I_\nu(\tfrac{1}{2}\Delta \tau)$ is the intensity that is
+observed by the telescope. An approximation for $I_\nu^{\mathrm{out}}$ which
+works well to within a few percent is the following modifed version of the
+Eddington-Barbier approximation:
+\begin{equation}\label{eq-modif-eddington-barbier}
+I_\nu^{\mathrm{out}} \simeq \left(1-e^{-\Delta\tau/\mu}\right)\,S_\nu\left((\tfrac{1}{2}\Delta\tau-\tau_\nu)/\mu=2/3\right)
+\end{equation}
+where
+\begin{equation}
+S_\nu(\tau_\nu) = \epsilon_\nu B_\nu(T_d) + (1-\epsilon_\nu) J_\nu(\tau_\nu)
+\end{equation}
+is the source function. The results are shown in Fig.~\ref{fig-scatter-parscan}.
+
+For small optical depths ($\Delta\tau\ll 1$) the role of scattering vanishes,
+and the solution approaches: $I_\nu^{\mathrm{out}} \rightarrow \epsilon_\nu\Delta\tau
+B_\nu(T_d)$.  This is the same limiting solution as when
+$\kappa_\nu^{\mathrm{scat}}$ is set to zero but $\kappa_\nu^{\mathrm{abs}}$ is
+kept the same. For high optical depth the outcoming intensity does not saturate
+to the Planck function, but a bit below, if the albedo is non-zero. This is the
+well-known effect that scattering makes objects appear cooler than they really
+are.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.47\textwidth]{scatter_parscan.pdf}}
+\caption{\label{fig-scatter-parscan}The intensity $I_\nu$, in units of the
+  Planck function, emerging from a slab seen face on, with total optical depth
+  $\Delta\tau$, a constant temperature, and an albedo of
+  $\eta_\nu=1-\epsilon_\nu$. See Appendix \ref{sec-oned-radtrans-model} for
+  details. The solid lines are the results of numerical integration of
+  Eq.~(\ref{eq-fte-slab-with-scat}). The dotted lines are the result of the
+  modified Eddington-Barbier approximation
+  (Eq.~\ref{eq-modif-eddington-barbier}).}
+\end{figure}
 
@@
 
-\section{Gauss fitting procedure}
-\label{sec-gauss-fitting}
-%
-In this paper we study each dust ring individually, and try to understand it in
-terms of the trapping of dust in a pressure trap. In Appendix
-\ref{sec-steady-state-analytic-trap-model} we find that for a Gaussian pressure
-bump the solution to the radial dust mixing and drift problem is, to first
-approximation, also a Gaussian, albeit a narrower one. Fitting the dust trapping
-model to the data can therefore be done in two stages: first fitting a Gaussian
-radial profile to the observed rings, then interpreting these Gaussian fits
-through the dust trapping model. In this appendix
-
 \section{Computing the dust mass trapped in the rings}
@@
 %
-We can compute the dust mass trapped in each of the rings by integrating over
-the measured brightness temperature profiles. However, the profiles do not go
-down to zero far way from the peak, and so it is not well defined what belongs
-to the ring and what not. Instead, we will compute the dust mass from the
-Gaussian profile that is fitted to the ring emission. This will slightly
-underestimate the dust mass, because the excess emission in the flanks of the
-profiles (see Fig.~\ref{fig-obs-gaussfits}) will not be included. But there are
-far more serious issues to worry about, such as our limited knowledge of the
-disk temperature, the uncertainties of the dust opacities, as well as optical
-depth effects. So we will estimate the mass of the best fitting Gaussian model
-only.
-
-We will assume that the intrinsic dust temperature is as given in Table
-\ref{tab-gauss-params} (column $T_{\mathrm{d}}$), which was computed based on
-a simple flaring angle irradiation model. We consider the Gaussian profile with
-the deconvolved width, i.e.\ $w_d$ (not $\sigma$). The peak linear brightness
-temperature is taken to be
-\begin{equation}
-A_{\mathrm{deconv}} = \frac{\sigma}{w_d}\,A
-\end{equation}
-where $A$, $\sigma$ and $w_d$ are from Table \ref{tab-gauss-params}. We do not
-account for geometric terms in the deconvolution, which is justified as long as
-$\sigma\ll r$. The deconvolved Gaussian brightness profile is then
-\begin{equation}\label{eq-tbrlin-deconvolved}
-T_{\mathrm{br,lin,deconv}}(r) = A_{\mathrm{deconv}}\,\exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
-\end{equation}
-
-Next we have to worry about optical depth effects. From Table
-\ref{tab-gauss-params} one can see that for most rings 
-$A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{d,br,lin}}$.
-However, for the rings of AS 209 the ratios
-$A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}$ are around 0.5, so optical depth
-effects start to play a role. Given a dust surface density profile
-$\Sigma_{\mathrm{d}}(r)$ as a function of radial coordinate $r$, and
-given an absorption opacity $\kappa_{\mathrm{a}}$ at the wavelength we
-are measuring, the absorption optical depth $\tau_{\mathrm{a}}(r)$ is
-\begin{equation}\label{eq-tau-afo-sig-kap}
-\tau_{\mathrm{a}}(r) = \Sigma_{\mathrm{d}}(r)\kappa_{\mathrm{a}}
-\end{equation}
-We assume that scattering does not play a role. The emission from this
-dusty ring is then, expressed as linear brightness temperature:
-\begin{equation}
-T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{d,br,lin}}
-\end{equation}
-If $\tau_{\mathrm{a}}(r)\ll 1$ for all $r$ this reduces to
-\begin{equation}\label{eq-int-afo-tau-bnu}
-T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{d,br,lin}}
-\end{equation}
-But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}
-=\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{d}}))$ can be as large as
-$\sim 0.5$, this condition is not always fulfilled.  On the other hand, these
-optical depth effects cannot be too strong, because then the Gaussian shape
-would flatten off at the top, which is not what we see in the observations. We
-will therefore continue our computation without accounting for these (mild)
-optical depth effects. This will lead to a slight underestimation of the
-mass. The reader can estimate the magnitude of this underestimation by comparing
-$(1-\exp(-\tau_{\mathrm{a}}))$ to $\tau_{\mathrm{a}}$.
-
-Using Eq.~(\ref{eq-int-afo-tau-bnu}) we can, from the deconvolved
-$T_{\mathrm{br,lin,deconv}}(r)$ (Eq.~\ref{eq-tbrlin-deconvolved}), solve for
-$\tau_{\mathrm{a}}(r)$ and obtain
-\begin{equation}
-  \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{d,br,lin}}}\,
-  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
-\end{equation}
-leading with Eq.~(\ref{eq-tau-afo-sig-kap}) to
-\begin{equation}
-  \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\,
-  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
-\end{equation}
-Integrating over $r$, ignoring geometric terms leads to
-\begin{equation}\label{eq-ringmass-from-obs}
-  \begin{split}
-    M_{\mathrm{d}} &\simeq 2\pi r_0\int_{r_{\mathrm{in}}}^{r_{\mathrm{out}}}\Sigma_{\mathrm{d}}(r)dr\\
-    &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\\
-    &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}
-  \end{split}
-\end{equation}
-
-\begin{figure}
-\centerline{\includegraphics[width=0.47\textwidth]{kappas_alma_band_6.pdf}}
-\caption{\label{fig-opacities}The dust opacity at $\lambda=0.13\,\mathrm{cm}$
-  as a function of the grain radius $a$. Plotted are the absorption opacity of
-  \citet{2010A&A...512A..15R}, the corresponding scattering opacity, as well as
-  the analytic absorption opacity of \citet{1997MNRAS.291..121I}.}
-\end{figure}
-
-As the final step we have to make an estimate of the absorption opacity
-$\kappa_{\mathrm{a}}$. This is the biggest uncertainty because it is not well
-understood what the composition of the dust is, what the grain sizes are and
-what the effect of the geometry is on the opacity. We will use two estimates: on
-the one hand we will employ the opacity used by \citet{2010A&A...512A..15R},
-which is a mixture of astronomical silicate and amorphous carbon (see
-Fig.~\ref{fig-opacities}). On the other hand we will use the dummy opacity model
-of \citet{1997MNRAS.291..121I}. These two opacities serve as two extremes that
-bracket the likely true value. It is known that at long wavelengths the
-opacities computed with the Mie procedure from laboratory measured optical
-constants are likely to be underestimates by up to a factor of 10
-\citep{2017A&A...600A.123D}. The more realistic opacities of
-\citep{2017A&A...600A.123D} are in between our two bracketing opacities.  For
-the mass estimate we use the opacity at $\lambda=0.13\,\mathrm{cm}$ for small
-grains ($a=1\mu\mathrm{m}$). For the Ricci opacity this is
-$\kappa_{\mathrm{a,ri}}=1.70\,\mathrm{cm}^2/\mathrm{g}$, while for the Ivezic
-opacity this is $\kappa_{\mathrm{a,ri}}=31.3\,\mathrm{cm}^2/\mathrm{g}$. The
-masses estimated from Eq.~(\ref{eq-ringmass-from-obs}) using these two
-bracketing opacities are listed in Table \ref{fig-obs-gaussfits}. Note that the
-opacity of \citet{2010A&A...512A..15R} has a peak around a grain size of 2
-$\mathrm{mm}$. That peak value happens to be very close to the Ivezic value of
-$31.3\,\mathrm{cm}^2/\mathrm{g}$. We can thus also regard the Ivezic mass
-estimate to be the mass if all grains have a radius of 2 mm.
-
+
+{\bf THIS APPENDIX HAS TO BE COMPLETELY REWRITTEN: FOCUS ON OPACITY
+  UNCERTAINTY INSTEAD OF COMPUTATION OF DUST MASS, BECAUSE THE LATTER IS
+NOW IN THE MAIN TEXT.}
+% 
+% We can compute the dust mass trapped in each of the rings by integrating over
+% the measured brightness temperature profiles. However, the profiles do not go
+% down to zero far way from the peak, and so it is not well defined what belongs
+% to the ring and what not. Instead, we will compute the dust mass from the
+% Gaussian profile that is fitted to the ring emission. This will slightly
+% underestimate the dust mass, because the excess emission in the flanks of the
+% profiles (see Fig.~\ref{fig-obs-gaussfits}) will not be included. But there are
+% far more serious issues to worry about, such as our limited knowledge of the
+% disk temperature, the uncertainties of the dust opacities, as well as optical
+% depth effects. So we will estimate the mass of the best fitting Gaussian model
+% only.
+% 
+% We will assume that the intrinsic dust temperature is as given in Table
+% \ref{tab-gauss-params} (column $T_{\mathrm{d}}$), which was computed based on
+% a simple flaring angle irradiation model. We consider the Gaussian profile with
+% the deconvolved width, i.e.\ $w_d$ (not $\sigma$). The peak linear brightness
+% temperature is taken to be
+% \begin{equation}
+% A_{\mathrm{deconv}} = \frac{\sigma}{w_d}\,A
+% \end{equation}
+% where $A$, $\sigma$ and $w_d$ are from Table \ref{tab-gauss-params}. We do not
+% account for geometric terms in the deconvolution, which is justified as long as
+% $\sigma\ll r$. The deconvolved Gaussian brightness profile is then
+% \begin{equation}\label{eq-tbrlin-deconvolved}
+% T_{\mathrm{br,lin,deconv}}(r) = A_{\mathrm{deconv}}\,\exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+% \end{equation}
+% 
+% Next we have to worry about optical depth effects. From Table
+% \ref{tab-gauss-params} one can see that for most rings 
+% $A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{d,br,lin}}$.
+% However, for the rings of AS 209 the ratios
+% $A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}$ are around 0.5, so optical depth
+% effects start to play a role. Given a dust surface density profile
+% $\Sigma_{\mathrm{d}}(r)$ as a function of radial coordinate $r$, and
+% given an absorption opacity $\kappa_{\mathrm{a}}$ at the wavelength we
+% are measuring, the absorption optical depth $\tau_{\mathrm{a}}(r)$ is
+% \begin{equation}\label{eq-tau-afo-sig-kap}
+% \tau_{\mathrm{a}}(r) = \Sigma_{\mathrm{d}}(r)\kappa_{\mathrm{a}}
+% \end{equation}
+% We assume that scattering does not play a role. The emission from this
+% dusty ring is then, expressed as linear brightness temperature:
+% \begin{equation}
+% T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{d,br,lin}}
+% \end{equation}
+% If $\tau_{\mathrm{a}}(r)\ll 1$ for all $r$ this reduces to
+% \begin{equation}\label{eq-int-afo-tau-bnu}
+% T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{d,br,lin}}
+% \end{equation}
+% But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{d,br,lin}}
+% =\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{d}}))$ can be as large as
+% $\sim 0.5$, this condition is not always fulfilled.  On the other hand, these
+% optical depth effects cannot be too strong, because then the Gaussian shape
+% would flatten off at the top, which is not what we see in the observations. We
+% will therefore continue our computation without accounting for these (mild)
+% optical depth effects. This will lead to a slight underestimation of the
+% mass. The reader can estimate the magnitude of this underestimation by comparing
+% $(1-\exp(-\tau_{\mathrm{a}}))$ to $\tau_{\mathrm{a}}$.
+% 
+% Using Eq.~(\ref{eq-int-afo-tau-bnu}) we can, from the deconvolved
+% $T_{\mathrm{br,lin,deconv}}(r)$ (Eq.~\ref{eq-tbrlin-deconvolved}), solve for
+% $\tau_{\mathrm{a}}(r)$ and obtain
+% \begin{equation}
+%   \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{d,br,lin}}}\,
+%   \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+% \end{equation}
+% leading with Eq.~(\ref{eq-tau-afo-sig-kap}) to
+% \begin{equation}
+%   \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\,
+%   \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
+% \end{equation}
+% Integrating over $r$, ignoring geometric terms leads to
+% \begin{equation}\label{eq-ringmass-from-obs}
+%   \begin{split}
+%     M_{\mathrm{d}} &\simeq 2\pi r_0\int_{r_{\mathrm{in}}}^{r_{\mathrm{out}}}\Sigma_{\mathrm{d}}(r)dr\\
+%     &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}\\
+%     &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{d,br,lin}}}
+%   \end{split}
+% \end{equation}
+% 
+% \begin{figure}
+% \centerline{\includegraphics[width=0.47\textwidth]{kappas_alma_band_6.pdf}}
+% \caption{\label{fig-opacities}The dust opacity at $\lambda=0.13\,\mathrm{cm}$
+%   as a function of the grain radius $a$. Plotted are the absorption opacity of
+%   \citet{2010A&A...512A..15R}, the corresponding scattering opacity, as well as
+%   the analytic absorption opacity of \citet{1997MNRAS.291..121I}.}
+% \end{figure}
+% 
+% As the final step we have to make an estimate of the absorption opacity
+% $\kappa_{\mathrm{a}}$. This is the biggest uncertainty because it is not well
+% understood what the composition of the dust is, what the grain sizes are and
+% what the effect of the geometry is on the opacity. We will use two estimates: on
+% the one hand we will employ the opacity used by \citet{2010A&A...512A..15R},
+% which is a mixture of astronomical silicate and amorphous carbon (see
+% Fig.~\ref{fig-opacities}). On the other hand we will use the dummy opacity model
+% of \citet{1997MNRAS.291..121I}. These two opacities serve as two extremes that
+% bracket the likely true value. It is known that at long wavelengths the
+% opacities computed with the Mie procedure from laboratory measured optical
+% constants are likely to be underestimates by up to a factor of 10
+% \citep{2017A&A...600A.123D}. The more realistic opacities of
+% \citep{2017A&A...600A.123D} are in between our two bracketing opacities.  For
+% the mass estimate we use the opacity at $\lambda=0.13\,\mathrm{cm}$ for small
+% grains ($a=1\mu\mathrm{m}$). For the Ricci opacity this is
+% $\kappa_{\mathrm{a,ri}}=1.70\,\mathrm{cm}^2/\mathrm{g}$, while for the Ivezic
+% opacity this is $\kappa_{\mathrm{a,ri}}=31.3\,\mathrm{cm}^2/\mathrm{g}$. The
+% masses estimated from Eq.~(\ref{eq-ringmass-from-obs}) using these two
+% bracketing opacities are listed in Table \ref{fig-obs-gaussfits}. Note that the
+% opacity of \citet{2010A&A...512A..15R} has a peak around a grain size of 2
+% $\mathrm{mm}$. That peak value happens to be very close to the Ivezic value of
+% $31.3\,\mathrm{cm}^2/\mathrm{g}$. We can thus also regard the Ivezic mass
+% estimate to be the mass if all grains have a radius of 2 mm.
+% 
 \end{document}

### 2018-08-13 projects/dullemond/paper_dusttraps/ms.tex
Slight update of dust trapping paper.

@@
   data, \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
-  for HD 163296, and \paperdsharphuangrings{} for the rest.}
+  for HD 163296, and \paperdsharphuangrings{} for the rest.
+{\bf TODO: y-labels}}
 \end{figure*}
@@
 
-{\bf [ADD APPENDIX WITH DETAILS OF FITS AND ERROR ESTIMATES]}
-
 \begin{figure*}
@@
   FWHM beam size of the observations. The grey band around the blue curve shows
-  the estimated uncertainty of the data.}
+  the estimated uncertainty of the data. {\bf TODO: y-labels}}
 \end{figure*}
@@
     for the two sources with two partly overlapping rings: HD 163296 and
-    HD 143006.}
+    HD 143006. {\bf TODO: y-labels}}
 \end{figure*}
@@
 where $b$ is the beam width expressed as standard deviation in units of
-$\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
-$b_{\mathrm{fwhm,as}}$ is the beam full-width at half-maximum (FWHM) in units of
-arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
-parsec). In using Eq.~(\ref{eq-simple-deconvolve-gauss}) we have ignored the
-elliptical shape of the beam as well as the inclination of the disk.
+$\mathrm{au}$. It is computed from $b_{\mathrm{fwhm,as}}$ (the beam full-width
+at half-maximum (FWHM) in units of arcseconds) through the equation
+$b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where $d_{\mathrm{pc}}$ is the
+distance to the source in units of parsec. In using
+Eq.~(\ref{eq-simple-deconvolve-gauss}) we have ignored the elliptical shape of
+the beam as well as the inclination of the disk.
 
@@
 
-The width of the dust rings
-in units of the local pressure scale height $h_p$ (see Table
-\ref{tab-gauss-params}) can only be roughly estimated, because we do not know
-the disk midplane temperature. Our estimate of the midplane disk temperature
-is made using the following simple irradiated flaring disk recipe:
+The width of the dust rings in units of the local pressure scale height $h_p$
+(see Table \ref{tab-gauss-params}) can only be roughly estimated, because we do
+not know $h_p$ well enough due to our lack of information about the disk midplane
+temperature. Our estimate of the midplane disk temperature is made using the
+following simple irradiated flaring disk recipe:
 \begin{equation}\label{eq-disk-temperature-model}
@@
 Lup and the ring 1 of HD 143006), but the data are inconsistent with radially
-unresolved rings. For AS 209 This is confirmed by more detailed analysis in the
+unresolved rings. For AS 209 this is confirmed by more detailed analysis in the
 uv-plane by \paperdsharpguzmann{}. {\bf CHECK IF HUANG DOES THIS, TOO.}
@@
 model we can derive the total dust mass trapped in the ring:
-\begin{equation}
+\begin{equation}\label{eq-dust-mass-estimate}
   M_d = \int_0^\infty 2\pi r \Sigma_d(r) dr \simeq
@@
 estimated dust surface density $\Sigma_d$ and dust mass $M_d$ derived from the
-optically thin assumption ( Eqs.~\ref{eq-optthin-conversion-linbright-sigmad}
-and {\bf XXX}, respectively) are only mildly affected by the optical depth
-effects.
+optically thin assumption (Eqs.~\ref{eq-optthin-conversion-linbright-sigmad} and
+\ref{eq-dust-mass-estimate}, respectively) are only mildly affected by the
+optical depth effects.
 
@@
 maps could thus be compatible with the sub-unity absorption optical depth we
-derived in out Gaussian fitting procedure (see Table \ref{tab-gauss-params}).
+derived in our Gaussian fitting procedure (see Table \ref{tab-gauss-params}).
 
@@
 %
-Given the well separated narrow rings in the above sources, we can analyze these
-rings on a one-by-one basis. The hypothesis we are going to test is that the
+The hypothesis we are now going to test is that the
 rings are caused by dust trapping in axisymmetric pressure bumps. For simplicity
@@
 
-{\bf Complete this section. Add an example of a corner plot (only one is
-presumably enough) and provide a more detailed table of error estimates.}
-
-The fitting was done in the
-image plane, in order to make it possible to fit to a limited radial range,
-i.e.\ selecting a single ring.
+The radial intensity profiles (expressed as linear brightness temperature) were
+extracted from the images using a procedure similar to that described by
+\paperdsharphuangrings{}. This procedure involves the fitting of an ellipse to
+describe the inclined ring shape, the deprojection into a ring, and the
+averaging of the intensity along the ring. This averaging procedure enhances the
+signal-to-noise ratio considerably, by a factor $\sqrt{N}$, where $N$ is the
+number of beams that fit along the ring. We estimate the intrinsic noise simply
+by computing the standard deviation along the ring. The resulting averaged
+radial intensity profile $I_\nu^{\mathrm{lb}}(r)$ thus obtains also an error
+estimate $\varepsilon(r)$, which is typically of the order of $\sim$1\% of
+the peak intensity.
+
+The rings display themselves as bumps in $I_\nu^{\mathrm{lb}}(r)$. We choose by eye
+a radial domain around the bump where we believe a Gaussian description is justified.
+The inner and outer radii of this domain are listed in Table \ref{tab-gauss-params}.
+By choosing this domain we can select a specific ring to fit, which is not possible
+when doing the fitting procedure in the uv-plane.
+
+We now fit a Gaussian profile to this bump
+\begin{equation}
+I_\nu^{\mathrm{lb,gauss}}(r) = A\exp\left(-\frac{(r-r_0)^2}{2\sigma^2}\right)
+\end{equation}
+We use the code {\small\sf emcee} \citep{2013PASP..125..306F} to perform a
+Markov Chain Monte Carlo procedure to find the set of parameters
+$(A,r_0,\sigma)$ which have the highest likelihood. Given that the radial
+sampling of $I_\nu^{\mathrm{lb,gauss}}(r)$ is about $N\simeq 70$ points per beam, we
+have to multiply the error estimate of the datapoints by $\sqrt{N}$ before
+feeding it into {\small\sf emcee}, otherwise the MCMC procedure would treat each
+sampling point as an independent measuring point, which is not the case.
+
+We show the corner plot of the fitting of ring 1 of AS 209 in
+Fig.~\ref{fig-corner-gaussfit-as209-1}. The plots for the other rings are
+similar, and are provided in the online material. {\bf TODO: PUT OTHER PLOTS
+IN ONLINE MATERIAL.}
+
+\begin{figure}
+\centerline{\includegraphics[width=0.45\textwidth]{AS_209_ring1_cornerplot_gaussfit.pdf}}
+\caption{\label{fig-corner-gaussfit-as209-1}The correletion plot for the Gauss
+  fitting of ring 1 of AS 209, from the MCMC procedure described in Appendix
+  \ref{sec-gauss-fitting-procedure}.}
+\end{figure}
+
+The most likely parameter values and their error estimates are given in
+Table \ref{tab-gauss-fit-errors}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|cc|ccc|}
+\hline
+\hline
+Source & Ring & $A$ & $r_0$ & $\sigma$ \\
+\hline
+AS 209     & 1 & $  3.637_{-0.047}^{+0.048}$ & $ 74.129_{-0.066}^{+0.063}$ & $  3.885_{-0.090}^{+0.098}$  \\
+AS 209     & 2 & $  2.949_{-0.031}^{+0.030}$ & $120.309_{-0.068}^{+0.065}$ & $  4.651_{-0.111}^{+0.115}$  \\
+Elias 24   & 1 & $  5.214_{-0.068}^{+0.066}$ & $ 76.851_{-0.092}^{+0.088}$ & $  5.104_{-0.172}^{+0.186}$  \\
+HD 163296  & 1 & $  8.475_{-0.056}^{+0.056}$ & $ 67.403_{-0.055}^{+0.057}$ & $  7.370_{-0.098}^{+0.098}$  \\
+HD 163296  & 2 & $  4.918_{-0.026}^{+0.027}$ & $ 99.862_{-0.061}^{+0.059}$ & $  5.741_{-0.116}^{+0.119}$  \\
+GW Lup     & 1 & $  1.467_{-0.052}^{+0.055}$ & $ 85.178_{-0.259}^{+0.316}$ & $  4.882_{-0.402}^{+0.530}$  \\
+HD 143006  & 1 & $  3.202_{-0.084}^{+0.082}$ & $ 41.080_{-0.169}^{+0.190}$ & $  5.133_{-0.320}^{+0.373}$  \\
+HD 143006  & 2 & $  2.473_{-0.043}^{+0.044}$ & $ 64.831_{-0.248}^{+0.230}$ & $  8.029_{-0.501}^{+0.593}$  \\
+\hline
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{tab-gauss-fit-errors}The Gaussian fit values with their error
+  estimates from the MCMC procedure described in Appendix \ref{sec-gauss-fitting-procedure}.}
+\end{table}
 
@@
 
-Thes results show that up to $\tau_\nu^{\mathrm{peak}}\lesssim 1$ the
+These results show that up to $\tau_\nu^{\mathrm{peak}}\lesssim 1$ the
 differences between the best-fit Gauss and the real curve are at most $\sim

### 2018-08-16 projects/huang/rings/continuumanalysis3.py
updated radial profiles

@@ -127,6 +127,24 @@ class Continuum:
 
         return radial_profile
 
+    def profilemin(self,radialbins): 
+        """
+        Parameters
+        ==========
+        """
+        rbins = radialbins
+        rwidth = (rbins[1]-rbins[0])
+
+        radial_profile = np.zeros( len(rbins) )
+
+
+        for i in range(len(rbins)):
+            annulus_intensities = self.image[(self.r>=(rbins[i]-0.5*rwidth)) & (self.r<(rbins[i]+0.5*rwidth))]
+
+            radial_profile[i] = np.min(annulus_intensities)
+
+        return radial_profile
+
     def azunwrap(self,radialbins, tbins = -179.+2*np.arange(180), yaxis = 'intensity', theta_exclusion = np.array([])): 
         """
         Parameters
@@ -350,6 +368,7 @@ class Continuum:
             else:
                 ticks = np.arange(-np.floor(size)/2., np.floor(size)/2.+0.001, 0.5)
 
+
             ax.tick_params(axis = "both", direction = 'in',which = 'major', colors = 'white', length = 3)
             ax.set_xticks(ticks)
             ax.set_yticks(ticks)
@@ -359,9 +378,6 @@ class Continuum:
             for ytick in ax.get_yticklabels():
                 ytick.set_color('black')
 
-            #plt.setp(ax.get_xticklabels())
-
-            #plt.setp(ax.get_yticklabels())
 
 
         plt.gca().invert_xaxis() #since RA increases right to left 

### 2018-08-18 projects/dullemond/paper_dusttraps/ms.tex
Updated dust trapping paper: new radial profiles.

@@
   with the width of the ALMA beam. For a detailed description of these
-  data, \paperdsharpguzmann{} for AS 209, \paperdsharpisella{}
-  for HD 163296, and \paperdsharphuangrings{} for the rest.
-{\bf TODO: y-labels}}
+  data, see \paperdsharphuangrings{}.}
 \end{figure*}
@@
   FWHM beam size of the observations. The grey band around the blue curve shows
-  the estimated uncertainty of the data. {\bf TODO: y-labels}}
+  the estimated uncertainty of the data.}
 \end{figure*}
@@
     for the two sources with two partly overlapping rings: HD 163296 and
-    HD 143006. {\bf TODO: y-labels}}
+    HD 143006.}
 \end{figure*}
@@
 % Note: Added A_dec
-AS 209     & 1 & 37 &  69 --  79 &   3.64 &   4.18 &  74.1 &  75.4 &   3.87 &   3.38 &  15.8 &  10.9 &   0.6 &   0.48 &  28.5\\
-AS 209     & 2 & 37 & 115 -- 125 &   2.95 &   3.23 & 120.3 &  90.3 &   4.64 &   4.23 &  12.4 &   7.7 &   0.4 &   0.55 &  60.7\\
-Elias 24   & 1 & 35 &  72 --  82 &   5.22 &   5.69 &  76.9 &  88.0 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &   0.40 &  33.4\\
-HD 163296  & 1 & 43 &  60 --  76 &   8.48 &   8.76 &  67.4 & 171.7 &   7.36 &   7.13 &  30.8 &  25.6 &   1.6 &   0.42 &  43.8\\
-HD 163296  & 2 & 43 &  94 -- 104 &   4.92 &   5.20 &  99.9 & 133.6 &   5.73 &   5.42 &  25.3 &  20.2 &   0.8 &   0.30 &  38.0\\
-GW Lup     & 1 & 44 &  79 --  89 &   1.48 &   1.87 &  85.1 &  72.1 &   4.74 &   3.76 &  10.3 &   5.7 &   0.5 &   0.40 &  34.1\\
-HD 143006  & 1 & 45 &  35 --  45 &   3.21 &   4.11 &  41.0 &  72.2 &   5.06 &   3.95 &  27.2 &  22.0 &   1.9 &   0.21 &  10.0\\
-HD 143006  & 2 & 45 &  59 --  72 &   2.48 &   2.71 &  64.8 & 112.5 &   7.88 &   7.22 &  21.6 &  16.6 &   2.0 &   0.18 &  21.6\\
+AS 209     & 1 & 37 &  69 --  79 &   3.66 &   4.17 &  74.1 &  77.6 &   3.99 &   3.51 &  15.8 &  10.9 &   0.6 &   0.48 &  29.2\\
+AS 209     & 2 & 37 & 115 -- 125 &   2.98 &   3.26 & 120.4 &  91.9 &   4.72 &   4.32 &  12.4 &   7.7 &   0.4 &   0.55 &  62.3\\
+Elias 24   & 1 & 35 &  72 --  82 &   5.50 &   6.03 &  76.7 &  85.3 &   4.92 &   4.49 &  22.3 &  17.2 &   0.6 &   0.43 &  34.1\\
+HD 163296  & 1 & 43 &  60 --  76 &   8.75 &   9.08 &  67.7 & 161.4 &   6.92 &   6.67 &  30.8 &  25.6 &   1.5 &   0.44 &  43.0\\
+HD 163296  & 2 & 43 &  94 -- 104 &   5.18 &   5.54 & 100.0 & 120.3 &   5.16 &   4.82 &  25.3 &  20.2 &   0.7 &   0.32 &  36.6\\
+GW Lup     & 1 & 44 &  79 --  89 &   1.29 &   1.49 &  85.5 &  87.7 &   5.77 &   4.99 &  10.2 &   5.7 &   0.7 &   0.30 &  33.5\\
+HD 143006  & 1 & 45 &  35 --  45 &   3.31 &   4.22 &  41.0 &  72.6 &   5.09 &   3.99 &  27.2 &  22.0 &   1.9 &   0.21 &  10.3\\
+HD 143006  & 2 & 45 &  59 --  72 &   2.58 &   2.80 &  65.2 & 114.3 &   8.01 &   7.37 &  21.6 &  16.5 &   2.0 &   0.19 &  22.9\\
 \hline
@@
 dust-to-gas ratio within the pressure bump stays constant. Only in the very
-wings of the Gaussian pressure bump is dust trapping significant, but this
+wings of the Gaussian pressure bump is dust trapping then significant, but this
 effect can only be studied using the more accurate non-gaussian solution of
@@
 Eq.~(\ref{eq-analytic-sol-radial-trapping-summary}) has only two parameters:
-$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. Both can be fairly directly
-measured from the observations, though $\Sigma_{\mathrm{d0}}$ will depend on the
-opacity, and hence on the unknown grain size (see discussion in Appendix
-\ref{sec-computing-dust-mass}). The values of $w_{\mathrm{d}}$ for
-the rings in our sample can be directly taken from Table \ref{tab-gauss-params}.
+$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. As we have shown in Sections
+\ref{sec-gauss-fits}, \ref{sec-opt-thin-analysis} and
+\ref{sec-optical-depth-effects}, both parameter can be extracted from the
+observations. The main uncertainty lies in $\Sigma_{\mathrm{d0}}$, due to the
+uncertainty in the dust opacity. The values of $w_{\mathrm{d}}$ for the rings in
+our sample can be directly taken from Table \ref{tab-gauss-params}.
 
@@
 \end{equation}
-Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
-Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
-weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
-$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
+Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed. Here
+$\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
+$G$ is the gravitational constant and $\Sigma_{\mathrm{g}}$ the gas
 surface density. Taking the disk midplane
@@
 % Note: New table with new beam size estimates 2018.08.06
-AS 209     & 1 &  5.59 & 19.6 & 1.1e-02 & 2.0e-01 & 1.6e+01 & 15.7 & 3.9e-03 & 3.1e-02 & 5.7e-01 & 1.2e-04\\
-AS 209     & 2 & 10.25 & 19.6 & 1.2e-02 & 2.3e-01 & 6.9e+00 &  5.8 & 9.1e-03 & 4.9e-02 & 2.1e-01 & 4.4e-04\\
-Elias 24   & 1 &  7.26 & 17.1 & 9.7e-03 & 1.8e-01 & 1.8e+01 & 18.8 & 3.6e-03 & 8.0e-02 & 7.0e-01 & 2.9e-04\\
-HD 163296  & 1 &  7.13$^{*}$ & 13.8 & 1.1e-02 & 1.9e-01 & 4.1e+01 & 40.0 & 1.5e-03 & 3.6e-01 & -- & 5.6e-04\\
-HD 163296  & 2 &  7.07 & 13.8 & 7.8e-03 & 1.4e-01 & 2.0e+01 & 27.3 & 3.1e-03 & 1.8e-01 & 1.4e+00 & 5.6e-04\\
-GW Lup     & 1 &  7.49 &  9.9 & 8.3e-03 & 1.5e-01 & 7.8e+00 &  9.8 & 8.0e-03 & 1.7e-01 & 3.4e-01 & 1.3e-03\\
-HD 143006  & 1 &  3.95$^{*}$ & 10.1 & 4.7e-03 & 8.6e-02 & 7.5e+01 & 167.1 & 8.4e-04 & 1.8e-01 & -- & 1.5e-04\\
-HD 143006  & 2 &  7.22$^{*}$ & 10.1 & 4.8e-03 & 8.8e-02 & 3.4e+01 & 73.1 & 1.9e-03 & 1.0e+00 & -- & 1.9e-03\\
+AS 209     & 1 &  5.59 & 19.6 & 1.1e-01 & 1.7e-01 & 1.6e+01 &  1.5 & 3.9e-03 & 3.3e-02 & 6.5e-01 & 1.3e-04\\
+AS 209     & 2 & 10.26 & 19.6 & 1.2e-01 & 1.9e-01 & 6.9e+00 &  0.6 & 9.1e-03 & 5.1e-02 & 2.2e-01 & 4.6e-04\\
+Elias 24   & 1 &  7.24 & 17.1 & 1.0e-01 & 1.5e-01 & 1.8e+01 &  1.8 & 3.6e-03 & 7.4e-02 & 6.3e-01 & 2.6e-04\\
+HD 163296  & 1 &  6.67$^{*}$ & 13.8 & 1.0e-01 & 1.6e-01 & 4.0e+01 &  4.1 & 1.6e-03 & 3.1e-01 & -- & 4.7e-04\\
+HD 163296  & 2 &  7.08 & 13.8 & 8.0e-02 & 1.2e-01 & 2.0e+01 &  2.7 & 3.1e-03 & 1.4e-01 & 8.6e-01 & 4.3e-04\\
+GW Lup     & 1 &  7.53 &  9.9 & 7.6e-02 & 1.2e-01 & 7.8e+00 &  1.1 & 8.1e-03 & 3.4e-01 & 7.8e-01 & 2.8e-03\\
+HD 143006  & 1 &  3.99$^{*}$ & 10.1 & 5.6e-02 & 8.4e-02 & 7.5e+01 & 14.0 & 8.3e-04 & 1.9e-01 & -- & 1.5e-04\\
+HD 143006  & 2 &  7.37$^{*}$ & 10.1 & 4.9e-02 & 7.4e-02 & 3.4e+01 &  7.0 & 1.9e-03 & 1.1e+00 & -- & 2.1e-03\\
 \hline
@@
 that. Depending on which of the two opacities we use (see Appendix
-\ref{sec-computing-dust-mass}) we arrive at different estimates of
+\ref{sec-computing-dust-mass}) {\bf TODO: OPACITIES AND IMPLICATIONS} we arrive at different estimates of
 $\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
@@
 keep $w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive any value for
-$\alpha_{\mathrm{turb}}/\mathrm{St}$. But for other rings we can compute
+$\alpha_{\mathrm{turb}}/\mathrm{St}$. But for other rings (those not marked
+with *) we can compute
 $\alpha_{\mathrm{turb}}/\mathrm{St}$. The resulting values for both choices of
@@
   2018ApJ...856..117F} {\bf [more?]}. However, it has to be kept in mind that
-the values of $\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ were derived
-for an example choice of parameters: : $w=w_{\mathrm{max}}$,
+the values of $\alpha_{\mathrm{turb}}=\alpha_{\mathrm{exmp}}$ were derived for
+an example choice of parameters: : $w=w_{\mathrm{max}}$,
 $\Sigma_{\mathrm{g}}=\Sigma_{\mathrm{g,max}}$ and $a=0.02\,\mathrm{cm}$. For a
 smaller value of $w$, a lower value of $\Sigma_{\mathrm{g}}$, or larger grains,
-the computed value of $\alpha_{\mathrm{turb}}$ will increase. So that it is hard
-to set a true upper limit on $\alpha_{\mathrm{turb}}$ from these observations.
+the computed value of $\alpha_{\mathrm{turb}}$ will increase. This means that it
+is hard to set a true upper limit on $\alpha_{\mathrm{turb}}$ from these
+observations.
 
@@
 confirm this, we need to dig a bit deeper by including size distributions
-(Section \ref{sec-model-with-grain-size-distribution}) and non-perfect
-(``leaky'') dust traps (Section {\bf XXXX}).
-
-{\bf We can make predictions for pressure slopes in CO}
-
-{\bf CONSIDER PRESSURE OFFSET!}
+(Section \ref{sec-model-with-grain-size-distribution}).
+
+\subsection{Non-perfect trapping: leakage of dust}
+\label{sec-leaky-traps}
+%
+So far we have assumed that the Gaussian pressure bump remains perfectly
+Gaussian well into the wings. This is obviously an unrealistic assumption. In
+practice, the bump will level off into a background pressure profile
+$p_{\mathrm{bg}}(r)$ that varies much slower with $r$ than the bump. In its
+simplest form we can investigate the following case:
+\begin{equation}
+  p(r) = \mathrm{max}\left(
+  p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right),
+  p_{\mathrm{bg}}(r)\right)
+\end{equation}
+which simply switches between the Gaussian bump of
+Eq.~(\ref{eq-gaussian-pressure-bump}) and the background profile
+$p_{\mathrm{bg}}(r)$, depending on which of the two is higher. Usually the
+background pressure profile has a negative slope: $dp_{\mathrm{bg}}(r)/dr<0$, so
+that any dust that reaches .. {\bf [NOT SURE IF THIS IS CORRECT]}
+
+To estimate
+the time it takes for the trapped dust to leak out
 
@@
 \begin{acknowledgements}
-  The authors acknowledge support
-  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
-  Datenverarbeitung of the University of T\"ubingen, the state of
-  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
-  through grant no INST 37/935-1 FUGG. This research was initiated at the
-  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
-  April 2018. Part of this work was also funded by the DFG Forschergruppe
-  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
+  The authors acknowledge support by the High Performance and Cloud Computing
+  Group at the Zentrum f\"ur Datenverarbeitung of the University of T\"ubingen,
+  the state of Baden-W\"urttemberg through bwHPC and the German Research
+  Foundation (DFG) through grant no INST 37/935-1 FUGG. Part of this work was
+  also funded by the DFG Forschergruppe FOR 2634 ``Planet Formation Witnesses
+  and Probes: Transition Disks''.
 \end{acknowledgements}
@@
 \hline
-AS 209     & 1 & $  3.637_{-0.047}^{+0.048}$ & $ 74.129_{-0.066}^{+0.063}$ & $  3.885_{-0.090}^{+0.098}$  \\
-AS 209     & 2 & $  2.949_{-0.031}^{+0.030}$ & $120.309_{-0.068}^{+0.065}$ & $  4.651_{-0.111}^{+0.115}$  \\
-Elias 24   & 1 & $  5.214_{-0.068}^{+0.066}$ & $ 76.851_{-0.092}^{+0.088}$ & $  5.104_{-0.172}^{+0.186}$  \\
-HD 163296  & 1 & $  8.475_{-0.056}^{+0.056}$ & $ 67.403_{-0.055}^{+0.057}$ & $  7.370_{-0.098}^{+0.098}$  \\
-HD 163296  & 2 & $  4.918_{-0.026}^{+0.027}$ & $ 99.862_{-0.061}^{+0.059}$ & $  5.741_{-0.116}^{+0.119}$  \\
-GW Lup     & 1 & $  1.467_{-0.052}^{+0.055}$ & $ 85.178_{-0.259}^{+0.316}$ & $  4.882_{-0.402}^{+0.530}$  \\
-HD 143006  & 1 & $  3.202_{-0.084}^{+0.082}$ & $ 41.080_{-0.169}^{+0.190}$ & $  5.133_{-0.320}^{+0.373}$  \\
-HD 143006  & 2 & $  2.473_{-0.043}^{+0.044}$ & $ 64.831_{-0.248}^{+0.230}$ & $  8.029_{-0.501}^{+0.593}$  \\
+AS 209     & 1 & $  3.663_{-0.045}^{+0.048}$ & $ 74.134_{-0.067}^{+0.069}$ & $  3.989_{-0.097}^{+0.106}$  \\
+AS 209     & 2 & $  2.981_{-0.031}^{+0.031}$ & $120.392_{-0.074}^{+0.073}$ & $  4.722_{-0.117}^{+0.129}$  \\
+Elias 24   & 1 & $  5.498_{-0.061}^{+0.060}$ & $ 76.650_{-0.072}^{+0.073}$ & $  4.923_{-0.122}^{+0.138}$  \\
+HD 163296  & 1 & $  8.749_{-0.073}^{+0.079}$ & $ 67.683_{-0.069}^{+0.071}$ & $  6.922_{-0.116}^{+0.120}$  \\
+HD 163296  & 2 & $  5.178_{-0.035}^{+0.033}$ & $ 99.963_{-0.059}^{+0.066}$ & $  5.161_{-0.114}^{+0.117}$  \\
+GW Lup     & 1 & $  1.292_{-0.036}^{+0.034}$ & $ 85.526_{-0.295}^{+0.399}$ & $  5.771_{-0.471}^{+0.597}$  \\
+HD 143006  & 1 & $  3.312_{-0.073}^{+0.075}$ & $ 40.996_{-0.193}^{+0.228}$ & $  5.087_{-0.319}^{+0.371}$  \\
+HD 143006  & 2 & $  2.577_{-0.051}^{+0.056}$ & $ 65.161_{-0.267}^{+0.243}$ & $  8.011_{-0.544}^{+0.672}$  \\
 \hline
@@
 \end{equation}
-where $\Delta\tau = \rho_d\, \kappa_\nu^{\mathrm{tot}} \Delta z$ and $b$ is
+where $\Delta\tau = \rho_d\, \kappa_\nu^{\mathrm{tot}} \Delta z$, and $b$ is
 \begin{equation}
@@
 where $\mu=\cos\theta$. We start at $\tau_\nu=-\tfrac{1}{2}\Delta \tau$ with
-$I_\nu=0$ and integrate to $\tau_\nu=+\tfrac{1}{2}\Delta \tau$. The resulting
+$I_\nu=0$, and integrate to $\tau_\nu=+\tfrac{1}{2}\Delta \tau$. The resulting
 $I_\nu^{\mathrm{out}}=I_\nu(\tfrac{1}{2}\Delta \tau)$ is the intensity that is
@@
 \begin{equation}\label{eq-v-radial-drift}
-  v_{\mathrm{dr}} = \frac{\mathrm{St}}{1+\mathrm{St}^{2}}
+  v_{\mathrm{dr}} = \frac{\mathrm{1}}{1+\mathrm{St}^{2}}v_{\mathrm{gr}} +
+  \frac{\mathrm{St}}{1+\mathrm{St}^{2}}
   \left(\frac{d\ln p}{d\ln r}\right)\frac{c_s^2}{\Omega_Kr}
@@
 \end{equation}
-where $t_{\mathrm{stop}}$ is the stopping time of the grains. The diffusion
+where $t_{\mathrm{stop}}$ is the stopping time of the grains. We assume that the
+gas radial velocity is zero: $v_{\mathrm{gr}}=0$, but we will briefly discuss
+below how the solution shifts slightly away from the peak of the pressure bump
+for $v_{\mathrm{gr}}\neq 0$.
+
+The diffusion
 coefficient for the dust is \citep{2007Icar..192..588Y}:
@@
 \end{equation}
-The normalization constant $\Sigma_{\mathrm{d0}}$ can be approximately expressed
-in terms of the total dust mass trapped in the pressure bump:
+As a side remark, we note that if $v_{\mathrm{gr}}$ is non-zero and
+inward-pointing, this solution shifts inward. We then replace $(r-r_0)$
+in Eq.~(\ref{eq-analytic-sol-radial-trapping}) by $(r-r_0-\delta r)$, with
+\begin{equation}
+  \delta r=\frac{w_d^2v_{\mathrm{gr}}}{D_{\mathrm{d}}(1+\mathrm{St}^2)}
+  \simeq \left(\frac{w}{h_p}\right)^2\left(\frac{v_{\mathrm{gr}}}{v_K}\right)
+  \frac{1}{\mathrm{St}}\,r_0
+\end{equation}
+One can see that this shift is independent of the width of the dust ring set by
+the turbulence. Note that in the above shift it is assumed that
+$v_{\mathrm{gr}}$ is constant across the pressure bump, which breaks mass
+conservation for the gas. The above treatment of $v_{\mathrm{gr}}\neq 0$ is
+therefore only a rough approximation. We will from here onward return to our
+assumption that $v_{\mathrm{gr}}=0$.
+
+The normalization constant $\Sigma_{\mathrm{d0}}$ in
+Eq.~(\ref{eq-analytic-sol-radial-trapping}) can be approximately expressed in
+terms of the total dust mass trapped in the pressure bump:
 \begin{equation}
@@
 \mathrm{Sc}\,\mathrm{St}$ (and equivalently $w_d\ll w$) is easily broken for
-small grains and/or non-weak turbulence. This is similar to the case for
+small grains and/or non-weak turbulence. In that case our assumption of
+a constant $\mathrm{St}$ becomes invalid. Dust will be turbulently mixed
+to distances $|r-r_0|\gtrsim w$, {\bf FINISH SENSTENCE}
+
+This is similar to the case for
 vertical settling and mixing: Small grains tend not to settle below a few
@@
 below which they are well mixed with the gas, and above which they are strongly
-depleted \citep{2004A&A...421.1075D}. \citet{2009A&A...496..597F} give an
-analytic solution to the settling-mixing problem (their Eq.~19) that reproduces
-the solutions of \citet{1995Icar..114..237D} for $\alpha_{\mathrm{turb}}\ll
+depleted \citep{2004A&A...421.1075D}. The reason for this is that the Stokes
+number $\mathrm{St}$ can only be assumed to be constant for $|r-r_0|\lesssim w$.
+For $|r-r_0|\gtrsim w$, the Stokes number (assuming we are in the Epstein limit)
+increases inversely proportional to the gas
+pressure. \citet{2009A&A...496..597F} give an analytic solution to the
+settling-mixing problem (their Eq.~19) that reproduces the solutions of
+\citet{1995Icar..114..237D} for $\alpha_{\mathrm{turb}}\ll
 \mathrm{Sc}\,\mathrm{St}$, but at the same time remains valid also for
 $\alpha_{\mathrm{turb}}\gtrsim \mathrm{Sc}\,\mathrm{St}$, as long as
-$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant with
-height above the midplane.
+$\alpha_{\mathrm{turb}}$ and $\mathrm{Sc}$ remain constant with height above the
+midplane, and the gas drag law remains in the Epstein regime.
 
@@
   \Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\Bigg[
-    -\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}}
+    -\frac{\mathrm{Sc}\,\mathrm{St}_0}{\alpha_{\mathrm{turb}}}
      \left(\exp\left(\frac{\Delta r^2}{2w^2}\right)-1\right)
@@
 \end{equation}
-The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) is valid for
-any value of $\alpha_{\mathrm{turb}}/\mathrm{Sc}\,\mathrm{St}$, as long as
-$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant along
-the radial width of the dust trap, and as long as $w\ll r_0$, to prevent
-geometric terms from the cilindrical coordinates from dominating.  One can
-easily verify that Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) reproduces
-the simpler Gaussian solution Eq.~(\ref{eq-analytic-sol-radial-trapping}) for
-$\Delta r\ll w$. One can also verify that for $\alpha_{\mathrm{turb}}\gg
-\mathrm{Sc}\,\mathrm{St}$ the shape of $\Sigma_{\mathrm{d}}(r)$ follows the
-shape of the gas pressure profile $p(r)$ (Eq.~\ref{eq-gaussian-pressure-bump-repeat}).
+and $\mathrm{St}_0$ is the value of the Stokes number at the peak of the
+pressure bump.  The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better})
+is valid for any value of $\alpha_{\mathrm{turb}}/\mathrm{Sc}\,\mathrm{St}_0$,
+as long as $\alpha_{\mathrm{turb}}$ and $\mathrm{Sc}$ remain constant along the
+radial width of the dust trap, the grains remain in the Epstein regime, and
+$w\ll r_0$, to prevent geometric terms from the cilindrical coordinates
+from dominating.  One can easily verify that
+Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) reproduces the simpler
+Gaussian solution Eq.~(\ref{eq-analytic-sol-radial-trapping}) for $\Delta r\ll
+w$. One can also verify that for $\alpha_{\mathrm{turb}}\gg
+\mathrm{Sc}\,\mathrm{St}_0$ the shape of $\Sigma_{\mathrm{d}}(r)$ follows the
+shape of the gas pressure profile $p(r)$
+(Eq.~\ref{eq-gaussian-pressure-bump-repeat}).
 
@@
 can be easily mistaken for a Gaussian profile with a narrower width than
-both the gas bump and the simple analytic solution.
+both the gas bump and the simple analytic solution. {\bf REWRITE THIS}
 
